### 2017-09-16 manuscript/manuscript.Rmd
Upload files

---
title:  |
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author:
  - "Danilo Freire^[PhD Candidate, Department of Political Economy, King's College London. Email:   [danilofreire@gmail.com](mailto:danilofreire@gmail.com).] &nbsp; John Meadowcroft^[Senior Lecturer in Public Policy, Department of Political Economy, King's College London. Email:  [john.meadowcroft@kcl.ac.uk](mailto:john.meadowcroft@kcl.ac.uk).] &nbsp; David Skarbek^[Associate Professor, Department of Political Science, Brown University. Email: [david_skarbek@brown.edu](david_skarbek@brown.edu). Corresponding author.] &nbsp; Eugenia Guerrero^[Software Developer, Tempo UK.]"
date: "`r format(Sys.time(), '%B %d, %Y')`"
thanks: "Data and replication information are available at [http://github.com/danilofreire/pinochet-victims-data](http://github.com/danilofreire/pinochet-victims-data)."
abstract: "How did the Pinochet regime use deaths and disappearances after the violent military coup in September 1973? The coup resulted in more than three decades of violence as General Pinochet's government fought against left-wing movements for the sake of the 'national wellbeing.' While these efforts were significant, they were not uniformly distributed spatially or over time during his regime. This paper provides new data on observed variation of regime-based violence. It codes and compiles data from two volumes of the Report of the Chilean National Commission on Truth and Reconciliation, a systematic investigation of the Pinochet regime published in 1991. The data documents the initially high-level and then steep decline of these atrocities, the extent to which victims were targeted, and other demographic information about the victims. Our data also includes geographic information system data that allows researchers to map the spatial variation of atrocities over time."
keywords: "authoritarianism, Chile, human rights, Pinochet, truth commission"
fontsize: 12pt
monospace-url: no
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    template: template.latex
    latex_engine: pdflatex
---

# Introduction

\setlength{\parindent}{1cm}
\setlength{\parskip}{0pt}

On 11 September 1973, General Augusto Pinochet led a coup against the democratically elected socialist government of Chile and its leader President Salvador Allende. The coup was the beginning of a seventeen-year military dictatorship that ended when the regime conceded in 1990 to popular demands for a return to democratic government. The Pinochet regime was notable for the violence of the coup -- in which President Allende committed suicide with his own rifle as troops stormed the Presidential Palace -- and of the period that followed, in which there were extra-judicial executions, systematic torture and other human rights abuses against alleged enemies of the regime. These abuses took place while the regime undertook a rapid liberalization of the economy under the direction of a small group of US-trained economists -- the so-called Chicago Boys. The success of those reforms, and the relationship of the reforms to the violence of the regime, remains controversial [@silva1991technocrats; @valdes1995pinochet]. The allegation that two Nobel Prize-winning economists, Milton Friedman and F. A. Hayek, supported the regime's free market policies despite its appalling human rights record, is also a matter of ongoing controversy [@farrant2014can; @farrant2012preventing; @meadowcroft2014hayek].

While there have been many historical and military studies of the Pinochet regime, there has hitherto been no attempt to offer a systematic, quantitative analysis of the regime's actions. Mapping the acts of violence during this period may offer insight into the role of violence in the transition from democratic to autocratic government, in the maintenance of the regime's power, and in the quick transition from one economic model to another. Past studies in political science have focused on the violence associated with wars between nation states [@lacina2006declining; @rummel1997power], civil wars [@azam2002violence; @balcells2010rivalry; @kalyvas2006logic], and conflict with criminal organizations [@lessing2015logics; @snyder2009does]. This paper instead contributes to the emergent literature that maps out and theorizes about autocratic violence during transition periods.

Violence during transitions has characterized a large number of regimes, from General Franco's purges in Spain to the recent ousting of Hosni Mubarak in Egypt. This pattern is also common in Latin America. The transitions to autocratic rule in Brazil (1964) and Argentina (1976) lead to frequent human rights violations, and Alberto Fujimori's _autogolpe_ (1992) explicitly allowed the president to "[apply] drastic punishments" towards alleged terrorists [@dabene1997narcodemocracias, 97; @samtleben2013constitucion, 56]. Much of the studies on regime transitions have been historical and qualitative in nature [e.g. @guest1990behind; @o1986transitions; @skidmore1990politics]. On Chile in particular, previous works have focused on how human rights violations affects the political system [@angell2005elecciones; @huneeus2014democracia; @portales2000chile], how various forms of social resistance emerged [@adams2002art; @gomez2008memory], and the role transitional justice plays in modern Chilean history [@collins2009human; @correa1991dealing; @ensalaco2000chile; @grandin2005instruction]. Unfortunately, quantitative studies on these topics remain rather limited. @ferrara2014assessing [16] writes that quantitative evaluations of long-term impacts of transitional justice are still sparse, mainly "due to the lack of necessary data and the difficulty of measure these types of effects in a statistical manner." Researchers have tried to overcome these problems by resorting to news reports, but events-based data are subject to severe distortions such as selection bias, under-reporting of important events or over-reporting of minor ones [@brysk1994politics; @cook2017two; @poe2004decision].

This article aims to fill that gap by presenting new data based on the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991] that has documented the organized violence that took place in Chile during the time of the Pinochet regime. The Truth and Reconciliation Report describes the findings of a national committee established after the end of the Pinochet regime to record the murders and disappearances during this period. The eight-member commission was chaired by Raúl Retting, who had been Chilean Ambassador to Brazil at the time of the coup, and its report was published in February 1991 and an English translation in 1993. Including all Appendices, the English translation of the report is 907 pages long and lists a total of 2,706 people known or believed to have died at the hands of the regime. In most cases, there is some detail as to the personal characteristics of the victim, and where and when they died or disappeared, and in many cases there is information as to how that person died. The new data presented here is a quantitative presentation of the qualitative data gathered in the Truth and Reconciliation Report.

The data was collected by recording each incidence described in the report into a quantitative dataset. It was possible to code for the geographical location of the incidents, the sociological characteristics of the victim, the affiliation of the victim (where known), the type of violence that took place, whether the victim was interrogated, tortured or in some other way mistreated (if known), and who were the perpetrators of the violence. There are a large number of cases where one or more of the variables are unknown. This reflects the incomplete nature of the original report in the sense that much remains unknown about what happened to a large number of the victims of the regime. Nevertheless, we believe that the new dataset does provide important information about the violence that took place between 11 September 1973 and the end of the regime.

# The Dataset

Our dataset comprises 2,398 observations and 57 variables. Each observation corresponds to a victim of the Pinochet regime and every individual has received a unique identifier in the data. There are several variables describing personal information of the victims, such as first and last age, his or her age, gender, nationality, occupation, and political affiliation if available. The dataset also contains information about the methods employed by the perpetrators and geographical coordinates for a number of the incidents.

To provide some context, consider first the total amount and types of violence carried out during this period. The report distinguishes between different types of violence. _Deaths_ are cases where the Commission signals a definite and known death of the victim. _Disappearances_ are cases where the victims are presumed to be dead at the hands of government agents. Government agents are assumed to have disposed of their bodies in a covert fashion, making their retrieval impossible at the time of the publication of the Report. _Disappearance, information of death_ are cases which the Commission has information that signals that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [-@report1991 44]. _Unresolved_ cases are those where insufficient information or evidence is available.

From the report, we were able to identify 1,331 victims who were killed, 853 who disappeared, 108 cases of disappearances where some information about the victim was available, 93 unresolved cases, and 7 suicides (see Figure \ref{number-victim}).[^ambiguous] It is, of course, very likely that the people who disappeared were, in fact, killed. However, based on the report and our methodology, we can only determine that 65% of all of the cases led to a person being killed. Forty percent of the documented cases were instances of someone being disappeared.

[^ambiguous]: Two cases are ambiguously described in the Truth and Commission Report, so we have decided to treat them as missing data. They are Ruiter Enrique Correa Arce, a news stand owner accused of facilitating message exchanges between party leaders (id number 843), and Alonso Fernando Gahona Chavez, a communist leader of municipal workers of La Cisterna (id number 847).

![Number of Victims According to Type of Violence\label{number-victim}](number-victim.pdf){width=85%}

We coded all cases according to the type of weapon used. Excluding missing observations, the vast majority of cases, 1,138 of them (85.9%), indicate that the type of weapon used was a gun. School students and the military had the highest number of deaths caused by guns, 90.4% and 88.5%, respectively. Less frequently, the report indicates that violence was also sometimes committed with a knife, bomb, tear gas, and poison, and involved beatings, hangings, asphyxiation, intentional car crashes, and electrocution. Seven members of the military (15%) were killed with explosives, and many non-military members of government, eleven in total (10.6%), were also victims of bomb attacks.

For those victims whose age was listed (roughly 52% of the total victims), the mean age was about 29 years. Figure \ref{victim-age} shows the distribution of ages, with the bulk of victims being in the age ranges from 21 to 35 years. Secondary and university students were an important part of the opposition [@guzman2012students, 6; @huneeus2016oposicion, 251], and the data indeed shows they were also targeted by regime. The vast majority of victims – 96.5% – were males. Likewise, nearly all of the victims were Chilean nationals – about 99% of the cases.

![Victim Age\label{victim-age}](victim-age.pdf){width=85%}

The report often provides information about the profession of victims (see Figure \ref{victim-occup}). The largest percentage of stated occupations that we classified would count as blue collar jobs. This was followed by white collar occupations, non-military government jobs, and university students. Figure \ref{type-occup} provides more detail about the relationship between the victim's previous occupation and the type violence they suffered.

![Victim Occupations\label{victim-occup}](victim-occup.pdf){width=85%}

![Type of Violence and Victim Occupation\label{type-occup}](type-occup.pdf){width=85%}

Moving on from the demographic of the victims, our data provides information on the nature of the incidents. For example, some of the cases specifically refer to whether victims were interrogated or tortured. However, in the majority of cases this is not known. Perhaps surprisingly, though, in a majority of cases where information is provided, victims were not typically subject to interrogation or torture. In Tables \ref{tab1} and \ref{tab2} below, the data shows that in 561 cases where there is information about the occurrence of an interrogation, it occurred in 85 cases (only about 15%). Likewise, there were 284 cases of torture in the 795 cases with information about whether torture happened (about 36% of cases).

\newpage

\begin{center}
  \begin{table}[H]
  		\begin{minipage}{.5\textwidth}
  		        \centering
			\caption{Interrogation\label{tab1}}
			\begin{tabular}{l r r}
				\toprule       & Total         & Percentage   \\
				\midrule Yes   & 85            & 3.5          \\
				No             & 476           & 19.8         \\
				Unknown        & 1837          & 76.6         \\
				\textbf{Total} & \textbf{2398} & \textbf{100} \\
				\bottomrule
			\end{tabular}
		\end{minipage}
		\begin{minipage}{.5\textwidth}
			\centering
			\caption{Torture\label{tab2}}
			\begin{tabular}{l r r}
				\toprule       & Total         & Percentage   \\
				\midrule Yes   & 284           & 11.8         \\
				No             & 511           & 21.3         \\
				Unknown        & 1603          & 66.9         \\
				\textbf{Total} & \textbf{2398} & \textbf{100} \\
				\bottomrule
			\end{tabular}
		\end{minipage}
	\end{table}
\end{center}

Although it is not clear why there is such a pattern in the data, one may conjecture a few explanatory hypotheses. It seems reasonable to assume that the data is not missing completely at random, that is, missingness is correlated with some unknown covariate [@heitjan1996distinguishing; @lall2016multiple; @little2014statistical; @rubin1976inference]. One such factors may be that former members of the Pinochet regime refused to collaborate with the Chilean Truth Commission, and consequently some of the most sensitive cases might have gone missing [@aguero2002militares, 54; @esperanza2007casi, 123].

Another possibility is that because of the reconciliatory purpose of the Truth and Commission Report, cases that would otherwise be classified as torture were deliberately omitted from the final text. @peris2008contradicciones [76] writes that the commission limited itself to documenting only disappearances and homicides, and about 40,000 possible victims of the regime were not included in the official records. As @grandin2005instruction [55] argues, faced with the need to confront Chile's recent history, the Commission treated some historical facts with "lawyerly ambivalence," minimising the likelihood that a higher number of incidents would be regarded as torture.

A comparison with Guatemala's truth commission offers some evidence in this regard. Guatemala's Commission for Historical Clarification (CEH) was established in 1994 to investigate human rights abuses during the 36 year civil war that ravaged the country [@tomuschat2001clarification, 233]. Like its Chilean counterpart, the Guatemalan CEH could not subpoena witnesses or records nor "individualize responsibility" [@grandin2005instruction, 58], but as the Commission staff investigated the role of informal institutions in the conflict, it was able to describe different types of human rights abuses in further detail [@chapman2001truth; @tepperman2002truth].

Much of the literature on violence in civil wars focuses on whether insurgents are targeted selectively for violence or whether violence is indiscriminate [@balcells2010rivalry; @kalyvas2006logic]. Applied to the Chilean context, the Pinochet Regime could only use selective violence if it had relatively reliable information about who is opposed to the regime and if these people are active in an area where the regime has control. In contrast, in areas controled by the insurgencies and other opposition groups, the government would resort to indiscriminate violence as it would not be able to distinguish between supporters and combatants. Based on Table \ref{tab3}, we see that the report provides information on the (relatively small) number of documented cases where victims were denounced, summoned, listed, voluntarily reported, or confronted in their home. It seems likely that for any of these types of encounters to occur, the regime would require specific information to target the individual.

|                      |   Total  | Percentage |
|:---------------------|---------:|-----------:|
| Denounced            |       10 |        0.4 |
| Summoned             |       62 |        2.6 |
| Listed               |       25 |        1.0 |
| Voluntarily reported |       51 |        2.1 |
| At home              |       15 |        0.6 |
| Unspecified          |     2235 |       93.2 |     
| **Total**            | **2398** |    **100** |

Table: Targeted Victims\label{tab3}

# Changes over Time

Our dataset allows researchers to analyze how violence varied during different stages of the Pinochet regime. Figure \ref{trend} shows the frequency of government-perpetrated incidents per month from 1973 to 1990. The data shows us that the vast majority of the acts of violence happened in the first two months after the coup. This is in line with past and recent historical accounts, including the frequent allusions to the "Caravan of Death", a group formed by paramilitaries in October 1973 to torture and kill dissidents [@davis1990they, 459; @verdugo2001chile].

![Incidents per Month\label{trend}](trend.pdf){width=85%}

In September 1973 alone, the government committed 633 acts of violence against civilians, of which 406 the Commission signals a definite and known death. The scale continued in October, with 504 incidences and 317 certified deaths. The numbers decline to about 10% of those levels in the next months, and then recede further in 1974.

In absolute numbers, the Chilean dictatorship was not as violent as its Argentinian counterpart,[^argentina] but the figures are notably high when compared to other South American countries. The Brazilian military regime made 434 victims from 1964 to 1985; Paraguay had 424 victims of human rights abuse from 1954 to 1989; Bolivia had 350 from 1964 to 1982; and Uruguay reported about 300 incidences from 1973 to 1985 [@agenciabrasil2014ditadura; @elpais2014ditadura; @fsp2016argentina]. The distribution of events over time seem to differ as well. Despite eventual periods of higher intensity, violence in authoritarian Brazil stretched somewhat evenly during the regime [@agenciabrasil2014ditadura; @elpais2014ditadura]. The Strossner regime in Paraguay also constantly employed violence against dissidents [@cvj2008paraguay].

[^argentina]: The Argentinian government disputes the figure of 30,000 deaths and disappearances suggested by human rights organisations [@bbc2016argentina; @elpais2016argentina]. Ceferino Reato, a writer and journalist, indicates that there were 7,158 incidences during the regime, including 6,415 cases of disappearances and 743 victims of "summary executions" [@clarin2016argentina]. Jorge Rafael Videla, who ruled Argentina from 1976 to 1981, also states that the regime murdered "about 7 or 8 thousand people" as "the price of winning the war against subversives" [@clarin2012videla].

After disaggregating the data into occupational status, we see that all groups were affected in the early periods of the regime. The later years show more variation. Figure \ref{trend-occup} displays the previous time trend disaggregated by the victim profession. We present the results for six groups: white collar and blue collar workers, school and university students, members of the military, and non-military civil servants.

Violence against white and blue collar workers were concentrated in September to December 1973. One of the possible causes for this trend is the fact that main guerilla group operating at the time, the _Movimiento de Izquierda Revolucionaria_ (Movement of the Revolutionary Left, MIR), was heavily attacked in the weeks that followed the coup. In addition to that, members of so-called "war councils" attacked civilians in urban squatter settlements and factories [@valdivia2003terrorism, 187]. After this period, violence recedes and does not reach previous levels.

![Incidents per Month, by Occupation\label{trend-occup}](trend-occup.pdf)

Similarly, violence against school and university students were also concentrated in the early stages of the regime. Nevertheless, the data indicates that the military repressed a noticeable number of students in the mid-1980s, a period that coincides with the _jornadas de protesta nacional_. The _protestas_ were a series of massive social upheavals spurred by the economic crisis that affected Chile in the early 1980s [@nsa2016bombing; @tironi1986revuelta; @weinstein1989jovenes]. The most shocking incident involved Rodrigo Rojas and Carmen Quintana, a young photographer and an engineering student, who were burned alive by members of the military after participating in a national protest [@agosin1987generals; @cummings2015democracy, 58].

The data suggests that the state-sponsored violence extended to the military personnel. The case of general Alberto Bachelet illustrates this fact. In 1974, General Bachelet was imprisoned and tortured because of his refusal to support the military coup. His wife Angela Jeria and his daughter Michelle were also tortured before fleeing to exile [@baesa2011reconstruccion; @quiroga2008primera, 8]. Other dissidents from the armed forces had a similar fate, such as Major Luis Iván Lavanderos Lataste who was found dead in the War Academy in 1973 after being accused of supporting political prisoners; and General Carlos Prats González, former member of the Allende cabinet, who was murdered along with his wife in Argentina in 1974 [@carmona2013lataste; @huneeus2016regimen, 23; @paredes2004operacion].

Lastly, non-military members of the government were targeted less frequently after the consolidation of the regime, where 62 people were attacked. Whereas we do observe abuses against this category in the later years of the regime, with 4 cases in July 1989, we observe an average of less than 1 incidence per month.

# Changes across Space

We were able to match a roughly 81% of the victims with data on the location in latitude and longitude of the incident reported. This allows us to analyze the spatial variation of the human rights abuses. Figures \ref{map-continent} provide this information in maps. We can see that all types of violence were mostly concentrated around Santiago and neighbouring cities. The results make intuitive sense. Santiago has been Chile's most densely populated city since the colonial times and in 1975 the city was home to 3,186,000 inhabitants, or about 30% of the country's population [@un1976yearbook].

Nevertheless, the types of abuses present some significant spatial variation. Homicides and disappearances are more widespread across the country, and a few incidents occurred even beyond the Chilean borders. This gives further evidence to the scope of the Operation Condor, a region-wide military effort to track down, and in many instances eliminate, political opponents of right-wing regimes in South America [@dinges2005condor; @mcsherry2002tracking; @mcsherry2012predatory; @slack1996operation]. The operation targeted high-profile members of the Allende government, such as the aforementioned General Carlos Prats, and Orlando Letelier, who was killed in a car bomb in Washington, D.C. [@freed1980death]. Moreover, the Chilean secret police also attempted to assassinate important political figures such as Bernardo Leighton, a Christian Democratic Party politician, and Carlos Altamirano, a prominent Chilean socialist [@wilson1999prosecuting; @wright2007chilean].

![Spatial Distribution of Human Rights Incidents\label{map-continent}](map-continent.pdf){width=110%, height=110%}

We can disaggregate the data to analyze violence patterns in further detail. Figure \ref{map-occup} provides a simple example. We analyze the spatial distribution of government-sponsored violence clustered by the victim's occupation. Violence against housewives and ex-military were concentrated in the capital, but the regime attacked individuals throughout the country regardless of their profession. Blue and white collar workers, the most numerous groups in our dataset, provide a clear example of how violence was widespread throughout Chile.

Moreover, the graph highlights that the government actively tracked down individuals abroad, and this was not strongly associated with their professional or educational background. University students seem to have been a special target, with a number of human right violations taking place in Argentina. The reason why this particular group was subject to violence abroad remains an open question.

In contrast, the regime attacked school students and members of the military only within the Chilean borders. This makes intuitive sense. During the dictatorship, the military were very active in the country due to the "internal war" between the regime and rebel groups [@ensalaco2000chile; @valdivia2003terrorism]; whereas high school students frequently clashed with the police in the 1980s [@kornbluh2004pinochet; @weinstein1989jovenes].

![Spatial Distribution of Human Rights Incidents\label{map-occup}](map-occup.pdf){width=150%, height=150%}

# Discussion and Outlook

The database discussed in this article is the first quantitative presentation of the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], which described more than two thousand human rights abuses perpetrated by the Pinochet regime from 1973 to 1990. The dataset includes rich information about the victims, the geographical location of the incidents, and the specific data where the violations occurred. The graphs and maps shown in this article provide some preliminary results about the temporal and spatial variation of state-sponsored violence during the recent military period in Chile.

We believe the dataset opens new potential research avenues. First, scholars can evaluate the long-term impact of human rights violations on social polarisation, political outcomes, and economic development in Chile. As incidences included in the dataset can be clustered into states, cities, neighbourhoods, or any other spatial unit of analysis, researchers can perform fine-grained statistical analyzes at the sub national level. Moreover, sociologists and criminologists can explore whether the intensity of the human rights abuse correlate with pre- or post-regime levels of interpersonal violence or civil conflict.

Second, the dataset also enables researchers to better understand the inner dynamics of authoritarian governments, a topic still understudied in political science. By linking the incidence of human right abuses to changes in the Pinochet government and his support coalition, historians and political scientists can explore whether the violations affected levels of civilian or bureaucratic collaboration to the regime. Additional studies can also make use of small-$n$ comparative designs to understand the reasons why civil society collaborates or opposes repressive regimes in Latin America.

Finally, scholars can investigate the links between the connections between the international realm and domestic politics in repressive regimes. As more documents are being declassified by both the Chilean government and the American intelligent services, this is a promising area of research. Scholars may be able to test whether pressure from foreign governments, international organizations and NGOs had any influence over the levels of human rights abuses in Chile. In this regard, our data allow researchers to shed more light on the possible impact of international actors over repressive regimes. We hope our dataset is useful for scholars interested in these and other questions, and that the information it contains elicits hypotheses not only about the Pinochet era, but about authoritarian governments more generally.

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{6pt}

### 2017-09-16 manuscript/online-appendix.Rmd
Upload files

---
title:  |
        | **Data Codebook and Appendix for**
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author:
  - "Danilo Freire^[PhD Candidate, Department of Political Economy, King's College London. Email:   [danilofreire@gmail.com](mailto:danilofreire@gmail.com).] &nbsp; John Meadowcroft^[Senior Lecturer in Public Policy, Department of Political Economy, King's College London. Email:  [john.meadowcroft@kcl.ac.uk](mailto:john.meadowcroft@kcl.ac.uk).] &nbsp; David Skarbek^[Associate Professor, Department of Political Science, Brown University. Email: [david_skarbek@brown.edu](david_skarbek@brown.edu). Corresponding author.] &nbsp; Eugenia Guerrero^[Software Developer, Tempo UK.]"
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: How did the Pinochet regime use deaths and disappearances after the violent military coup in September 1973? The coup resulted in more than three decades of violence as General Pinochet's government fought against left-wing movements for the sake of the 'national wellbeing.' While these efforts were significant, they were not uniformly distributed spatially or over time during his regime. This paper provides new data on observed variation of regime-based violence. It codes and compiles data from two volumes of the Report of the Chilean National Commission on Truth and Reconciliation, a systematic investigation of the Pinochet regime published in 1991. The data documents the initially high-level and then steep decline of these atrocities, the extent to which victims were targeted, and other demographic information about the victims. Our data also includes geographic information system data that allows researchers to map the spatial variation of atrocities over time.
keywords: authoritarianism, Chile, human rights, Pinochet, truth commission
fontsize: 12pt
monospace-url: no
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: no
    keep_tex: no
    template: template.latex
    latex_engine: pdflatex
---

\newpage

# Introduction

\setlength{\parindent}{0pt}

This document contains the codebook, data and code necessary to replicate the results in "Deaths and Disappearances in the Pinochet Regime" by @freire2017pinochet-regime-data. The `dataset.csv` file includes the complete dataset. The original Excel file, also included in the paper [GitHub repository](https://github.com/danilofreire/pinochet-victims-data), has been pre-processed to reduce compatibility problems with other statistical software such as R or Stata. All data pre-processing steps are described below. The analyses were performed in the R statistical language, version 3.4.1 [@rstats2017].

# Background

The violent military coup in Chile in September 1973 resulted in over three decades of sustained violence, deaths, and forced disappearances as General Pinochet’s government fought against left-wing movements for the sake of the 'national wellbeing'. The fall of Pinochet’s regime and the traumatic consequences of his authoritarian government left Chile deeply divided. In order to move forward and avoid a repetition of the past, the Chilean government highlighted the need of an impartial and comprehensive report detailing the extent of the deaths and disappearances.

By compiling and analyzing the data from both volumes of the Report of the Chilean National Commission on Truth and Reconciliation, we seek to understand the logic and dynamics of the violence, deaths, and disappearances under Pinochet’s Chile.

# Sources

Report of the Chilean National Commission on Truth and Reconciliation, Vol. 1.

Report of the Chilean National Commission on Truth and Reconciliation, Vol. 2.

\newpage

# Variables

**Region**. Report of the Chilean National Commission on Truth and Reconciliation, Vol. 1.

**Case number**. Variable label: “Case number in the dataset”.

**Sub-case number**. Variable label: “Sub-case number in the dataset”.

**Merged ID**. Variable label: “Identification number in the dataset”. The variable ‘Merged ID’ was created so as to have a complete row of ID numbers recognized by the mapping software. The previous two variables label cases by number (e.g. 1098) and subdivide the cases within that number (e.g. 1098 A, 1098 B).

**Page**. Variable label: “Page related to the identification number in the dataset”.

**Start Date**. Variable label: “Start date of incident”.

**End Date**. Variable label: “End date of incident”. This variable attempts to pinpoint when the individual died, was found, or was last presumed to be seen.

**LastName**. Variable label: “Last name(s) of individual”.

**FirstName**. Variable label: “First name(s) of individual”.

**Minor**. Variable label: “Minor or not”. Categories:

* 0 - Adult
* 1 - Minor
* NA - Unspecified

**Age**. Variable label: “Age of individual”. Range: from 2 to 85 years old.

**Female**. Variable label: “Female”. Categories:

* 0 - Male
* 1 - Female
* NA - Unknown

**Nationality**. Variable label: “Country of nationality label”. Categories:

* Argentinean
* Austrian
* Bolivian
* Brazilian
* Chilean
* Dual British and Chilean
* Dual French and Chilean
* Dual Swiss and Chilean
* Ecuadorian
* French
* Mexican
* Spanish
* Uruguayan
* Vietnamese

**Occupation**. Variable label: “Identification number of individual’s occupation”. Categories:

* 1 - White collar
* 2 - Blue collar
* 3 - School student
* 4 - University student
* 5 - Military
* 6 - Non-military government
* 7 - Unemployed
* 8 - Unknown
* 9 - Ex-military
* 10 - Housewife

**Occupation (string)**. Variable label: “Description of individual’s occupation”.

\newpage

**Victim Affiliation** Variable label: “Individual’s political affiliation”. Categories:

* 1 - _Opposition (Allende/Left)_: this category encompasses individuals loyal to the ex-president Allende, as well as individuals who were affiliated or leaders of left-wing groups
* 2 - _Regime (Pinochet)_: this category encompasses individuals who belonged to the Pinochet
government.
* 3 - _Unknown_: this category refers to individuals whose political affiliation was unknown or
unspecified.
* 4 - _None_: this category pertains to individuals whom the report states had no known political
affiliation.

**Victim Affiliation (string)**. Variable label: “Description of individual’s political affiliation” This complements the previous variable by giving as much information as possible on the victims’ political affiliations. Categories:

* _Opposition_
    - Communist
        - Young Communist
    - Christian Democrat/Christian Left
    - FER-MIR
    - MAPU (Movimiento de Acción Popular Unitaria)
    - MIR (Revolutionary Left Movement)
    - National Party
    - Pro-Democracy party (PPD)
    - Revolutionary Radical Party
        - Revolutionary Radical Youth Party
    - Socialist
        - Young Socialist
* _Regime_. This variable encompasses individuals who belonged to the Pinochet government.

\newpage

**Violence**. Categories:

* 1 - Killed: cases where the Commission signals a definite and known death.
* 2 - Suicide
* 3 - Disappearance: cases where the victims are presumed to be dead at the hands of government
agents. Government agents are assumed to have disposed of their bodies in a covert fashion, making their retrieval impossible at the time of the publication of the Report.
* 4 - Disappearance, information of death: cases which the Commission has information that signals that “the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims’ mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion” (p.44).
* 5 - Unresolved: insufficient information or evidence

**Method**. Categories:

* 1 - Gun
* 2 - Knife
* 3 - Unknown
* 4 - Beaten
* 5 - Hung
* 6 - Denied medical treatment
* 7 - Acute loss of blood
* 8 - Asphyxiated
* 9 - Intentional car crash
* 10 - Cardio-respiratory arrest
* 11 - Torture
* 12 - Jumped
* 13 - Immersion
* 14 - Poisoned
* 15 - Bomb
* 16 - Burned
* 17 - Tear gas
* 18 - Electrocuted

**Interrogation**. Variable label: “Individual was interrogated”. Categories:

* 1 - Yes
* 2 - No
* 3 - Unknown

**Torture**. Variable label: “Individual was tortured”. Categories:

* 1 - Yes
* 2 - No
* 3 - Unknown

“For the purposes of this Convention, torture means any act by which severe pain or suffering, whether physical or mental, is intentionally inflicted on a person for such purposes as obtaining from him or a third person information or confession, punishing him for an act he or a third person has committed or is suspected of having committed, or intimidating or coercing him or a third person, or for any reason based on discrimination of any kind, when such pain or suffering is inflicted by or at the instigation of, or with the consent or acquiescence of a public official or other person acting in an official capacity. It does not include pain or suffering arising only from, inherent in or incidental to, lawful sanctions” (pp. 38-39).

“These agencies – the CNI, the DICOMAR, and the group called COVEMA – practiced torture. CNI’s use of torture was systematic but more selective than that of the DINA which tortured practically everyone who passed through some of its secret facilities. The main torture methods continued to be the use of electricity, especially on the sensitive parts of the body, all kinds of beatings, and plunging the person’s head down into the water to the brink of asphyxiation, and then doing it again/ There are also indications that DICOMAR and COVEMA practiced torture, even to the point of death, as this Commission has verified and as will be noted in the case material.” (p. 644)

**Mistreatment**. Variable label: “Individual was mistreated”.

**Mistreatment means**. Categories:

* 1 - Yes
* 2 - No
* 3 - Unknown

**Targeted**. Variable label: “Individual was targeted”. Categories:

* 1 - Denounced
* 2 - Summoned
* 3 - Listed
* 4 - Voluntarily reported
* 5 - At home
* 6 - Unspecified

**Press**. Variable label: “Individual’s disappearance or death was justified in the press”. Accounts of an individual’s disappearance or death often appeared in the press, but the Truth Report refutes many of these accounts on the basis that it covers up or legitimizes the violence stemming from Pinochet’s regime. For example, fictional gun battles were fabricated, legitimizing the arbitrary and often unprovoked deaths of those rebelling against the government. Categories:

* 1 - Yes
* 2 - Unspecified

**War tribunal**. Variable label: “Individual was subjected to a war tribunal”. Categories:

* 1 - Yes
* 2 - Unspecified

**Previous arrests**. Variable label: “Individual was previously arrested”. Categories:

* 0 - Reported to the authorities and was allowed to go free (no arrest)
* 1 - One arrest
* 2 - Two arrests
* 3 - Three arrests
* 4 - Four arrests

**Perpetrator affiliation**. Variable label: “Perpetrator’s political affiliation”. Categories:

* 1 - Opposition (Allende/Left): “private citizens who committed terrorist acts or other kinds of politically motivated attacks” (p.43)
* 2 - Regime (Pinochet):
* 3 - Unknown
* 4 - None

    - _‘Opposition’_ (1):
    - _‘Government agents’_ (2):
        - “In this report the Commission is offering as much information as it could obtain about actions committed by government agents [...]: when such information is available the report names the branch or branches said to have participated, and specifically the regiment, base, police precinct, garrison, or group from which the official forces came” (p.43)
        - “When the Commission was unable to obtain such information but did come to the conviction that the person was killed by, or disappeared in the hands of, government agents, it has stated so” (p.43)
    - When the perpetrator is _‘unknown’_ (3): “The Commission has not presumed that government agents were involved in the death of individuals, even when it is clear that they were killed by firearms and when there is every reason to believe that the motivation was political, unless there are some grounds for that judgment” (p.43). See also “References to private citizens” on p. 44
    - “When the Commission here refers to the perpetrators as private citizens acting for political reasons or pretexts, it does not always mean that these were people who were opposed to the military government. In some cases the political motives of such private citizens were quite the contrary, that is, they supported the government. In some of those cases, which will be narrated in the chapter on those killed during protest actions, the Commission does not rule out the possibility that such private citizens were really security agents in civilian dress”

**Perpetrator (string)**. Variable label: “Description of perpetrator’s political affiliation”.

**Locations**. Place. Categories:

* 1 - Home
* 2 - Work
* 3 - University
* 4 - In custody
* 5 - In public
* 6 - In hospital
* 7 - Unknown

**Location (string)**. Variable label: “Location where victim was spotted/reported to be seen”.

**Latitude**. Variable label: “Latitudinal coordinates pertaining to the string ‘location’ variable”.

**Longitude**. Variable label: “Longitudinal coordinates pertaining to the string ‘location’ variable”.

**Exactitude of coordinates**. Variable label: “Exactitude of coordinates pertaining to the string ‘location’ variable”. Categories:

* 0 - Exact
* 1 - Area/neighborhood level
* 2 - City/town level

_The determination of coordinates_. The compilation of the ‘Location’ variable was based completely on information given in the Truth Report. However, since this information was in a string format (e.g. intersection of Calle Grecia and Avenida Rosa), creating a new variable incorporating each location’s latitude and longitude was necessary to pursue further analysis of the trends in deaths and disappearances. The format chosen was decimal coordinates.

_Sources used for the determination of coordinates_

Three main websites were used to find the decimal coordinates pertaining to the string ‘Location’ variable:

* 1 - Latitude: http://latitude.to/. Latitude was mostly used for sourcing more inexact coordinates, such as the coordinates for cities or towns. It was also used to convert non-decimal coordinates into the decimal format.
* 2 - Wikimapia: http://wikimapia.org/. Wikimapia was mostly used for locating shantytowns, neighborhoods, military bases, or police stations that were not found on Google maps. Its strength is that it is an editable map, meaning that individual users can collaborate so as to produce a more comprehensive map (filled with local knowledge) than other standard maps such as Google maps.
* 3 - Google maps: http://maps.google.co.uk/. Google maps was useful for finding intersections of roads or precise addresses, as well as for searching for police stations, some military bases, and other governmental buildings.

\newpage

# Data Pre-Processing

## Load required packages
```{r, message=FALSE,warning=FALSE}
if (!require("RCurl")) {
        install.packages("RCurl")
}
if (!require("plyr")) {
        install.packages("plyr")
}
if (!require("tidyverse")) {
        install.packages("tidyverse")
}
if (!require("devtools")) {
        install.packages("devtools")
}
if (!require("narnia")) {
        install_github("njtierney/narnia")
}
if (!require("rnaturalearth")) {
        install.packages("rnaturalearth")
}

devtools::install_github("tidyverse/ggplot2")
library(ggplot2)
```

## Load data
```{r, message=FALSE,warning=FALSE,cache=TRUE, tidy=FALSE}
df <- readr::read_csv("https://goo.gl/F1tgk1")
```

## Data manipulation
```{r, message=FALSE,warning=FALSE,cache=FALSE}
# Convert names to lowercase
names(df) <- tolower(names(df))

# Rename variables
df <- plyr::rename(df,
                   c("merged id" = "merged_id",
                     "start date" = "start_date",
                     "end date" = "end_date",
                     "last name" = "last_name",
                     "first name" = "first_name",
                     "sex (c)" = "female",
                     "occupation (c)" = "occupation",
                     "occupation" = "occupation_2",                     
                     "victim affiliation (c)" = "victim_affiliation",
                     "victim affiliation" = "victim_affiliation2",
                     "violence (c)" = "violence",
                     "method (c)" = "method",
                     "interrogation (c)" = "interrogation",
                     "torture (c)" = "torture",
                     "war tribunal" = "war_tribunal",
                     "previous arrests (#)" = "number_previous_arrests",
                     "perpetrator affiliation (c)" = "perpetrator_affiliation",
                     "place-1" = "place_1",
                     "start location-1" = "start_location_1",
                     "latitude-1" = "latitude_1",
                     "longitude-1" = "longitude_1",
                     "exact coordinates-1" = "exact_coordinates_1",
                     "place-2" = "place_2",
                     "location-2" = "location_2",
                     "latitude-2" = "latitude_2",
                     "longitude-2" = "longitude_2",
                     "exact coordinates-2" = "exact_coordinates_2",
                     "place ©-3" = "place_3",
                     "end location-3" = "end_location_3",
                     "latitude-3" = "latitude_3",
                     "longitude-3" = "longitude_3",
                     "exact coordinates-3" = "exact_coordinates_3",
                     "place (c)-4" = "place_4",
                     "end location-4" = "end_location_4",
                     "latitude-4" = "latitude_4",
                     "longitude-4" = "longitude_4",
                     "exact coordinates-4" = "exact_coordinates_4",
                     "place (c)-5" = "place_5",
                     "end location-5" = "end_location_5",
                     "latitude-5" = "latitude_5",
                     "longitude-5" = "longitude_5",
                     "exact coordinates-5" = "exact_coordinates_5",
                     "place (c)-6" = "place_6",
                     "end location-6" = "end_location_6",
                     "latitude-6" = "latitude_6",
                     "longitude-6" = "longitude_6",
                     "exact coordinates-6" = "exact_coordinates_6"
                     ))

# Transform variables
df <- transform(df,
                merged_id = as.integer(merged_id),
                start_date = as.Date(start_date, "%d/%m/%Y"),
                end_date = as.Date(end_date, "%d/%m/%Y"),
                minor = as.factor(minor),
                age = as.numeric(age),
                female = female - 1,
                violence = as.integer(violence),
                method = as.integer(method),
                number_previous_arrests = as.integer(number_previous_arrests),
                perpetrator_affiliation = as.integer(perpetrator_affiliation),
                place_1 = as.integer(place_1),
                exact_coordinates_1 = as.numeric(exact_coordinates_1),
                exact_coordinates_2 = as.numeric(exact_coordinates_2),
                exact_coordinates_3 = as.numeric(exact_coordinates_3),
                exact_coordinates_4 = as.numeric(exact_coordinates_4),
                exact_coordinates_5 = as.numeric(exact_coordinates_5)
                )

# Recode torture and interrogation
df$torture <- ifelse(df$torture > 2 | is.na(df$torture), "NA",
                     ifelse(df$torture == 2, 0, df$torture))
df$interrogation <- ifelse(df$interrogation > 2 | is.na(df$interrogation), "NA",
                           ifelse(df$interrogation == 2, 0, df$interrogation))

# Revalue variables
df$minor <- plyr::revalue(df$minor, c("y" = 1, "n" = 0))

# Save data
readr::write_csv(df, "dataset.csv")
```

## Visualise Missing Data

```{r, warning=FALSE,error=FALSE,message=FALSE,fig.width=7, fig.height=7}
# Proportion of variables that contain any missing values
miss_var_pct(df)

# Number of missings in each variable
gg_miss_var(df)
```

\newpage

## Figure 1
```{r, warning=FALSE,error=FALSE,message=FALSE,cache=TRUE}
df1 <- df %>%
        filter(!is.na(violence)) %>%
        group_by(violence) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3)) %>%
        as.data.frame() %>%
        mutate(violence = factor(violence,
                                 labels = c("1" = "Killed",
                                            "2" = "Suicide",
                                            "3" = "Disappearance",
                                            "4" = "Disappearance\nwith Information",
                                            "5" = "Unresolved"))) %>%
        print()

ggplot(df1, aes(x = reorder(violence, -n), y = n)) +
        geom_histogram(stat = "identity") +
        scale_x_discrete("\nType of Violence") +
        geom_text(stat = "identity", aes(label = n),
                  vjust = -0.5, size = 2.6) +
        ylab("Number of Victims\n") +
        theme_minimal()  +
        theme(axis.text.y = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank())
```

\newpage

## Figure 2
```{r, warning=FALSE,error=FALSE,message=FALSE}
df2 <- df %>%
        select(age) %>%
        na.omit() %>%
        unlist() %>%
        as.numeric() %>%
        cut(breaks = c(0, 16, 20, 25, 30, 35, 40, 45, 50, 85)) %>%
        as.vector()

my_labels <- c("Under 16", "16-20", "20-25", "25-30", "30-35",
                "35-40", "40-45", "45-50", "Over 50")

ggplot(as.data.frame(df2), aes(x = df2)) +
        geom_histogram(stat = "count") +
        scale_x_discrete("", labels = my_labels) +
        geom_text(stat = "count", aes(label = ..count..),
                  vjust = -1, size = 2.6) +
        ylab("") +
        theme_minimal() +
        theme(axis.text.y = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank())
```

\newpage

## Figure 3
```{r, warning=FALSE,error=FALSE,message=FALSE, fig.width=8.75, fig.height=5}
df3 <- df %>%
        select(occupation) %>%
        filter(occupation %in% c(1:10)) %>%
        na.omit() %>%
        group_by(occupation) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3)) %>%
        as.data.frame() %>%
        mutate(occupation = factor(occupation,
                                   labels = c("White\nCollar", "Blue\nCollar",
                                              "School\nStudent",
                                              "University\nStudent", "Military",
                                              "Non-Military\nGovernment",
                                              "Unemployed", "Unknown", "Ex-Military",
                                              "Housewife")))

ggplot(df3, aes(x = reorder(occupation, -n), y = n)) +
        geom_histogram(stat = "identity") +
        scale_x_discrete("") +
        geom_text(stat = "identity", aes(label = n),
                  vjust = -0.5,  size = 2.6) +
        ylab("") +
        theme_minimal() +
        theme(axis.text.y = element_blank(),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              axis.text.x = element_text(angle = 45, hjust = 1, size = 11))
```

\newpage

## Tables 1, 2 and 3
```{r}
df %>%
        select(interrogation) %>%
        group_by(interrogation) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3))

df %>%
        select(torture) %>%
        group_by(torture) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3))

df %>%
        select(targeted) %>%
        group_by(targeted) %>%
        summarise(n = n()) %>%
        mutate(freq = round(n / sum(n), 3))

```

\newpage

## Figure 4
```{r, warning=FALSE,error=FALSE,message=FALSE, fig.width=8, fig.height=5.55}
df5 <- df

df5$female <- factor(df5$female,
                     levels = c(0, 1),
                     labels = c("Male", "Female"))

df5$violence <- factor(df5$violence,
                       levels = c(1, 2, 3, 4, 5),
                       labels = c("1" = "Killed",
                                  "2" = "Suicide",
                                  "3" = "Disappearance",
                                  "4" = "Disappearance\nwith Information",
                                  "5" = "Unresolved"))

par(mar = c(0, 0, 0, 0))
mosaicplot(female ~ violence, data = df5, las = 1, color = TRUE,
           xlab = "Gender", main = "", ylab = "Type of Violence")

table(df5$violence, df5$female)

round(prop.table(table(df5$violence, df5$female)), 2)
```

\newpage

## Figure 5
```{r, warning=FALSE,error=FALSE,message=FALSE}
df4 <- df %>%
        select(violence, start_date) %>%
        na.omit() %>%
        mutate(violence_dummy = ifelse(violence >= 1, 1, 0)) %>%
        mutate(day = format(start_date, "%d"),
               month = format(start_date, "%m"),
               year = format(start_date, "%Y")) %>%
        filter(year < 1992 & year > 1972) %>%
        group_by(year, month) %>%
        summarise(total = sum(violence_dummy)) %>%
        mutate(dates = paste(month, year, sep="-")) %>%
        mutate(day = "01", date = paste0(day, "-", dates),
               date = lubridate::parse_date_time(date, "dmY"))

ggplot(df4, aes(date, total)) +
        geom_line() +
        scale_x_datetime(name = "",
                         date_breaks = "15 months",
                         expand = c(0, 0),
                         date_labels = "%b\n%Y") +
        ylab("") +
        theme_minimal(base_size = 12) +
        theme(panel.grid.minor = element_blank(),
              axis.text.x = element_text(hjust = 1))
```

\newpage

## Figure 6
```{r}
df4 <- df %>%
        select(violence, start_date, occupation) %>%
        na.omit() %>%
        filter(occupation %in% c(1:6)) %>%
        mutate(violence_dummy = ifelse(violence >= 1, 1, 0)) %>%
        mutate(day = format(start_date, "%d"),
               month = format(start_date, "%m"),
               year = format(start_date, "%Y")) %>%
        filter(year < 1992 & year > 1972) %>%
        group_by(year, month, occupation) %>%
        summarise(total = sum(violence_dummy)) %>%
        mutate(dates = paste(month, year, sep="-")) %>%
        mutate(day = "01", date = paste0(day, "-", dates),
               date = lubridate::parse_date_time(date, "dmY")) %>%
        as.data.frame() %>%
        mutate(occupation = factor(occupation,
                                   labels = c("White\nCollar", "Blue\nCollar",
                                              "School\nStudent", "University\nStudent",
                                              "Military", "Non-Military\nGovernment")))

ggplot(df4, aes(date, total)) +
        geom_line() +
        scale_x_datetime(name = "",
                         date_breaks = "18 months",
                         expand = c(0, 0),
                         date_labels = "%b\n%Y") +
        ylab("") +
        theme_minimal(base_size = 12) +
        scale_y_continuous(expand = c(0, 0)) +
        theme(panel.grid.minor = element_blank(),
              axis.text.x = element_text(hjust = 1)) +
        facet_wrap( ~ occupation, scales = "free", ncol = 2)
```

\newpage

## Figure 7
```{r, warning=FALSE,error=FALSE,message=FALSE,cache=TRUE, fig.width=11,fig.height=7}
df7 <- df %>%
        select(violence, longitude_1, latitude_1) %>%
        na.omit()

df7$violence <- factor(df7$violence,
                       levels = c(1, 2, 3, 4, 5),
                       labels = c("1" = "Killed",
                                  "2" = "Suicide",
                                  "3" = "Disappearance",
                                  "4" = "Disappearance\nwith Information",
                                  "5" = "Unresolved"))

sf_data <- rnaturalearth::ne_countries(returnclass = "sf", continent = c("South America", "North America"))

ggplot() +
  geom_sf(data = sf_data, colour = "black", size = .2, fill = "#eeeeee") +
  theme_minimal() +
  geom_point(data = df7, aes(x = longitude_1, y = latitude_1,
                             color = factor(violence)),
             alpha = 0.4, size = 1) +
        theme(axis.title.x = element_blank()) +
        theme(axis.title.y = element_blank()) +
        theme(legend.position = "none") +
        facet_wrap(~ violence, ncol = 3) +
        theme(plot.margin = unit(c(0, 0, 0.5, 0), "cm"))
```

\newpage

## Figure 8
```{r, warning=FALSE,error=FALSE,message=FALSE,cache=TRUE,fig.width=11,fig.height=7}
df8 <- df %>%
        select(occupation, violence, longitude_1, latitude_1) %>%
        filter(occupation %in% c(1:10)) %>%
        na.omit() %>%
        group_by(occupation) %>%
        as.data.frame() %>%
        mutate(occupation = factor(occupation,
                                   labels = c("White\nCollar", "Blue\nCollar",
                                              "School\nStudent",
                                              "University\nStudent", "Military",
                                              "Non-Military\nGovernment",
                                              "Unemployed", "Unknown",
                                              "Ex-Military",
                                              "Housewife"))) %>%
        mutate(violence = ifelse(violence >= 1, 1, 0))

sf_data <- rnaturalearth::ne_countries(returnclass = "sf", continent = c("South America", "North America"))

ggplot() +
  geom_sf(data = sf_data, colour = "black", size = .2, fill = "#eeeeee") +
  theme_minimal() +
  geom_point(data = df8, aes(x = longitude_1, y = latitude_1,
                             color = factor(occupation)),
             alpha = 0.4, size = .8) +
        theme(axis.title.x = element_blank()) +
        theme(axis.title.y = element_blank()) +
        theme(legend.position = "none") +
        facet_wrap(~ occupation, ncol = 4) +
        theme(plot.margin = unit(c(0, 0, 0.5, 0), "cm"))
```

\newpage
```{r}
sessionInfo()
```

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{6pt}

### 2017-09-16 R/script.R

@@ -1,0 +1,365 @@
+###########################################################################
+############ Deaths and Disappearances in the Pinochet Regime #############
+### Danilo Freire, John Meadowcroft, David Skarbek and Eugenia Guerrero ###
+###########################################################################
+
+## Load required packages
+if (!require("plyr")) {
+        install.packages("plyr")
+}
+if (!require("tidyverse")) {
+        install.packages("tidyverse")
+}
+if (!require("devtools")) {
+        install.packages("devtools")
+}
+if (!require("narnia")) {
+        install_github("njtierney/narnia")
+}
+if (!require("rnaturalearth")) {
+        install.packages("rnaturalearth")
+}
+
+devtools::install_github("tidyverse/ggplot2")
+library(ggplot2)
+
+## Load data
+df <- readr::read_csv("https://goo.gl/F1tgk1")
+
+## Data manipulation
+# Convert names to lowercase
+names(df) <- tolower(names(df))
+
+# Rename variables
+df <- plyr::rename(df,
+                   c("merged id" = "merged_id",
+                     "start date" = "start_date",
+                     "end date" = "end_date",
+                     "last name" = "last_name",
+                     "first name" = "first_name",
+                     "sex (c)" = "female",
+                     "occupation (c)" = "occupation",
+                     "occupation" = "occupation_2",                     
+                     "victim affiliation (c)" = "victim_affiliation",
+                     "victim affiliation" = "victim_affiliation2",
+                     "violence (c)" = "violence",
+                     "method (c)" = "method",
+                     "interrogation (c)" = "interrogation",
+                     "torture (c)" = "torture",
+                     "war tribunal" = "war_tribunal",
+                     "previous arrests (#)" = "number_previous_arrests",
+                     "perpetrator affiliation (c)" = "perpetrator_affiliation",
+                     "place-1" = "place_1",
+                     "start location-1" = "start_location_1",
+                     "latitude-1" = "latitude_1",
+                     "longitude-1" = "longitude_1",
+                     "exact coordinates-1" = "exact_coordinates_1",
+                     "place-2" = "place_2",
+                     "location-2" = "location_2",
+                     "latitude-2" = "latitude_2",
+                     "longitude-2" = "longitude_2",
[…]

### 2019-03-08 manuscript/manuscript2.tex
Add new branch

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{amssymb,amsmath,amssymb}
\usepackage[stable]{footmisc}
\usepackage{lmodern}
\usepackage{libertine}
\usepackage[libertine]{newtxmath}
\usepackage[scale=.95]{inconsolata}
\usepackage[authoryear]{natbib}
\usepackage{babelbib}
\usepackage{booktabs, makecell, longtable}
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.0,0.0,0.55}
\setcitestyle{aysep={}} 
\usepackage{etoolbox}
\makeatletter
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
	 \NAT@nmfmt{\NAT@nm}%
	 \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
	 \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
	 \NAT@nmfmt{\NAT@nm}%
	 \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
   {\@citeb\@extra@b@citeb}%
	 \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
  {}{}
\makeatother
\usepackage{setspace}
\usepackage[top=2cm,bottom=2cm,left=2.2cm,right=2.2cm]{geometry}
\usepackage[backref,pagebackref]{hyperref}
\renewcommand*{\backref}[1]{}
\renewcommand*{\backrefalt}[4]{%
	\ifcase #1 (Not cited.)%
	\or        Cited on page~#2.%
	\else      Cited on pages~#2.%
	\fi}
\renewcommand{\backreftwosep}{ and~}
\renewcommand{\backreflastsep}{ and~}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usepackage{mathtools}
\usepackage{caption}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\cleanlookdateon
\exhyphenpenalty=1000
\hyphenpenalty=1000
\widowpenalty=10000
\clubpenalty=1000

\hypersetup{pdftitle={Deaths and Disappearances in the Pinochet Regime: A New Dataset},
        pdfauthor={Danilo Freire, John Meadowcroft, David Skarbek and Eugenia Guerrero},
        pdfsubject={political science},
        pdfkeywords={authoritarianism, Chile, human rights, Pinochet, truth commission},
		pdfborder={0 0 0},
		breaklinks=true,
		linkcolor=Mahogany,
		citecolor=Mahogany,
		urlcolor=darkblue,
		colorlinks=true}

\onehalfspacing

\title{Deaths and Disappearances in the Pinochet Regime:\\ A New Dataset\thanks{We thank Umberto Mignozzetti and Robert Myles McDonnell for his helpful comments on previous versions of this manuscript. The dataset and all replication information are available on GitHub at \href{http://github.com/danilofreire/death-disappearances-pinochet}{\texttt{http://github.com/danilofreire/death-disappearances-pinochet}}.}}

\author{Danilo Freire\thanks{Postdoctoral Research Associate, Political Theory Project, Brown University, 8 Fones Alley, Providence RI 02912; \href{http://danilofreire.github.io}{\texttt{http://danilofreire.github.io}},  \href{mailto:danilo\_freire@brown.edu}{\texttt{danilo\_freire@brown.edu}}. Corresponding author.} 
\and John Meadowcroft\thanks{Senior Lecturer in Public Policy, Department of Political Economy, King's College London, Bush House (North East Wing), 30 Aldwych, London UK WC2B 4BG; \href{https://johnmeadowcroft.net}{\texttt{https://johnmeadowcroft.net}},  \href{mailto:john.meadowcroft@kcl.ac.uk}{\texttt{john.meadowcroft@kcl.ac.uk}}.} 
\and David Skarbek\thanks{Associate Professor, Department of Political Science and Political Theory Project, Brown University, 8 Fones Alley, Providence RI 02912; \href{http://davidskarbek.com}{\texttt{http://davidskarbek.com}},  \href{mailto:david_skarbek@brown.edu}{\texttt{david\_skarbek@brown.edu}}.} 
\and Eugenia Guerrero\thanks{Software Developer, Tempo, 8 Greencoat Place, London UK SW1P 1PL.}
}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}
\noindent In September 1973, General Augusto Pinochet staged a military coup against the Chilean President Salvador Allende. The coup resulted in 17 years of violence in which the Pinochet regime persecuted left-wing groups and government dissidents. While these efforts were significant, the repression was not uniformly distributed across space or over time. In this paper, we present a dataset containing all documented cases of victimisation during the Pinochet regime and include new geocoded information about the repression episodes. In that regard, we expand the findings of the Report of the Chilean National Commission on Truth and Reconciliation (1991) and provide novel statistics about the temporal and spatial variation of state-sponsored violence during the Pinochet era. Our data document the initially high-level and then steep decline of these atrocities, the extent to which victims were targeted, and show how violence was largely concentrated in Chile's metropolitan areas. Our dataset allows scholars to test new hypotheses on the dynamics of regime transition and contributes to the emerging literature on micro-level dynamics of state violence.

 \vspace{.75cm}
 \noindent
 \textbf{Keywords}: authoritarianism, Chile, human rights, Pinochet, truth commission
\end{abstract}

\newpage

\section{Introduction}
\label{sec:intro}

\doublespacing

On 11 September 1973, General Augusto Pinochet led a coup against the democratically elected socialist government of Chile and its leader President Salvador Allende. The coup marked the beginning of a seventeen-year military dictatorship that severely curtailed civil and political liberties in the country. The Pinochet regime was notable for the violence of the coup and of the period that followed. Deposed President Allende committed suicide with his own rifle as troops stormed the Presidential Palace, and in the following months the new regime committed a high number of extra-judicial executions, systematic torture and other human rights abuses. These atrocities happened while the regime undertook a rapid liberalisation of the economy under the direction of a small group of US-trained economists -- the so-called Chicago Boys. The success of those reforms, and the relationship of the reforms to the violence of the regime, remains controversial \citep{silva1991technocrats,valdes1995pinochet}. The allegation that two Nobel Prize-winning economists, Milton Friedman and F. A. Hayek, supported the regime's free market policies despite its appalling human rights record, is also a matter of ongoing controversy \citep{farrant2014can,farrant2012preventing,meadowcroft2014hayek}.

In this article, we present new individual-level data on the deaths and disappearances of the Pinochet regime. Our dataset updates and expands the information available in the 1128-page long Report of the Chilean National Commission on Truth and Reconciliation (\citeyear{report1991}), which documented the organised violence that took place in Chile under Pinochet's rule.\footnote{The report is available at \href{https://www.usip.org/publications/1990/05/truth-commission-chile-90}{\texttt{https://www.usip.org/publications/1990/05/truth-commission-chile-90}}.} We make three contributions to the original data. First, we code all textual information about the victims and perpetrators into a spreadsheet that can be easily imported into any statistical software. Scholars are thus able to query data previously available only in print format, such as the victims' sociological characteristics, their affiliation (where known), the type of violence that took place, whether the victim was interrogated, tortured or in some other way mistreated (if known), and who were the perpetrators of the violence. Second, we add new information about political partisanship, an important predictor of victimisation in the Pinochet regime.\footnote{The law decree number 77 (27/03/1973), issued only 16 days after the coup d'état, formally banned all left-wing groups associated with former President Salvador Allende. The text explicitly stated that ``[\dots] the Communist Party of Chile, the Socialist Party, the Socialist Popular Union, MAPU, Radicals, the Christian Left, the Independent Popular Action, the Popular Unity Party and all other entities, groups, factions or movements that support the Marxist doctrine shall be considered illegal.'' See: \href{https://www.leychile.cl/Navegar?idNorma=5730}{\texttt{https://www.leychile.cl/Navegar?idNorma=5730}}. Access: August 2018.} Also, we add the geographical location of all instances of repression in which such information was available. We have matched 87.5\% of the cases in our sample, or 2110 victims, to coordinates of latitude and longitude. These novel data allow researchers to investigate, with new levels of detail, how state repression varied over time and across space during the Chilean dictatorship.

Our dataset is part of a recent movement in comparative politics and international relations that emphasises the use of disaggregated data to understand the dynamics of conflict and cooperation \citep[e.g.,][]{cederman2009disaggregating, kalyvas2006logic, kalyvas2008microdynamics,lyall2010coethnics,raleigh2012violence}. More specifically, we contribute to the literature on state violence, regime transitions and authoritarian politics \citep[e.g.,][]{gandhi2008political,huneeus2014democracia,huneeus2016regimen,slater2010ordering}, which despite their many recent progresses, still contend with significant data limitations \citep[365]{art2012we}. Individual-level datasets allow scholars to test hypotheses about patterns of state repression, how they evolve over time, and what are their short and long-run consequences. For instance, disaggregated data have been used to analyse the incidence of homicides \citep{karstedt2012contextualizing}, lynchings \citep{godoy2002lynchings}, and sexual violence  \citep{sivakumaran2007sexual,wood2006variation,wood2009armed} in conflicts. Moreover, data from truth commissions can also help in understanding the impact of large-scale violence on state-building \citep{call2003democratisation}, societal reconciliation \citep{gibson2006contributions}, and mental health \citep{brouneus2010trauma, godoy2011something}. Similar efforts include the coding of the Liberian and Sierra Leone truth commission reports by the Human Rights Data Analysis Group \citep{hrgd2009liberia,hrgd2010sierra}, the estimation of human rights violations in Guatemala from 1960 to 1996 \citep{ball1999guatemala}, as well as the compilation of geocoded data from violence episodes in El Salvador \citep{mason2012elsalvador}.

Out data may also be valuable for scholars in area studies, specially those interested in Latin American politics. Chile is a particularly important case to study the logic of authoritarianism in the region. The Pinochet rule lasted for more than 15 years, was virtually unchallenged, and exerted considerable influence in Chile's democratisation process \citep{angell2005elecciones,barros2002constitutionalism,policzer2009rise, portales2000chile}.

This article proceeds as follows. In the following section, we give an overview of the dataset and describe its main variables. Section~\ref{sec:space} presents the spatial variation in repression by the Pinochet regime using our newly-collected geocoded data. Section \textit{XXXXX gives an empirical example}. Section~\ref{sec:limitations} discusses the potential limitations of the data and of statistical inference made with convenience samples. Section~\ref{sec:discussion} concludes and indicates possible avenues for further research.

\section{The Dataset}
\label{sec:data}

Our dataset comprises 2,399 observations and 57 variables. Each observation corresponds to one victim of the Pinochet regime and every individual has received a unique identifier. The data contain a wealth of personal information about the victims, such as their first and last names, his or her age, gender, nationality, occupation, and political affiliation if available. We have also included details about

the methods employed by the perpetrators and geographical coordinates for a number of the incidents.

To provide some context, consider first the total amount and types of violence carried out during this period. The report distinguishes between different types of violence. \textit{Deaths} are cases where the Commission signals a definite and known death of the victim. \textit{Disappearances} are cases where the victims are presumed to be dead at the hands of government agents. Government agents are assumed to have disposed of their bodies in a covert fashion, making their retrieval impossible at the time of the publication of the Report. \textit{Disappearance, information of death} are cases which the Commission has information that signals that ``the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion'' 
\citep[44]{report1991}. \textit{Unresolved} cases are those where insufficient information or evidence is available.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=.9\textwidth, height=10cm]{fig1.eps}
    \caption{Types of Violence}
    \label{fig:number-victims}
\end{figure}

%%%% Add information about gender to the graph

From the report, we were able to identify 1,331 victims who were killed, 853 who disappeared, 108 cases of disappearances where some information about the victim was available, 93 unresolved cases, and 7 suicides.\footnote{Two cases are ambiguously described in the Truth and Commission Report, so we have decided to treat them as missing data. They are Ruiter Enrique Correa Arce, a news stand owner accused of facilitating message exchanges between party leaders (id number 843), and Alonso Fernando Gahona Chavez, a communist leader of municipal workers of La Cisterna (id number 847).} It is, of course, very likely that the people who disappeared were, in fact, killed. However, based on the report and our methodology, we can only determine that 65\% of all of the cases led to a person being killed. Forty percent of the documented cases were instances of someone being disappeared. Note too that the regime targeted males much more frequently than females. In our dataset, about 95\% of the total number of victims were men, and in no category they represent less than 90\% of the sample.

We coded all cases according to the type of weapon used. Excluding missing observations, the vast majority of cases, 1,138 of them (86\%), indicate that the type of weapon used was a gun. School students and the military had the highest number of deaths caused by guns, 90\% and 88\%, respectively. Less frequently, the report indicates that violence was also sometimes committed with a knife, bomb, tear gas, and poison, and involved beatings, hangings, asphyxiation, intentional car crashes, and electrocution. Seven members of the military (15\%) were killed with explosives, and many non-military members of government, eleven in total (11\%), were also victims of bomb attacks.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=.9\textwidth, height=10cm]{fig2.eps}
    \caption{Victims' Age Distribution}
    \label{fig:victim-age}
\end{figure}

%% Again, add gender information here

For those victims whose age was listed (roughly 52\% of the total victims), the mean age was about 29 years. Figure \ref{fig:victim-age} shows the distribution of ages, with the bulk of victims being in the age ranges from 21 to 35 years. Secondary and university students were an important part of the opposition \citep[251]{guzman2012students,huneeus2016oposicion}, and the data indeed shows they were also targeted by regime. The vast majority of victims -- 96.5\% -- were males. Likewise, nearly all of the victims were Chilean nationals -- about 99\% of the cases.

Moving on from the demographic of the victims, our data provides information on the nature of the incidents. For example, some of the cases specifically refer to whether victims were interrogated or tortured. However, in the majority of cases this is not known. Perhaps surprisingly, though, in a majority of cases where information is provided, victims were not typically subject to interrogation or torture. In Tables \ref{tab1} and \ref{tab2} below, the data shows that in 561 cases where there is information about the occurrence of an interrogation, it occurred in 85 cases (only about 15\%). Likewise, there were 284 cases of torture in the 795 cases with information about whether torture happened (about 36\% of cases).

\begin{center}
  \begin{table}[H]
  		\begin{minipage}{.5\textwidth}
  		        \centering
			\caption{Interrogation\label{tab1}}
			\begin{tabular}{l r r}
				\toprule       & Total         & Percentage   \\
				\midrule Yes   & 85            & 3.5          \\
				No             & 476           & 19.8         \\
				Unknown        & 1837          & 76.6         \\
				\textbf{Total} & \textbf{2398} & \textbf{100} \\
				\bottomrule
			\end{tabular}
		\end{minipage}
		\begin{minipage}{.5\textwidth}
			\centering
			\caption{Torture\label{tab2}}
			\begin{tabular}{l r r}
				\toprule       & Total         & Percentage   \\
				\midrule Yes   & 284           & 11.8         \\
				No             & 511           & 21.3         \\
				Unknown        & 1603          & 66.9         \\
				\textbf{Total} & \textbf{2398} & \textbf{100} \\
				\bottomrule
			\end{tabular}
		\end{minipage}
	\end{table}
\end{center}

%%%% check the number of unknown cases in tables 1 and 2

Out dataset also has information about the political affiliation of XX\% of victims of the regime. The Pinochet regime followed strict ideological lines, thus political partisanship appears as a robust predictor of victimisation during that period. The repression against left-wing demonstrators was particularly prevalent in universities, where the government forces conducted what \citet[249]{constable1993nation} call a ``massive ideological purge''.

%%%%%%%%%% Insert graph with information about political affiliation

\newpage

\section{Variation Across Space}
\label{sec:space}

\section{Data Limitations}
\label{sec:limitations}

Although the data we present in this article constitute the most comprehensive account of the repression that took place during the Pinochet years, they also have some important limitations. Such limitations, however, are not unique to our data; in fact, most historical studies of political violence and authoritarian regimes suffer from the problems we describe below, like reporting bias, systematic missingness, and inconsistent information  \citep[e.g.,][]{art2012we,ball1999guatemala,kalyvas2006logic,king2009data,lustick1996history}. We discuss how these issues relates to our dataset and suggest some strategies scholars can adopt to minimise these concerns.

First, our dataset is a convenience sample, that is, it was not collected according a  probability-based selection method. In other words, the data are neither a random sample nor the full distribution of the events. Despite the wide-reaching scope of the Chilean Truth Commission, obtaining reliable information about repressive regimes is a difficult undertaking, and one cannot ensure that the report describes all violence episodes.

For instance, it is likely that the data are subject to report bias, as certain individuals may be more willing to report their victimisation that others. Moreover, former government officials who know about missing cases have few incentives to speak the truth if they will incriminate themselves or others by doing so. Also, human rights violations that occurred in urban areas are often overrepresented in the sample, as observers, journalists, legal practitioners and scholars are clustered in cities \citep{kalyvas2004urban}.

Finally, some observations may include conflicting information as the reports rely on individual recollections of the past. Historical memory can suffer from several types of bias, such as holding overly favourable views of the ingroup \citep{sahdra2007group}, ascribing positive attributes to one's choices \citep{mather2000choice}, using evidence to confirm one's prior beliefs \citep{nickerson1998confirmation}, or framing past events as predictable when they were not \citep{fischhoff2007early}.

Unfortunately, we cannot measure the extend of the selection bias without access to the underlying population distribution, what is impossible in our case. So we recommend authors proceed with some caution when conducting statistical inference with the data. However, there are ways of mitigating these problems. One suggestion is to use sensitivity tests to measure how large the selection bias should be to invalidate a given result, as proposed by \citet{blackwell2014selection}, \citet{cinelli2018making}, \citet{oster2017unobservable}, \citet{rosenbaum1983assessing}, and others. % This is the approach we adopt in this paper.

Another suggestion would be to use other sources of evidence to complement or verify the information available in the dataset. \citet{balcells2017rivalry} uses individual testimonies and archival records to analyse the Spanish Civil War; \citet{kalyvas2006logic} employs mixed methods to understand how groups use lethal force in internal conflicts;

\section{Discussion}
\label{sec:discussion}

The database discussed in this article is a quantitative presentation of the Report of the Chilean National Commission on Truth and Reconciliation (\citeyear{report1991}), which described more than two thousand human rights abuses perpetrated by the Pinochet regime from 1973 to 1990. The dataset includes rich information about the victims, the geographical location of the incidents, and the specific data where the violations occurred. The graphs and maps shown in this article provide some preliminary results about the temporal and spatial variation of state-sponsored violence during the recent military period in Chile.

As we discussed in the previous section, scholars should be cautious when using inferential methods with non-random samples, and it is important to describe to readers the limitations of such analyses. Nevertheless, we believe that the dataset opens new potential research avenues. First, scholars can evaluate the long-term impact of human rights violations on social polarisation, political outcomes, and economic development in Chile. As incidences included in the dataset can be clustered into states, cities, neighbourhoods, or any other spatial unit of analysis, researchers can perform fine-grained statistical analyses at the sub-national level. Moreover, sociologists and criminologists can explore whether the intensity of the human rights abuse correlate with pre- or post-regime levels of interpersonal violence or civil conflict.

Second, the dataset also enables researchers to better understand the inner dynamics of authoritarian governments, a topic still understudied in political science. By linking the incidence of human right abuses to changes in the Pinochet government and his support coalition, historians and political scientists can explore whether the violations affected levels of civilian or bureaucratic collaboration to the regime. Additional studies can also make use of small-$n$ comparative designs to understand the reasons why civil society collaborates or opposes repressive regimes in Latin America.

Finally, scholars can investigate the links between the connections between the international realm and domestic politics in repressive regimes. As more documents are being declassified by both the Chilean government and the American intelligent services, this is a promising area of research. Scholars may be able to test whether pressure from foreign governments, international organisations and NGOs had any influence over the levels of human rights abuses in Chile. In this regard, our data allow researchers to shed more light on the possible impact of international actors over repressive regimes. We hope our dataset is useful for scholars interested in these and other questions, and that the information it contains elicits hypotheses not only about the Pinochet era, but about authoritarian governments more generally.

\newpage

\bibliography{references}
\bibliographystyle{apalike}
\end{document}

### 2019-03-08 rap/article.Rmd
Add new branch

---
title:  |
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author: 
  - "Danilo Freire[^df]"
  - "John Meadowcroft[^jm]"
  - "David Skarbek[^ds]"
  - "Eugenia Guerrero[^eg]"
date: \today
thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments. Data and replication information are available at [`http://github.com/danilofreire/pinochet-dataset`](http://github.com/danilofreire/pinochet-dataset)."
abstract: "... However, . This paper presents a georeferenced event dataset on "
keywords: "authoritarian regimes, Chile, geocoded event data, Pinochet, truth commission"
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    template: template.latex
---

# Introduction\label{sec:intro}

\setlength{\parindent}{1cm}
\setlength{\parskip}{0pt}

# Conclusion: New Avenues for Research

In this paper, we discuss a quantitative presentation of the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], which described more than two thousand human rights abuses perpetrated by the Pinochet regime from 1973 to 1990.

The dataset includes rich information about more than 2,0000 victims of the Pinochet regime, the geographical location of the incidents, and the specific date when the violations occurred. The graphs and maps shown in this article provide some preliminary results about the temporal and spatial variation of state-sponsored violence during the recent military period in Chile.

We believe our data opens a number of potential research avenues. First, scholars can use the dataset to evaluate the long-term impacts of human rights violations on several outcomes of interest. For example, XXX and XXX highlight that state-led violence has enduring effects on social polarisation, political outcomes, and economic development. Researchers can test whether the Pinochet regime caused a similar impact on Chile's economy and society. Moreover, sociologists and criminologists may be interested in the relationship between the intensity of the human rights abuse and post-regime levels of social unrest or interpersonal violence.

Second, the dataset enables researchers to analyse the inner dynamics of authoritarian governments. This topic remains understudied in political science due to a lack of disaggregated information. By linking human right abuses to changes in Pinochet's support coalition, scholars can explore whether the violations affected levels of civilian or bureaucratic collaboration to the regime. XXXXXXXXXXXXX.

Finally, scholars can investigate the connections between international legitimacy and domestic politics in repressive regimes. This is a promising area of research as the Chilean government and American intelligence services continue to declassify documents from the Pinochet era. One relevant question is whether pressure from foreign governments and organisations had any influence over the levels of human rights abuses in Chile. We hope our dataset is useful for scholars interested in these and other questions, and that the information it contains elicits hypotheses not only about the Pinochet period, but about authoritarian governments more generally.

[^df]: Postdoctoral Research Fellow, The Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), [`http://danilofreire.github.io`](http://danilofreire.github.io). Corresponding author.

[^jm]: Senior Lecturer in Public Policy, Department of Political Economy, King's College London, [`john.meadowcroft@kcl.ac.uk`](mailto:john.meadowcroft@kcl.ac.uk), [`http://johnmeadowcroft.net`](http://johnmeadowcroft.net).

[^ds]: Associate Professor, Department of Political Science and the Political Theory Project, Brown University, [`david_skarbek@brown.edu`](david_skarbek@brown.edu), [`http://davidskarbek.com`](http://davidskarbek.com).

[^eg]: Software Developer, Tempo UK.

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{6pt}

### 2019-03-12 manuscript/article.Rmd
Add files via upload

--- 2019-03-08 rap/article.Rmd
@@
 monospace-url: yes
-spacing: double
+spacing: single
 papersize: a4paper
@@
 
-In this paper, we discuss a quantitative presentation of the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], which described more than two thousand human rights abuses perpetrated by the Pinochet regime from 1973 to 1990.
+The dataset we introduce in this paper describes more than 2,000 human rights violations perpetrated by the Pinochet regime from 1973 to 1990.
+We include rich information about the events, such as the specific dates when the abuses occurred, personal details of the victims, and the geographical locations of the incidents.
+In that regard, our dataset provides a quantitative presentation of, and an extension to, the findings of Report of the Chilean National Commission on Truth and Reconciliation [-@report1991].
+All information contained in the Report, plus our own spatial coordinates, are available in spreadsheet format which scholars can readily employ in their analyses.
 
-The dataset includes rich information about more than 2,0000 victims of the Pinochet regime, the geographical location of the incidents, and the specific date when the violations occurred.
-The graphs and maps shown in this article provide some preliminary results about the temporal and spatial variation of state-sponsored violence during the recent military period in Chile.
-
-We believe our data opens a number of potential research avenues.
-First, scholars can use the dataset to evaluate the long-term impacts of human rights violations on several outcomes of interest.
-For example, XXX and XXX highlight that state-led violence has enduring effects on social polarisation, political outcomes, and economic development.
-Researchers can test whether the Pinochet regime caused a similar impact on Chile's economy and society.
+Our data opens a number of potential avenues of research.
+For instance, @lupu2017legacy, @rozenas2017political and @zhukov2018stalin highlight that state repression has enduring effects on the political preferences and social attitudes of the victims.
+Researchers can test whether victims of the Pinochet regime show similar changes in political attitudes.
 Moreover, sociologists and criminologists may be interested in the relationship between the intensity of the human rights abuse and post-regime levels of social unrest or interpersonal violence.
 
-Second, the dataset enables researchers to analyse the inner dynamics of authoritarian governments.
-This topic remains understudied in political science due to a lack of disaggregated information.
+The dataset also enables researchers to examine how political coalitions affect the use of lethal violence in authoritarian regimes.
+While the topic has a long tradition in political science [e.g., @de2010leader; @gandhi2007authoritarian], only recently authors have
+
+The internal dynamics of autocratic governments remains an understudied topic in political science due to a lack of fine-grained information. But there are some noteworthy exceptions.
+
+@fjelde2010generals and @rivera2017authoritarian are noteworthy exceptions. Both authors note that
+
 By linking human right abuses to changes in Pinochet's support coalition, scholars can explore whether the violations affected levels of civilian or bureaucratic collaboration to the regime.

### 2019-03-19 rap/article.Rmd
update

@@
 
-In this paper, we discuss a quantitative presentation of the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], which described more than two thousand human rights abuses perpetrated by the Pinochet regime from 1973 to 1990.
+In this paper, we introduce a dataset with rich information about more than 2,000 human rights abuses perpetrated by the Pinochet regime from 1973 to 1990.
+We manually coded all information available in the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], and added new variables on the geographical location of the incidents and, whenever possible, the specific dates when each of the violations occurred.
+The graphs and maps included in this article provide some preliminary results about the temporal and spatial variation of state violence during Chile's last military government.
 
-The dataset includes rich information about more than 2,0000 victims of the Pinochet regime, the geographical location of the incidents, and the specific date when the violations occurred.
-The graphs and maps shown in this article provide some preliminary results about the temporal and spatial variation of state-sponsored violence during the recent military period in Chile.
+We believe our data open a number of potential research avenues.
+For instance, @lupu2017legacy, @rozenas2017political and @zhukov2018stalin highlight that state repression has enduring effects on the political preferences and social attitudes of the victims.
+Researchers can test whether victims of the Pinochet regime show similar changes in political behaviour.
+Moreover, sociologists and criminologists may be interested in the relationship between human rights abuses and post-regime levels of interpersonal violence.
+Recent studies have shown that democracies which arise after military regimes have higher homicide rates [@frantz2018legacy; @karstedt2006democracy], and our data can show whether areas with significant levels of military repression are also more violent today.
 
-We believe our data opens a number of potential research avenues.
-First, scholars can use the dataset to evaluate the long-term impacts of human rights violations on several outcomes of interest.
-For example, XXX and XXX highlight that state-led violence has enduring effects on social polarisation, political outcomes, and economic development.
-Researchers can test whether the Pinochet regime caused a similar impact on Chile's economy and society.
-Moreover, sociologists and criminologists may be interested in the relationship between the intensity of the human rights abuse and post-regime levels of social unrest or interpersonal violence.
+The dataset also enables researchers to examine how political coalitions affect the use of lethal violence in authoritarian regimes.
+Although the topic has received increasing attention in the last years [e.g., @fjelde2010generals; @gandhi2007authoritarian; @rivera2017authoritarian], the internal dynamics of autocratic governments remains understudied due to a lack of fine-grained information.
+By linking human right abuses to changes in Pinochet's support coalition, scholars can explore whether the civilian or bureaucratic collaboration to the regime impacted the number and type of violations.
+Since our data are presented at the individual level, they can be easily combined with legislative or bureaucratic records aggregated at any level.
 
-Second, the dataset enables researchers to analyse the inner dynamics of authoritarian governments.
-This topic remains understudied in political science due to a lack of disaggregated information.
-By linking human right abuses to changes in Pinochet's support coalition, scholars can explore whether the violations affected levels of civilian or bureaucratic collaboration to the regime.
-XXXXXXXXXXXXX.
+Qualitative scholars may find the rich personal details about the victims particularly useful.
 
-Finally, scholars can investigate the connections between international legitimacy and domestic politics in repressive regimes.
+Lastly, scholars can investigate the connections between international legitimacy and domestic politics in repressive regimes.
 This is a promising area of research as the Chilean government and American intelligence services continue to declassify documents from the Pinochet era.

### 2019-03-19 rap/article.Rmd
Update article.Rmd

@@
 
-In this paper, we introduce a dataset with rich information about more than 2,000 human rights abuses perpetrated by the Pinochet regime from 1973 to 1990.
-We manually coded all information available in the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], and added new variables on the geographical location of the incidents and, whenever possible, the specific dates when each of the violations occurred.
+In this paper, we introduce a dataset with rich information about more than 2,000 victims of the Pinochet regime.
+Our data come from two sources.
+First, we manually coded all information available in the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991].
+Second, we added the geographical locations and the specific dates of the human rights abuses whenever we could retrieve them.
 The graphs and maps included in this article provide some preliminary results about the temporal and spatial variation of state violence during Chile's last military government.
 
-We believe our data open a number of potential research avenues.
-For instance, @lupu2017legacy, @rozenas2017political and @zhukov2018stalin highlight that state repression has enduring effects on the political preferences and social attitudes of the victims.
-Researchers can test whether victims of the Pinochet regime show similar changes in political behaviour.
-Moreover, sociologists and criminologists may be interested in the relationship between human rights abuses and post-regime levels of interpersonal violence.
-Recent studies have shown that democracies which arise after military regimes have higher homicide rates [@frantz2018legacy; @karstedt2006democracy], and our data can show whether areas with significant levels of military repression are also more violent today.
+We believe our data open new topics of research.
+For instance, @lupu2017legacy, @rozenas2017political and @zhukov2018stalin highlight that state repression has enduring effects on political preferences and social attitudes.
+Researchers can test whether the Pinochet regime has caused similar attitudinal changes in direct or indirect victims.
+Moreover, sociologists and criminologists can analyse the relationship between human rights abuses and post-regime levels of interpersonal violence.
+Recent studies show that democracies which arise after military regimes have higher homicide rates [@frantz2018legacy; @karstedt2006democracy].
+Our data can show if areas with significant levels of military repression are more violent today.
 
-The dataset also enables researchers to examine how political coalitions affect the use of lethal violence in authoritarian regimes.
-Although the topic has received increasing attention in the last years [e.g., @fjelde2010generals; @gandhi2007authoritarian; @rivera2017authoritarian], the internal dynamics of autocratic governments remains understudied due to a lack of fine-grained information.
-By linking human right abuses to changes in Pinochet's support coalition, scholars can explore whether the civilian or bureaucratic collaboration to the regime impacted the number and type of violations.
-Since our data are presented at the individual level, they can be easily combined with legislative or bureaucratic records aggregated at any level.
+Researchers can also examine how political coalitions affect the use of lethal violence in authoritarian regimes.
+Although the topic has received increasing attention [e.g., @fjelde2010generals; @gandhi2007authoritarian; @rivera2017authoritarian], the internal dynamics of autocratic governments remains understudied.
+The main reason is a lack of fine-grained information [@ferrara2014assessing, 16].
+By linking human right abuses to changes in Pinochet's coalition, scholars can explore whether civilian or bureaucratic support lead to higher incidence of state violence.
+The individual data presented here can be combined with government records at any level of aggregation.
 
-Qualitative scholars may find the rich personal details about the victims particularly useful.
+Qualitative scholars will find the personal details of the victims to be particularly useful.
+Historians willing to reconstruct the biographies of specific individuals are able to access pre-compiled information in a single digital file.
+Others might be interested in using our data as a starting point for network analysis or to collect oral testimony from survivors and acquaintaces.
+In that regard, the dataset can accommodate future qualitative information.
+As we include a unique identification number to each victim, it is easy to update the personal record of any individual with new data from public archives or personal correspondence.
 

### 2019-03-20 rap/article.Rmd
upd8 conclusion
update conclusion, intro

@@
 thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments.
-Data and replication information are available at [`http://github.com/danilofreire/pinochet-dataset`](http://github.com/danilofreire/pinochet-dataset)."
+Data and replication information are available at [`http://github.com/danilofreire/pinochet-data`](http://github.com/danilofreire/pinochet-data)."
 abstract: "... However, . This paper presents a georeferenced event dataset on "
-keywords: "authoritarian regimes, Chile, geocoded event data, Pinochet, truth commission"
+keywords: "authoritarian regimes, Chile, georeferenced event data, Pinochet, truth commission"
 fontfamily: libertine
@@
 \setlength{\parskip}{0pt}
+
+On 11 September 1973, General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende.
+The coup marked the beginning of a seventeen-year military dictatorship that undertook a rapid liberalisation of the Chilean economy while perpetrating systematic violence against the opposition [@meadowcroft2014hayek; @valdes1995pinochet].
+The extent of the human rights abuses during the Pinochet regime remains unknown.
+However, in 1991 then-President Patricio Aylwin created a commission to investigate the most serious violations that happened during the military period.
+The Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], or the Rettig Report[^rettig], records over 2,000 cases of murders and disappearances.
+In most cases, there is some detail as to the personal characteristics of the victim, and where and when they died or disappeared, and in many cases there is information as to how that person died.
+The Report proved a major milestone in the pursuit of justice for the victims of the regime.
+
+Although the Report is a valuable source, quantitative scholars cannot easily use the rich information it contains.
+In this paper, we present a manually-coded dataset with all information from the Rettig Report plus new variables we constructed to complement the original data.
+We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each of the victims, then matched the location of the human rights abuses to geographic coordinates when we could verify that information.
+
+Apart from the geographical location of the incidents, our dataset also includes: 1) the sociological characteristics of the victim; 2) their affiliation (where known); 3) the type of violence that took place during that particular event; 4) whether the victim was interrogated, tortured or in some other way mistreated (if known); 5) who were the perpetrators of the violence.
+If the Report does not have a particular information, we coded it as missing.
+As each individual receives their own ID, new information can be added to the dataset as archival work continues.
+The next sections of this paper describe the dataset, show the descriptive statistics of some variables, and then suggest how our data can help answer future research questions.
+
+# Historical Background
+
+# The Dataset
+
+Our dataset comprises 2,398 observations and 57 variables.
+As we noted above, each observation corresponds to a victim of the Pinochet regime and every individual has a unique ID.
+There are several variables describing personal information of the victims, such as age, gender, nationality, occupation, and political affiliation if available.
+The dataset also includes information about torture or assassination methods and geographical coordinates for a number of the incidents.
+
+Users can download the data as an Excel spreadsheet (`.xlsx`) or as a comma-separated values (`.csv`) file.
+Both data and codebook are also available to download as an `R` package at <http://github.com/danilofreire/pinochet-data>.
+The repository has detailed installation instructions for users new to `R`.
+We follow the principles of "tidy data", where each column represents one variable and each row is one case [@wickham2014tidy].
+
+## Types of Violence
+
+The Report distinguishes between different types of violence carried out by the Pinochet regime.
+The first type is _deaths_.
+These are cases where the Commission signals a definite and known death of the victim.
+Take the case of Benito Heriberto Torres Torres, one of the first victims of the Pinochet regime (id number 2).
+Our dataset shows that Mr Torres a 57-years old male who worked as a plumber, and that he was tortured on the way to the 26th police station in Santiago on the 12 September 1973, just one day after the military coup.
+He was executed and his body was later found in Las Barrancas, a commune close to Santiago.
+
+ _Disappearances_ are cases where the victims are presumed to be dead at the hands of government agents.
+Government agents are assumed to have disposed of their bodies in a covert fashion, making their retrieval impossible at the time of the publication of the Report.
+_Disappearance, information of death_ are cases which the Commission has information that signals that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [-@report1991 44].
+_Unresolved_ cases are those where insufficient information or evidence is available.
+
+From the report, we identify 1,331 murders, 853 disappearances, 108 cases of disappearances where some information about the victim was available, 93 unresolved cases, and 7 suicides[^ambiguous].
+However, it is very likely that the people who disappeared were, in fact, killed.
+Based on the report and our methodology, we can only determine that 65% of all of the cases led to an assassination.
+Forty percent of the documented cases are disappearances.
+
+## Geocoding
+
+We georeferenced the events using the Google Maps API.
+The process is straightforward.
+The Truth and Commission Report often includes the name or approximate location of hospitals, police districts, houses, or other sites government agents used to torture or kill dissidents.
+We then
+
+## Variables and Patterns
 
@@
 Historians willing to reconstruct the biographies of specific individuals are able to access pre-compiled information in a single digital file.
-Others might be interested in using our data as a starting point for network analysis or to collect oral testimony from survivors and acquaintaces.
+Others might be interested in using our data as a starting point for network analysis or to collect oral testimony from survivors and acquaintances.
 In that regard, the dataset can accommodate future qualitative information.
@@
 
+[^rettig]: Former Chilean ambassador Raúl Rettig chaired the Chilean National Commission for Truth and Reconciliation.
+
+[^ambiguous]: Two cases are ambiguously described in the Truth and Commission Report, so we treat them as missing data.
+The victims are Ruiter Enrique Correa Arce, a news stand owner accused of facilitating message exchanges between party leaders (id number 843), and Alonso Fernando Gahona Chavez, a communist leader of municipal workers of La Cisterna (id number 847).
+
 \newpage

### 2019-03-21 rap/article.Rmd
update

@@
 As each individual receives their own ID, new information can be added to the dataset as archival work continues.
-The next sections of this paper describe the dataset, show the descriptive statistics of some variables, and then suggest how our data can help answer future research questions.
+In the next sections of this paper describe the dataset, we show summary statistics for some variables, then suggest how our data can help answer future research questions.
 
@@
 Take the case of Benito Heriberto Torres Torres, one of the first victims of the Pinochet regime (id number 2).
-Our dataset shows that Mr Torres a 57-years old male who worked as a plumber, and that he was tortured on the way to the 26th police station in Santiago on the 12 September 1973, just one day after the military coup.
-He was executed and his body was later found in Las Barrancas, a commune close to Santiago.
+Our dataset shows that Mr Torres was male, 57 years-old, and that he worked as a plumber.
+On the 12 September 1973, just one day after the military coup, Mr Torres was torture on the way to the 26th police station in Santiago.
+The records indicate he executed and his body was later found in Las Barrancas, Santiago.
+The dataset also shows that we obtained this information from pages 159-160 in the Truth Commission Report.
 
- _Disappearances_ are cases where the victims are presumed to be dead at the hands of government agents.
-Government agents are assumed to have disposed of their bodies in a covert fashion, making their retrieval impossible at the time of the publication of the Report.
-_Disappearance, information of death_ are cases which the Commission has information that signals that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [-@report1991 44].
+The second type of violence recorded in the dataset is _disappearances_.
+These are cases where government agents are presumed to have killed and disposed of the body of the victims.
+One such example is that of the Brazilian engineer Tulio Roberto Quintiliano Cardozo (id number 5).
+Mr Cardozo was a member of the Communist party and troops took him to the Military Academy for interrogation also on the 12 September 1971.
+He was never seen again and is presumed dead.
+
+The third category refers to _disappearance, information of death_. As the name implies,
+
+the Commission has information
+
+ that signals that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [-@report1991 44].
 _Unresolved_ cases are those where insufficient information or evidence is available.

### 2019-03-21 rap/article.Rmd
update types

@@
 The records indicate he executed and his body was later found in Las Barrancas, Santiago.
-The dataset also shows that we obtained this information from pages 159-160 in the Truth Commission Report.
+The dataset also shows that we obtained this information from pages 159-160 of the Truth Commission Report.
 
@@
 
-The third category refers to _disappearance, information of death_. As the name implies,
+The third category is _disappearance, information of death_.
+As the name implies, these observations refer to cases the Commission confirms the individual died after being missing.
+The formal definition for this category is that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [-@report1991 44].
+The assassination of Humberto de las Nieves Fuentes Rodriguez is an example (id number 854).
+He was taken in custody to the Colina air base, then loaded onto a helicopter with other political prisoners.
+According to the Report, government agents drugged him, beat him with a metal bar, opened his stomach with a knife before throwing the former Communist alderman from a helicopter.
+While his body has not been found, there is enough information about the case to classify his death as a severe human rights violation.
 
-the Commission has information
-
- that signals that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [-@report1991 44].
-_Unresolved_ cases are those where insufficient information or evidence is available.
-
-From the report, we identify 1,331 murders, 853 disappearances, 108 cases of disappearances where some information about the victim was available, 93 unresolved cases, and 7 suicides[^ambiguous].
+The last group is that of _unresolved_ cases.
+These are defined as those where insufficient information or evidence is available.
+Our dataset counts 93 unresolved cases, many with georeferenced information about possible torture sites[^ambiguous].
 However, it is very likely that the people who disappeared were, in fact, killed.

### 2019-03-24 rap/article.Rmd
change pdf viewer

@@
 thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments.
-Data and replication information are available at [`http://github.com/danilofreire/pinochet-data`](http://github.com/danilofreire/pinochet-data)."
+Data and replication information are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
 abstract: "... However, . This paper presents a georeferenced event dataset on "
@@
 Users can download the data as an Excel spreadsheet (`.xlsx`) or as a comma-separated values (`.csv`) file.
-Both data and codebook are also available to download as an `R` package at <http://github.com/danilofreire/pinochet-data>.
-The repository has detailed installation instructions for users new to `R`.
-We follow the principles of "tidy data", where each column represents one variable and each row is one case [@wickham2014tidy].
+We have also created an `R` package called `pinochet` which includes the dataset in both formats plus the codebook.
+The files follow the principles of "tidy data", where each column represents one variable and each row is one case [@wickham2014tidy].
+The `pinochet` package is available for download on the Comprehensive R Archive Network (CRAN) and at <http://github.com/danilofreire/pinochet>.
+To make the data accessible, our GitHub repository has detailed installation instructions for users new to `R`.
 
@@
 As the name implies, these observations refer to cases the Commission confirms the individual died after being missing.
-The formal definition for this category is that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [-@report1991 44].
+The formal definition for this category is that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [@report1991 44].
 The assassination of Humberto de las Nieves Fuentes Rodriguez is an example (id number 854).
@@
 
+[comment]: <> (Table 1 around here)
+
+Although the Report provides a comprehensive list of the regime victims, the Chilean Truth Commission itself acknowledges that the data are incomplete.
+The Commission had only nine months to write the Report and it had no legal power to oblige citizens to cooperate.
+Therefore, we expect that some cases have not been recorded.
+But this problem is not unique to the Pinochet regime.
+Selection bias is a common feature in events data, even more so in datasets about armed conflict or government repression [@ball2019using; @chapman2001truth].
+As the cases might not be a representative sample of the population of victims, we suggest scholars to carefully consider issues of model specification.
+
 ## Geocoding
+
+We have georeferenced all observations for which the relevant data was available.
 

### 2019-03-26 rap/article.Rmd
update

@@
----
 title:  |
@@
 
-On 11 September 1973, General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende.
+On 11 September 1973, Army General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende.
 The coup marked the beginning of a seventeen-year military dictatorship that undertook a rapid liberalisation of the Chilean economy while perpetrating systematic violence against the opposition [@meadowcroft2014hayek; @valdes1995pinochet].
@@
 
-Although the Report is a valuable source, quantitative scholars cannot easily use the rich information it contains.
-In this paper, we present a manually-coded dataset with all information from the Rettig Report plus new variables we constructed to complement the original data.
-We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each of the victims, then matched the location of the human rights abuses to geographic coordinates when we could verify that information.
+Although the Report is a valuable resource for scholars across the humanities, quantitative researchers cannot easily use the rich information it contains.
+In this paper, we present a manually-coded dataset with all information from the Truth Commission Report plus new variables we constructed to complement the original data.
+We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each of the victims, then matched the location of the human rights abuses to geographic coordinates.
+We matched all cases which had spatial information available.
 
@@
 If the Report does not have a particular information, we coded it as missing.
-As each individual receives their own ID, new information can be added to the dataset as archival work continues.
-In the next sections of this paper describe the dataset, we show summary statistics for some variables, then suggest how our data can help answer future research questions.
+As each individual receives their own id number, new information can be added to the dataset as archival work on the Pinochet era progresses.
+In the following sections, we describe the dataset, show summary statistics for selected variables, then provide suggestions about how our data can answer future research questions.
 
@@
 Our dataset comprises 2,398 observations and 57 variables.
-As we noted above, each observation corresponds to a victim of the Pinochet regime and every individual has a unique ID.
+As we noted above, each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number.
 There are several variables describing personal information of the victims, such as age, gender, nationality, occupation, and political affiliation if available.
 The dataset also includes information about torture or assassination methods and geographical coordinates for a number of the incidents.
+Figure 01 shows an example entry from our dataset.
+
+[comment]: <> (Example Entry)
 
@@
 We have georeferenced all observations for which the relevant data was available.
+Based on data from the Truth Commission, we first created a variable indicating where the victim was allegedly tortured or murder.
+We assigned a number to the six locations that most frequently appear in the report, namely: 1) home; 2) work; 3) university; 4) in custody; 5) in public; and 6) in hospital.
+We assigned "unknown" if the location could not be identified.
+Then we added latitude and longitude coordinates to each location where the individuals were seen.
 
-We georeferenced the events using the Google Maps API.
-The process is straightforward.
-The Truth and Commission Report often includes the name or approximate location of hospitals, police districts, houses, or other sites government agents used to torture or kill dissidents.
-We then
+Since the Truth Commission Report includes spatial data in string format -- e.g., intersection of _Calle Grecia_ with _Avenida Rosa_ -- we mapped the locations using decimal degrees to increase precision.
+The mapping process is straightforward.
+We used three different websites to increase the coverage of our maps and include the largest possible number of observations.
+The first website we employed is Google Maps (<http://maps.google.co.uk>), the world's most widely used map software.
+We queried Google Maps' application programming interface (API) to search for street intersections and other precise addressed we retrieved from the Truth Commission Report.
+Additionally, we used the API to locate police stations and government buildings still in use.
+
+Next, we used Wikimapia (<https://wikimapia.org/>), a open-content collaborative platform that allows users to mark locations worldwide.
+As of March 2019, Wikimapia contains about 30 million places in its database[^wikimapia].
+We benefited from this crowdsourced information to identify shanty-towns, military bases, and small neighbourhoods that could be easily found in general-purpose maps.
+
+Our third source of geographic coordinates is Latitude (<http://latitude.to>).
+We used Latitude to source inexact coordinates, such as coordinates for cities and towns.
+If no precise information was available, our measure of latitude and longitude correspond to those of the control point of each city.
+We also used the Latitude.to website to convert non-decimal coordinates to decimal ones.
+For instance, some addresses in Wikimapia are in degrees, minutes, and seconds (DMS) format, and to keep the dataset comparable all coordinates follow the same standard.
 
 ## Variables and Patterns
+
+As mentioned previously, our dataset includes 57 variables and covers the seventeen years of the Pinochet Presidency (1973--1990).
+Figure \ref{trend} shows the frequency of government-perpetrated incidents per month from 1973 to 1990.
+The data show us that the vast majority of the acts of violence happened in the first two months after the military coup.
+The high incidence of human rights violations in the beginning of the Pinochet era reflects the activity of the "Caravan of Death", a group of paramilitaries established in October 1973 to torture and kill dissidents [@davis1990they, 459; @verdugo2001chile].
+
+[comment]: <> (Figure 02 - Time Trends Here)
+
+In September 1973 alone, the government committed 633 acts of violence against civilians, of which 406 the Commission signals a definite and known death.
+The scale continued in October, with 504 incidences and 317 certified deaths.
+The numbers decline to about 10% of those levels in the next months, and then recede further in 1974.
+
+In absolute numbers, the Chilean dictatorship was not as violent as its Argentinian counterpart,[^argentina] but the figures are notably high when compared to other South American countries.
+The Brazilian military regime made 434 victims from 1964 to 1985; Paraguay had 424 victims of human rights abuse from 1954 to 1989; Bolivia had 350 from 1964 to 1982; and Uruguay reported about 300 incidences from 1973 to 1985 [@agenciabrasil2014ditadura; @elpais2014ditadura; @fsp2016argentina].
+The distribution of events over time seem to differ as well.
+Despite eventual periods of higher intensity, violence in authoritarian Brazil stretched somewhat evenly during the regime [@agenciabrasil2014ditadura; @elpais2014ditadura].
+The Stroessner regime in Paraguay also constantly employed violence against the opposition [@cvj2008paraguay].
 
@@
 
+[^wikimapia]: Wikimapia statistics for March 2019 are available at <https://wikimapia.org/stats/action_stats/?fstat=2&period=3&year=2019&month=3>.
+
+[^argentina]: The Argentinian government disputes the figure of 30,000 deaths and disappearances human rights organisations often suggest [@bbc2016argentina; @elpais2016argentina].
+Ceferino Reato, a writer and journalist, indicates that there were 7,158 incidences during the regime, including 6,415 cases of disappearances and 743 victims of summary executions [@clarin2016argentina].
+However, General Jorge Rafael Videla, who ruled Argentina from 1976 to 1981, seems to agree with those figures.
+He once remarked that killing "about 7 or 8 thousand people" was "the price of winning the war against the subversives" [@clarin2012videla].
+
 \newpage

### 2019-03-26 rap/article.Rmd
add variables

@@
+---
 title:  |
@@
 Figure \ref{trend} shows the frequency of government-perpetrated incidents per month from 1973 to 1990.
-The data show us that the vast majority of the acts of violence happened in the first two months after the military coup.
+We see that the vast majority of the acts of violence happened in the first two months after the military coup.
 The high incidence of human rights violations in the beginning of the Pinochet era reflects the activity of the "Caravan of Death", a group of paramilitaries established in October 1973 to torture and kill dissidents [@davis1990they, 459; @verdugo2001chile].
@@
 
-In September 1973 alone, the government committed 633 acts of violence against civilians, of which 406 the Commission signals a definite and known death.
-The scale continued in October, with 504 incidences and 317 certified deaths.
-The numbers decline to about 10% of those levels in the next months, and then recede further in 1974.
+In September 1973 alone, the government committed 633 acts of violence against civilians, of which 406 the Commission signals as "definite and known deaths".
+The violations continued in October, with 504 incidences and 317 certified deaths.
+The numbers decline to about 10% of those levels in November and December, and then recede further in 1974.
 
-In absolute numbers, the Chilean dictatorship was not as violent as its Argentinian counterpart,[^argentina] but the figures are notably high when compared to other South American countries.
-The Brazilian military regime made 434 victims from 1964 to 1985; Paraguay had 424 victims of human rights abuse from 1954 to 1989; Bolivia had 350 from 1964 to 1982; and Uruguay reported about 300 incidences from 1973 to 1985 [@agenciabrasil2014ditadura; @elpais2014ditadura; @fsp2016argentina].
-The distribution of events over time seem to differ as well.
-Despite eventual periods of higher intensity, violence in authoritarian Brazil stretched somewhat evenly during the regime [@agenciabrasil2014ditadura; @elpais2014ditadura].
-The Stroessner regime in Paraguay also constantly employed violence against the opposition [@cvj2008paraguay].
+In absolute numbers, the Chilean dictatorship was not as violent as its Argentinian counterpart[^argentina], but the figures are high when compared to other South American countries.
+For example, the Brazilian military regime made 434 victims from 1964 to 1985; Paraguay had 424 victims of human rights abuse from 1954 to 1989; and Uruguay reported about 300 incidences from 1973 to 1985 [@elpais2014ditadura; @fsp2016argentina].
+The timing of repression also varies.
+Despite eventual periods of higher intensity, violence in authoritarian Brazil and Paraguay stretched somewhat evenly during the regime [@agenciabrasil2014ditadura; @elpais2014ditadura].
+This contrasts with the Chilean case, whose repression was concentrated in the early years of the regime.
+
+We can also explore how the government employed violence across space.
+We see that all types of violence were concentrated around Santiago and neighbouring cities.
+The results are in line with theoretical expectations.
+Santiago has been Chile's most densely populated city since the colonial times and in 1975 the city was home to about 30% of the country's population [@un1976yearbook].
+
+[comment]: <> (Figure 03 - Spatial variation. Add a few more lines above)
 

### 2019-03-27 rap/article.Rmd
add

@@
 On 11 September 1973, Army General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende.
-The coup marked the beginning of a seventeen-year military dictatorship that undertook a rapid liberalisation of the Chilean economy while perpetrating systematic violence against the opposition [@meadowcroft2014hayek; @valdes1995pinochet].
-The extent of the human rights abuses during the Pinochet regime remains unknown.
-However, in 1991 then-President Patricio Aylwin created a commission to investigate the most serious violations that happened during the military period.
-The Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], or the Rettig Report[^rettig], records over 2,000 cases of murders and disappearances.
-In most cases, there is some detail as to the personal characteristics of the victim, and where and when they died or disappeared, and in many cases there is information as to how that person died.
-The Report proved a major milestone in the pursuit of justice for the victims of the regime.
+The coup marked the beginning of a seventeen-year military dictatorship which combined large-scale political repression with rapid economic liberalisation [@meadowcroft2014hayek; @valdes1995pinochet].
+In 1991, after the transition to democracy, then-President Patricio Aylwin created a commission to investigate human rights abuses in the Pinochet regime.
+The Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], records over 2,000 cases of murders and disappearances from 1973 to 1990.
+In most cases, there are some details as to the personal characteristics of the victim, and where and when they died or disappeared.
+Despite its omissions[^omissions].
 
@@
 
+[^omissions]: The Truth and Commission Report investigated only cases of death and disappearances, leaving aside a high number of other serious human rights violations committed by the Pinochet government.
+For instance, the Report makes no mention of cases of torture or forced exile, which comprise a significant share of the reported violations [@ensalaco1994truth, 660].
+
 [^ambiguous]: Two cases are ambiguously described in the Truth and Commission Report, so we treat them as missing data.

### 2019-03-28 rap/article.Rmd
more text

@@
 
-On 11 September 1973, Army General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende.
-The coup marked the beginning of a seventeen-year military dictatorship which combined large-scale political repression with rapid economic liberalisation [@meadowcroft2014hayek; @valdes1995pinochet].
-In 1991, after the transition to democracy, then-President Patricio Aylwin created a commission to investigate human rights abuses in the Pinochet regime.
-The Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], records over 2,000 cases of murders and disappearances from 1973 to 1990.
-In most cases, there are some details as to the personal characteristics of the victim, and where and when they died or disappeared.
-Despite its omissions[^omissions].
+On 11 September 1973, General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende.
+The coup marked the beginning of a seventeen-year military dictatorship which combined rapid economic liberalisation with large-scale political repression [@meadowcroft2014hayek; @valdes1995pinochet].
+In 1991, after Chile's transition to democracy, then-President Patricio Aylwin convened a commission to investigate human rights abuses in the Pinochet regime.
+The Report of the National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], recorded over 2,000 cases of murders and disappearances from 1973 to 1990.
+The Commission directly confronted the abuses of the past and portrayed a careful reconstruction of state repression during the military government.
+Most importantly, the Report contributed to the erosion of authoritarianism in Chile, as it fostered a series of criminal investigations that culminated with the indictment of General Pinochet in 2000 [@brahm2007uncovering 26].
 
-Although the Report is a valuable resource for scholars across the humanities, quantitative researchers cannot easily use the rich information it contains.
-In this paper, we present a manually-coded dataset with all information from the Truth Commission Report plus new variables we constructed to complement the original data.
-We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each of the victims, then matched the location of the human rights abuses to geographic coordinates.
-We matched all cases which had spatial information available.
+Although the Report is a valuable resource to scholars, quantitative researchers cannot readily use the rich information it contains.
+In this paper, we present a manually-coded dataset with all information from the Truth Commission Report plus new variables we collected to complement the original data.
+We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each victim, then georeferenced the location of the human rights abuses.
+We matched all cases that had spatial information available.
 
-Apart from the geographical location of the incidents, our dataset also includes: 1) the sociological characteristics of the victim; 2) their affiliation (where known); 3) the type of violence that took place during that particular event; 4) whether the victim was interrogated, tortured or in some other way mistreated (if known); 5) who were the perpetrators of the violence.
+Apart from the geographical location of the incidents, our dataset includes variables on: 1) the sociological characteristics of the victim; 2) their affiliation (where known); 3) the type of violence that took place during that particular event; 4) whether the victim was interrogated, tortured or in some other way mistreated (if known); 5) who were the perpetrators of the violence.
 If the Report does not have a particular information, we coded it as missing.
-As each individual receives their own id number, new information can be added to the dataset as archival work on the Pinochet era progresses.
-In the following sections, we describe the dataset, show summary statistics for selected variables, then provide suggestions about how our data can answer future research questions.
+As each individual receives their own id number, new information can be added to the dataset as archival work progresses.
+In the following sections, we describe our dataset, show summary statistics for selected variables, and suggest how our data can help answer future research questions.
 
@@
 Our dataset comprises 2,398 observations and 57 variables.
-As we noted above, each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number.
-There are several variables describing personal information of the victims, such as age, gender, nationality, occupation, and political affiliation if available.
-The dataset also includes information about torture or assassination methods and geographical coordinates for a number of the incidents.
+Each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number.
+There are several variables describing the personal information of the victims, such as age, gender, nationality, occupation, and political affiliation if available.
+The data also feature geographical coordinates for a number of the incidents, as well as information about methods of torture and causes of death.
 Figure 01 shows an example entry from our dataset.
@@
 
-Users can download the data as an Excel spreadsheet (`.xlsx`) or as a comma-separated values (`.csv`) file.
-We have also created an `R` package called `pinochet` which includes the dataset in both formats plus the codebook.
+Users can download the data as an Excel spreadsheet (`.xlsx`) or as a comma-separated values text file (`.csv`).
+We have also created an `R` package called `pinochet` which includes the dataset in both formats plus the data codebook.
 The files follow the principles of "tidy data", where each column represents one variable and each row is one case [@wickham2014tidy].
@@
 On the 12 September 1973, just one day after the military coup, Mr Torres was torture on the way to the 26th police station in Santiago.
-The records indicate he executed and his body was later found in Las Barrancas, Santiago.
+The records indicate he was executed and that his body was later found in Las Barrancas, Santiago.
 The dataset also shows that we obtained this information from pages 159-160 of the Truth Commission Report.
@@
 One such example is that of the Brazilian engineer Tulio Roberto Quintiliano Cardozo (id number 5).
-Mr Cardozo was a member of the Communist party and troops took him to the Military Academy for interrogation also on the 12 September 1971.
+Mr Cardozo was a member of the Communist party and troops took him to the Military Academy for interrogation also on the 12 September 1973.
 He was never seen again and is presumed dead.
@@
 As the name implies, these observations refer to cases the Commission confirms the individual died after being missing.
-The formal definition for this category is that "the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [@report1991 44].
+The formal definition for this category is the following: "the victims are dead; they died at the hands of the government agents, or persons in their service; and these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [@report1991 44].
 The assassination of Humberto de las Nieves Fuentes Rodriguez is an example (id number 854).
@@
 We can also explore how the government employed violence across space.
-We see that all types of violence were concentrated around Santiago and neighbouring cities.
-The results are in line with theoretical expectations.
+We see that assassinations are concentrated around Santiago and neighbouring cities.
+The results are expected.
 Santiago has been Chile's most densely populated city since the colonial times and in 1975 the city was home to about 30% of the country's population [@un1976yearbook].
+Disappearances follow a similar pattern.
+Although there are cases in the northern parts of the country, the military government targeted more victims in Santiago and neighbouring areas.
 
-[comment]: <> (Figure 03 - Spatial variation. Add a few more lines above)
+[comment]: <> (Figure 03 - Spatial variation)
+
+We can further disaggregate the analysis and visualise how the regime persecuted particular groups.
+Students and members of left-wing parties are two categories of special interest.
+On 18 September 1073, only one week after the coup, Pinochet and his followers rounded up thousands of students and activists in Chile's National Stadium.
+Military forces tortured and murdered hundreds of prisoners over the course of a few days.
+An emblematic victim was Víctor Lidio Jara Martínez, a university teacher and song composer.
+According to witnesses, Mr Jara had his hand crushed and his wrists broken, and he was later shot 44 times with a machine gun.
+Mr Jara's body was left in the front of the stadium as a warning for other political prisoners [@mcsherry2015victor 56].
+The government again tortured and killed students and activists in its campaign against the Revolutionary Left Movement (_Movimiento de Izquierda Revolucionaria_, MIR).
+The MIR originated in student unions and in the late 1970s the movement promoted a series of attacks against government personnel and official buildings.
+Government forces strongly retaliated and the group resorted to terrorist tactics to confront the regime [@schlotterbeck2018beyond].
+
+[comment]: <> (Figure 04 - Students and Communist Leaders)
 
@@
 
-[^df]: Postdoctoral Research Fellow, The Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), [`http://danilofreire.github.io`](http://danilofreire.github.io).
+[^df]: Postdoctoral Research Fellow, the Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), [`http://danilofreire.github.io`](http://danilofreire.github.io).
 Corresponding author.

### 2019-03-28 rap/article.Rmd
add text

@@
 Data and replication information are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
-abstract: "... However, . This paper presents a georeferenced event dataset on "
+abstract: "This article presents a georeferenced dataset on human rights abuses in the Pinochet regime (1973-1990).
+Our dataset is based on the Chilean Truth Commission and Reconciliation Report and includes 57 variables on 2,398 uniquely-identified victims.
+We manually coded all information available in the two volumes of the Report and added geographical coordinates to the atrocities that could be identified.
+The data describe the initially high-level then steep decline of these violations, the extent to which victims were targeted, and the spatial variation of state repression.
+We also discuss our coding methodology and introduce an accompanying R package named `pinochet` that contains the data files and codebook.
+Lastly, we conclude with suggestions for potential uses of our dataset and new avenues of research on authoritarianism."
 keywords: "authoritarian regimes, Chile, georeferenced event data, Pinochet, truth commission"

### 2019-03-29 rap/article.Rmd
update'

--- 2019-03-28 rap/article.Rmd
@@
 Data and replication information are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
-abstract: "... However, . This paper presents a georeferenced event dataset on "
+abstract: "This article presents a georeferenced dataset on human rights violations in General Augusto Pinochet's dictatorship in Chile.
+We manually coded the personal details of 2,398 victims mentioned in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations.
+Our dataset comprises 57 variables from 1973 to 1990 and it is available as a stand-alone spreadsheet or as the `pinochet` package for `R`.
+The data describe the initially high-level then steep decline of human rights violations, the extent to which the military government targeted the opposition, and the spatial variation of state repression.
+We also discuss our coding procedures, show descriptive statistics, and conclude with suggestions for potential uses of our dataset."
 keywords: "authoritarian regimes, Chile, georeferenced event data, Pinochet, truth commission"
@@
 The Report of the National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], recorded over 2,000 cases of murders and disappearances from 1973 to 1990.
-The Commission directly confronted the abuses of the past and portrayed a careful reconstruction of state repression during the military government.
-Most importantly, the Report contributed to the erosion of authoritarianism in Chile, as it fostered a series of criminal investigations that culminated with the indictment of General Pinochet in 2000 [@brahm2007uncovering 26].
+The Commission confronted the abuses of the past and portrayed a careful reconstruction of state repression during the military government.
+Most important, the Report contributed to the erosion of authoritarianism in Chile, as it fostered a series of criminal investigations that culminated with the indictment of General Pinochet in 2000 [@brahm2007uncovering 26].
 
@@
 We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each victim, then georeferenced the location of the human rights abuses.
-We matched all cases that had spatial information available.
+We mapped all locations that appear in the report, such as places used to torture, interrogate or murder the victims.
 
-Apart from the geographical location of the incidents, our dataset includes variables on: 1) the sociological characteristics of the victim; 2) their affiliation (where known); 3) the type of violence that took place during that particular event; 4) whether the victim was interrogated, tortured or in some other way mistreated (if known); 5) who were the perpetrators of the violence.
+Apart from the geographical location of the incidents, our dataset includes variables on: 1) the sociological characteristics of the victim; 2) their affiliation (where known); 3) the type of violence that took place during that event; 4) whether the victim was interrogated, tortured or in some other way mistreated (if known); 5) who were the perpetrators of the violence.
 If the Report does not have a particular information, we coded it as missing.
 As each individual receives their own id number, new information can be added to the dataset as archival work progresses.
-In the following sections, we describe our dataset, show summary statistics for selected variables, and suggest how our data can help answer future research questions.
+In the following sections, we give some background about the Pinochet regime, describe our dataset, show summary statistics for selected variables, and suggest how our data can help answer future research questions.
+
+\newpage
 
 # Historical Background
+
+Violence during transitions has characterised regime change in Latin America.
+Transitions to autocratic rule in Brazil (1964) and Argentina (1976) prompted frequent human rights violations that only recently have been fully documented [@agosin1987generals; @elpais2016argentina].
+In some cases, autocrats declared unrestrained violence to be an explicit government policy.
+Peru's former President Alberto Fujimori openly stated that his _autogolpe_ (1992) granted him the rights to "[apply] drastic punishments" towards alleged terrorists [@dabene1997narcodemocracias, 97; @samtleben2013constitucion, 56].
+
+The Pinochet regime is also notable for its brutality.
+On the day of the coup, the military opened fire against the _Palacio La Moneda_, the Presidential Office, and President Salvador Allende committed suicide after troops stormed the building [@verdugo2001chile].
+The violence continued during the military government.
+The _junta_ killed thousands of students and members of labour organisations in 1973, with up to 200,000 Chileans -- about 2% of the country's population -- led into exile [@wright2007chilean, 31].
+
+The establishment of the Truth and Reconciliation Commission was the first attempt of
 
@@
 The second type of violence recorded in the dataset is _disappearances_.
-These are cases where government agents are presumed to have killed and disposed of the body of the victims.
+These are cases where the Truth Commission presumes government agents to have killed and disposed of the body of the victims.
 One such example is that of the Brazilian engineer Tulio Roberto Quintiliano Cardozo (id number 5).
@@
 The assassination of Humberto de las Nieves Fuentes Rodriguez is an example (id number 854).
-He was taken in custody to the Colina air base, then loaded onto a helicopter with other political prisoners.
+The military took in custody to the Colina air base, then loaded onto a helicopter with other political prisoners.
 According to the Report, government agents drugged him, beat him with a metal bar, opened his stomach with a knife before throwing the former Communist alderman from a helicopter.
@@
 The last group is that of _unresolved_ cases.
-These are defined as those where insufficient information or evidence is available.
+The Report defines them as cases where insufficient information or evidence is available.
 Our dataset counts 93 unresolved cases, many with georeferenced information about possible torture sites[^ambiguous].
@@
 Selection bias is a common feature in events data, even more so in datasets about armed conflict or government repression [@ball2019using; @chapman2001truth].
-As the cases might not be a representative sample of the population of victims, we suggest scholars to carefully consider issues of model specification.
+As the cases might not be a representative sample of the population of victims, we suggest scholars to consider issues of model specification.
 
@@
 Based on data from the Truth Commission, we first created a variable indicating where the victim was allegedly tortured or murder.
-We assigned a number to the six locations that most frequently appear in the report, namely: 1) home; 2) work; 3) university; 4) in custody; 5) in public; and 6) in hospital.
-We assigned "unknown" if the location could not be identified.
-Then we added latitude and longitude coordinates to each location where the individuals were seen.
+We assigned a number to the six locations that most often appear in the report: 1) home; 2) work; 3) university; 4) in custody; 5) in public; and 6) in hospital.
+We assigned "unknown" if we could not identify the location.
+Then we added latitude and longitude coordinates to each place.
 
@@
 We used three different websites to increase the coverage of our maps and include the largest possible number of observations.
-The first website we employed is Google Maps (<http://maps.google.co.uk>), the world's most widely used map software.
+The first website we employed is Google Maps (<http://maps.google.co.uk>), the world's most popular map software.
 We queried Google Maps' application programming interface (API) to search for street intersections and other precise addressed we retrieved from the Truth Commission Report.
@@
 As of March 2019, Wikimapia contains about 30 million places in its database[^wikimapia].
-We benefited from this crowdsourced information to identify shanty-towns, military bases, and small neighbourhoods that could be easily found in general-purpose maps.
+We benefited from this crowdsourced information to identify shanty-towns, military bases, and small neighbourhoods that we could not find in general-purpose maps.
 
@@
 Figure \ref{trend} shows the frequency of government-perpetrated incidents per month from 1973 to 1990.
-We see that the vast majority of the acts of violence happened in the first two months after the military coup.
+We see that the majority of the acts of violence happened in the first two months after the military coup.
 The high incidence of human rights violations in the beginning of the Pinochet era reflects the activity of the "Caravan of Death", a group of paramilitaries established in October 1973 to torture and kill dissidents [@davis1990they, 459; @verdugo2001chile].
@@
 Military forces tortured and murdered hundreds of prisoners over the course of a few days.
-An emblematic victim was Víctor Lidio Jara Martínez, a university teacher and song composer.
+An emblematic victim was Víctor Lidio Jara Martínez, a university teacher, theatre director and singer (id number 25).
 According to witnesses, Mr Jara had his hand crushed and his wrists broken, and he was later shot 44 times with a machine gun.

### 2019-03-29 rap/article.tex
update'

--- 2019-03-29 rap/article.Rmd
@@
----
-title:  |
-        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
-author: 
-  - "Danilo Freire[^df]"
-  - "John Meadowcroft[^jm]"
-  - "David Skarbek[^ds]"
-  - "Eugenia Guerrero[^eg]"
-date: \today
-thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments.
-Data and replication information are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
-abstract: "This article presents a georeferenced dataset on human rights violations in General Augusto Pinochet's dictatorship in Chile.
-We manually coded the personal details of 2,398 victims mentioned in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations.
-Our dataset comprises 57 variables from 1973 to 1990 and it is available as a stand-alone spreadsheet or as the `pinochet` package for `R`.
-The data describe the initially high-level then steep decline of human rights violations, the extent to which the military government targeted the opposition, and the spatial variation of state repression.
-We also discuss our coding procedures, show descriptive statistics, and conclude with suggestions for potential uses of our dataset."
-keywords: "authoritarian regimes, Chile, georeferenced event data, Pinochet, truth commission"
-fontfamily: libertine
-fontawesome: yes
-fontsize: 12pt
-monospace-url: yes
-spacing: double
-papersize: a4paper
-bibliography: references.bib
-biblio-style: apalike
-output:
-  pdf_document:
-    citation_package: natbib
-    fig_caption: yes
-    number_sections: yes
-    keep_tex: no
-    template: template.latex
----
+\documentclass[12pt,a4paper,]{article}
+\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
+\usepackage[]{libertine}
+% If you're using libertine, this math font looks nice:
+\usepackage[libertine]{newtxmath}
+\usepackage[T1]{fontenc}
+\usepackage[scaled=.95]{inconsolata}
+% Add the following packages to support kableExtra
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{etoolbox}
+% Uncomment the line below if you want to use XeLaTeX or 
+% write on ShareLaTeX or Overleaf
+% \setmonofont[Scale=.9,BoldFont=inconsolata-bold.ttf]{inconsolata-regular.ttf}
+\usepackage[usenames,dvipsnames]{xcolor}
+\definecolor{darkblue}{rgb}{0.0,0.0,0.55}
+\usepackage{setspace}
+\usepackage[top=2cm,bottom=2cm,left=2.2cm,right=2.2cm]{geometry}
+\usepackage[backref,pagebackref]{hyperref}
+\usepackage{dcolumn}
+\usepackage{graphicx}
+\usepackage{float}
+\floatplacement{figure}{H}
+\usepackage{pgf}
+\usepackage{tikz}
+\usetikzlibrary{arrows}
+\usetikzlibrary{positioning}
+\usepackage{mathtools}
+\usepackage{caption}
+\usepackage{ifthen}
+\usepackage[UKenglish]{babel}
+\usepackage[UKenglish]{isodate}
+\cleanlookdateon
+\exhyphenpenalty=1000
+\hyphenpenalty=1000
+\widowpenalty=1000
+\clubpenalty=1000
+\renewcommand*{\backref}[1]{}
+\renewcommand*{\backrefalt}[4]{%
+    \ifcase #1 (Not cited.)%
+    \or        Cited on page~#2.%
+    \else      Cited on pages~#2.%
+    \fi}
+\renewcommand{\backreftwosep}{ and~} 
+\renewcommand{\backreflastsep}{ and~}
+\hypersetup{
+  linkcolor=Mahogany,
+  citecolor=Mahogany,
+  urlcolor=darkblue, 
+  breaklinks=true, 
+  colorlinks=true,
+      pdfauthor={Danilo Freire; John Meadowcroft; David Skarbek; Eugenia Guerrero},
+      pdfkeywords={authoritarian regimes, Chile, georeferenced event data, Pinochet, truth
+commission},
+  }
+\setlength{\parindent}{1cm}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{amssymb,amsmath}
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
 
-# Introduction\label{sec:intro}
+\usepackage{natbib}
+\bibliographystyle{apalike}
+\usepackage{etoolbox}
+\makeatletter
+% Remove comma after author
+\setcitestyle{aysep={}}
+\patchcmd{\NAT@citex}
+	  {\@citea\NAT@hyper@{%
+		 \NAT@nmfmt{\NAT@nm}%
+		 \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
+		 \NAT@date}}
+	  {\@citea\NAT@nmfmt{\NAT@nm}%
+	   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
+	\patchcmd{\NAT@citex}
+	  {\@citea\NAT@hyper@{%
+		 \NAT@nmfmt{\NAT@nm}%
+		 \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
+		   {\@citeb\@extra@b@citeb}%
+		 \NAT@date}}
+	  {\@citea\NAT@nmfmt{\NAT@nm}%
+	   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
+	  {}{}
+% Patch case where name and year are separated by aysep
+\patchcmd{\NAT@citex}
+  {\@citea\NAT@hyper@{%
+     \NAT@nmfmt{\NAT@nm}%
+     \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
+     \NAT@date}}
+  {\@citea\NAT@nmfmt{\NAT@nm}%
+   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
+% Patch case where name and year are separated by opening bracket
+\patchcmd{\NAT@citex}
+  {\@citea\NAT@hyper@{%
+     \NAT@nmfmt{\NAT@nm}%
+     \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
+       {\@citeb\@extra@b@citeb}%
+     \NAT@date}}
+  {\@citea\NAT@nmfmt{\NAT@nm}%
+   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
+  {}{}
+\makeatother
+  \IfFileExists{parskip.sty}{%
+ \usepackage{parskip}
+ }{% else
+ \setlength{\parindent}{0pt}
+ \setlength{\parskip}{0pt}
+ }
+  \setlength{\emergencystretch}{3em}  % prevent overfull lines
+ \providecommand{\tightlist}{%
+   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+% % % Redefines (sub)paragraphs to behave more like sections
+% \ifx\paragraph\undefined\else
+% \let\oldparagraph\paragraph
+% \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+% \fi
+% \ifx\subparagraph\undefined\else
+% \let\oldsubparagraph\subparagraph
+% \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+% \fi
+% 
+\doublespacing
+
+\title{\textbf{Deaths and Disappearances in the Pinochet Regime: A New Dataset}\thanks{We thank Umberto Mignozzetti and Robert Myles McDonnell for their
+helpful comments. Data and replication information are available at
+\href{http://github.com/danilofreire/pinochet}{\texttt{http://github.com/danilofreire/pinochet}}.}}
+\author{Danilo Freire\footnote{Postdoctoral Research Fellow, the Political
+  Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912,
+  \href{mailto:danilofreire@gmail.com}{\texttt{danilofreire@gmail.com}},
+  \href{http://danilofreire.github.io}{\texttt{http://danilofreire.github.io}}.
+  Corresponding author.} \and John Meadowcroft\footnote{Senior Lecturer in Public Policy, Department
+  of Political Economy, King's College London,
+  \href{mailto:john.meadowcroft@kcl.ac.uk}{\texttt{john.meadowcroft@kcl.ac.uk}},
+  \href{http://johnmeadowcroft.net}{\texttt{http://johnmeadowcroft.net}}.} \and David Skarbek\footnote{Associate Professor, Department of Political
+  Science and the Political Theory Project, Brown University,
+  \href{david_skarbek@brown.edu}{\texttt{david\_skarbek@brown.edu}},
+  \href{http://davidskarbek.com}{\texttt{http://davidskarbek.com}}.} \and Eugenia Guerrero\footnote{Software Developer, Tempo UK.}}
+\date{\today}
+
+\begin{document}
+
+\maketitle
+
+\begin{abstract}
+\noindent This article presents a georeferenced dataset on human rights violations
+in General Augusto Pinochet's dictatorship in Chile. We manually coded
+the personal details of 2,398 victims mentioned in the Chilean Truth
+Commission Report and added geographical coordinates for all
+identifiable atrocity locations. Our dataset comprises 57 variables from
+1973 to 1990 and it is available as a stand-alone spreadsheet or as the
+\texttt{pinochet} package for \texttt{R}. The data describe the
+initially high-level then steep decline of human rights violations, the
+extent to which the military government targeted the opposition, and the
+spatial variation of state repression. We also discuss our coding
+procedures, show descriptive statistics, and conclude with suggestions
+for potential uses of our dataset.
+\vspace{.5cm}
+
+\noindent \textbf{Keywords}: authoritarian regimes, Chile, georeferenced event data, Pinochet, truth
+commission
+\vspace{.25cm}
+
+\end{abstract}
+\newpage
+
+\hypertarget{introduction}{%
+\section{\texorpdfstring{Introduction\label{sec:intro}}{Introduction}}\label{introduction}}
 
@@
 
-On 11 September 1973, General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende.
-The coup marked the beginning of a seventeen-year military dictatorship which combined rapid economic liberalisation with large-scale political repression [@meadowcroft2014hayek; @valdes1995pinochet].
-In 1991, after Chile's transition to democracy, then-President Patricio Aylwin convened a commission to investigate human rights abuses in the Pinochet regime.
-The Report of the National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], recorded over 2,000 cases of murders and disappearances from 1973 to 1990.
-The Commission confronted the abuses of the past and portrayed a careful reconstruction of state repression during the military government.
-Most important, the Report contributed to the erosion of authoritarianism in Chile, as it fostered a series of criminal investigations that culminated with the indictment of General Pinochet in 2000 [@brahm2007uncovering 26].
+On 11 September 1973, General Augusto Pinochet led a coup against
+Chile's socialist President Salvador Allende. The coup marked the
+beginning of a seventeen-year military dictatorship which combined rapid
+economic liberalisation with large-scale political repression
+\citep{meadowcroft2014hayek, valdes1995pinochet}. In 1991, after Chile's
+transition to democracy, then-President Patricio Aylwin convened a
+commission to investigate human rights abuses in the Pinochet regime.
+The Report of the National Commission on Truth and Reconciliation
+\citeyearpar{report1991}, also known as the Rettig Report\footnote{Former
+  Chilean ambassador Raúl Rettig chaired the Chilean National Commission
+  for Truth and Reconciliation.}, recorded over 2,000 cases of murders
+and disappearances from 1973 to 1990. The Commission confronted the
+abuses of the past and portrayed a careful reconstruction of state
+repression during the military government. Most important, the Report
+contributed to the erosion of authoritarianism in Chile, as it fostered
+a series of criminal investigations that culminated with the indictment
+of General Pinochet in 2000 \citep[ 26]{brahm2007uncovering}.
 
-Although the Report is a valuable resource to scholars, quantitative researchers cannot readily use the rich information it contains.
-In this paper, we present a manually-coded dataset with all information from the Truth Commission Report plus new variables we collected to complement the original data.
-We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each victim, then georeferenced the location of the human rights abuses.
-We mapped all locations that appear in the report, such as places used to torture, interrogate or murder the victims.
+Although the Report is a valuable resource to scholars, quantitative
+researchers cannot readily use the rich information it contains. In this
+paper, we present a manually-coded dataset with all information from the
+Truth Commission Report plus new variables we collected to complement
+the original data. We transcribed every personal detail from the 903
+pages of the English translation of the Report, assigned a unique
+identification number to each victim, then georeferenced the location of
+the human rights abuses. We mapped all locations that appear in the
+report, such as places used to torture, interrogate or murder the
+victims.
 
-Apart from the geographical location of the incidents, our dataset includes variables on: 1) the sociological characteristics of the victim; 2) their affiliation (where known); 3) the type of violence that took place during that event; 4) whether the victim was interrogated, tortured or in some other way mistreated (if known); 5) who were the perpetrators of the violence.
-If the Report does not have a particular information, we coded it as missing.
-As each individual receives their own id number, new information can be added to the dataset as archival work progresses.
-In the following sections, we give some background about the Pinochet regime, describe our dataset, show summary statistics for selected variables, and suggest how our data can help answer future research questions.
+Apart from the geographical location of the incidents, our dataset
+includes variables on: 1) the sociological characteristics of the
+victim; 2) their affiliation (where known); 3) the type of violence that
+took place during that event; 4) whether the victim was interrogated,
+tortured or in some other way mistreated (if known); 5) who were the
+perpetrators of the violence. If the Report does not have a particular
+information, we coded it as missing. As each individual receives their
+own id number, new information can be added to the dataset as archival
+work progresses. In the following sections, we give some background
+about the Pinochet regime, describe our dataset, show summary statistics
+for selected variables, and suggest how our data can help answer future
+research questions.
 
@@
 
-# Historical Background
+\hypertarget{historical-background}{%
+\section{Historical Background}\label{historical-background}}
 
-Violence during transitions has characterised regime change in Latin America.
-Transitions to autocratic rule in Brazil (1964) and Argentina (1976) prompted frequent human rights violations that only recently have been fully documented [@agosin1987generals; @elpais2016argentina].
-In some cases, autocrats declared unrestrained violence to be an explicit government policy.
-Peru's former President Alberto Fujimori openly stated that his _autogolpe_ (1992) granted him the rights to "[apply] drastic punishments" towards alleged terrorists [@dabene1997narcodemocracias, 97; @samtleben2013constitucion, 56].
+Violence during transitions has characterised regime change in Latin
+America. Transitions to autocratic rule in Brazil (1964) and Argentina
+(1976) prompted frequent human rights violations that only recently have
+been fully documented \citep{agosin1987generals, elpais2016argentina}.
+In some cases, autocrats declared unrestrained violence to be an
+explicit government policy. Peru's former President Alberto Fujimori
+openly stated that his \emph{autogolpe} (1992) granted him the rights to
+``{[}apply{]} drastic punishments'' towards alleged terrorists
+\citetext{\citealp[97]{dabene1997narcodemocracias}; \citealp[56]{samtleben2013constitucion}}.
 
-The Pinochet regime is also notable for its brutality.
-On the day of the coup, the military opened fire against the _Palacio La Moneda_, the Presidential Office, and President Salvador Allende committed suicide after troops stormed the building [@verdugo2001chile].
-The violence continued during the military government.
-The _junta_ killed thousands of students and members of labour organisations in 1973, with up to 200,000 Chileans -- about 2% of the country's population -- led into exile [@wright2007chilean, 31].
+The Pinochet regime is also notable for its brutality. On the day of the
+coup, the military opened fire against the \emph{Palacio La Moneda}, the
+Presidential Office, and President Salvador Allende committed suicide
+after troops stormed the building \citep{verdugo2001chile}. The violence
+continued during the military government. The \emph{junta} killed
+thousands of students and members of labour organisations in 1973, with
+up to 200,000 Chileans -- about 2\% of the country's population -- led
+into exile \citep[31]{wright2007chilean}.
 
-The establishment of the Truth and Reconciliation Commission was the first attempt of
+The establishment of the Truth and Reconciliation Commission was the
+first attempt of
 
-# The Dataset
+\hypertarget{the-dataset}{%
+\section{The Dataset}\label{the-dataset}}
 
-Our dataset comprises 2,398 observations and 57 variables.
-Each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number.
-There are several variables describing the personal information of the victims, such as age, gender, nationality, occupation, and political affiliation if available.
-The data also feature geographical coordinates for a number of the incidents, as well as information about methods of torture and causes of death.
-Figure 01 shows an example entry from our dataset.
+Our dataset comprises 2,398 observations and 57 variables. Each
+observation corresponds to a victim of the Pinochet regime and every
+individual has a unique id number. There are several variables
+describing the personal information of the victims, such as age, gender,
+nationality, occupation, and political affiliation if available. The
+data also feature geographical coordinates for a number of the
+incidents, as well as information about methods of torture and causes of
+death. Figure 01 shows an example entry from our dataset.
 
-[comment]: <> (Example Entry)
+Users can download the data as an Excel spreadsheet (\texttt{.xlsx}) or
+as a comma-separated values text file (\texttt{.csv}). We have also
+created an \texttt{R} package called \texttt{pinochet} which includes
+the dataset in both formats plus the data codebook. The files follow the
+principles of ``tidy data'', where each column represents one variable
+and each row is one case \citep{wickham2014tidy}. The \texttt{pinochet}
+package is available for download on the Comprehensive R Archive Network
+(CRAN) and at \url{http://github.com/danilofreire/pinochet}. To make the
+data accessible, our GitHub repository has detailed installation
+instructions for users new to \texttt{R}.
 
-Users can download the data as an Excel spreadsheet (`.xlsx`) or as a comma-separated values text file (`.csv`).
-We have also created an `R` package called `pinochet` which includes the dataset in both formats plus the data codebook.
-The files follow the principles of "tidy data", where each column represents one variable and each row is one case [@wickham2014tidy].
-The `pinochet` package is available for download on the Comprehensive R Archive Network (CRAN) and at <http://github.com/danilofreire/pinochet>.
-To make the data accessible, our GitHub repository has detailed installation instructions for users new to `R`.
+\hypertarget{types-of-violence}{%
+\subsection{Types of Violence}\label{types-of-violence}}
 
-## Types of Violence
+The Report distinguishes between different types of violence carried out
+by the Pinochet regime. The first type is \emph{deaths}. These are cases
+where the Commission signals a definite and known death of the victim.
+Take the case of Benito Heriberto Torres Torres, one of the first
+victims of the Pinochet regime (id number 2). Our dataset shows that Mr
+Torres was male, 57 years-old, and that he worked as a plumber. On the
+12 September 1973, just one day after the military coup, Mr Torres was
+torture on the way to the 26th police station in Santiago. The records
+indicate he was executed and that his body was later found in Las
+Barrancas, Santiago. The dataset also shows that we obtained this
+information from pages 159-160 of the Truth Commission Report.
 
-The Report distinguishes between different types of violence carried out by the Pinochet regime.
-The first type is _deaths_.
-These are cases where the Commission signals a definite and known death of the victim.
-Take the case of Benito Heriberto Torres Torres, one of the first victims of the Pinochet regime (id number 2).
-Our dataset shows that Mr Torres was male, 57 years-old, and that he worked as a plumber.
-On the 12 September 1973, just one day after the military coup, Mr Torres was torture on the way to the 26th police station in Santiago.
-The records indicate he was executed and that his body was later found in Las Barrancas, Santiago.
-The dataset also shows that we obtained this information from pages 159-160 of the Truth Commission Report.
+The second type of violence recorded in the dataset is
+\emph{disappearances}. These are cases where the Truth Commission
+presumes government agents to have killed and disposed of the body of
+the victims. One such example is that of the Brazilian engineer Tulio
+Roberto Quintiliano Cardozo (id number 5). Mr Cardozo was a member of
+the Communist party and troops took him to the Military Academy for
+interrogation also on the 12 September 1973. He was never seen again and
+is presumed dead.
 
-The second type of violence recorded in the dataset is _disappearances_.
-These are cases where the Truth Commission presumes government agents to have killed and disposed of the body of the victims.
-One such example is that of the Brazilian engineer Tulio Roberto Quintiliano Cardozo (id number 5).
-Mr Cardozo was a member of the Communist party and troops took him to the Military Academy for interrogation also on the 12 September 1973.
-He was never seen again and is presumed dead.
+The third category is \emph{disappearance, information of death}. As the
+name implies, these observations refer to cases the Commission confirms
+the individual died after being missing. The formal definition for this
+category is the following: ``the victims are dead; they died at the
+hands of the government agents, or persons in their service; and these
+or other agents disposed of the victims' mortal remains by throwing them
+into a river or a sea, by covertly burying them, or by disposing of them
+in some other secret fashion'' \citep[ 44]{report1991}. The
+assassination of Humberto de las Nieves Fuentes Rodriguez is an example
+(id number 854). The military took in custody to the Colina air base,
+then loaded onto a helicopter with other political prisoners. According
+to the Report, government agents drugged him, beat him with a metal bar,
+opened his stomach with a knife before throwing the former Communist
+alderman from a helicopter. While his body has not been found, there is
+enough information about the case to classify his death as a severe
+human rights violation.
 
-The third category is _disappearance, information of death_.
-As the name implies, these observations refer to cases the Commission confirms the individual died after being missing.
-The formal definition for this category is the following: "the victims are dead; they died at the hands of the government agents, or persons in their service; and these or other agents disposed of the victims' mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion" [@report1991 44].
-The assassination of Humberto de las Nieves Fuentes Rodriguez is an example (id number 854).
-The military took in custody to the Colina air base, then loaded onto a helicopter with other political prisoners.
-According to the Report, government agents drugged him, beat him with a metal bar, opened his stomach with a knife before throwing the former Communist alderman from a helicopter.
-While his body has not been found, there is enough information about the case to classify his death as a severe human rights violation.
+The last group is that of \emph{unresolved} cases. The Report defines
+them as cases where insufficient information or evidence is available.
+Our dataset counts 93 unresolved cases, many with georeferenced
+information about possible torture sites\footnote{Two cases are
+  ambiguously described in the Truth and Commission Report, so we treat
+  them as missing data. The victims are Ruiter Enrique Correa Arce, a
+  news stand owner accused of facilitating message exchanges between
+  party leaders (id number 843), and Alonso Fernando Gahona Chavez, a
+  communist leader of municipal workers of La Cisterna (id number 847).}.
+However, it is very likely that the people who disappeared were, in
+fact, killed. Based on the report and our methodology, we can only
+determine that 65\% of all of the cases led to an assassination. Forty
+percent of the documented cases are disappearances.
 
-The last group is that of _unresolved_ cases.
-The Report defines them as cases where insufficient information or evidence is available.
-Our dataset counts 93 unresolved cases, many with georeferenced information about possible torture sites[^ambiguous].
-However, it is very likely that the people who disappeared were, in fact, killed.
-Based on the report and our methodology, we can only determine that 65% of all of the cases led to an assassination.
-Forty percent of the documented cases are disappearances.
+Although the Report provides a comprehensive list of the regime victims,
+the Chilean Truth Commission itself acknowledges that the data are
+incomplete. The Commission had only nine months to write the Report and
+it had no legal power to oblige citizens to cooperate. Therefore, we
+expect that some cases have not been recorded. But this problem is not
+unique to the Pinochet regime. Selection bias is a common feature in
+events data, even more so in datasets about armed conflict or government
+repression \citep{ball2019using, chapman2001truth}. As the cases might
+not be a representative sample of the population of victims, we suggest
+scholars to consider issues of model specification.
 
-[comment]: <> (Table 1 around here)
+\hypertarget{geocoding}{%
+\subsection{Geocoding}\label{geocoding}}
 
-Although the Report provides a comprehensive list of the regime victims, the Chilean Truth Commission itself acknowledges that the data are incomplete.
-The Commission had only nine months to write the Report and it had no legal power to oblige citizens to cooperate.
-Therefore, we expect that some cases have not been recorded.
-But this problem is not unique to the Pinochet regime.
-Selection bias is a common feature in events data, even more so in datasets about armed conflict or government repression [@ball2019using; @chapman2001truth].
-As the cases might not be a representative sample of the population of victims, we suggest scholars to consider issues of model specification.
+We have georeferenced all observations for which the relevant data was
+available. Based on data from the Truth Commission, we first created a
+variable indicating where the victim was allegedly tortured or murder.
+We assigned a number to the six locations that most often appear in the
+report: 1) home; 2) work; 3) university; 4) in custody; 5) in public;
+and 6) in hospital. We assigned ``unknown'' if we could not identify the
+location. Then we added latitude and longitude coordinates to each
+place.
 
-## Geocoding
+Since the Truth Commission Report includes spatial data in string format
+-- e.g., intersection of \emph{Calle Grecia} with \emph{Avenida Rosa} --
+we mapped the locations using decimal degrees to increase precision. The
+mapping process is straightforward. We used three different websites to
+increase the coverage of our maps and include the largest possible
+number of observations. The first website we employed is Google Maps
+(\url{http://maps.google.co.uk}), the world's most popular map software.
+We queried Google Maps' application programming interface (API) to
+search for street intersections and other precise addressed we retrieved
+from the Truth Commission Report. Additionally, we used the API to
+locate police stations and government buildings still in use.
 
-We have georeferenced all observations for which the relevant data was available.
-Based on data from the Truth Commission, we first created a variable indicating where the victim was allegedly tortured or murder.
-We assigned a number to the six locations that most often appear in the report: 1) home; 2) work; 3) university; 4) in custody; 5) in public; and 6) in hospital.
-We assigned "unknown" if we could not identify the location.
-Then we added latitude and longitude coordinates to each place.
+Next, we used Wikimapia (\url{https://wikimapia.org/}), a open-content
+collaborative platform that allows users to mark locations worldwide. As
+of March 2019, Wikimapia contains about 30 million places in its
+database\footnote{Wikimapia statistics for March 2019 are available at
+  \url{https://wikimapia.org/stats/action_stats/?fstat=2\&period=3\&year=2019\&month=3}.}.
+We benefited from this crowdsourced information to identify
+shanty-towns, military bases, and small neighbourhoods that we could not
+find in general-purpose maps.
 
-Since the Truth Commission Report includes spatial data in string format -- e.g., intersection of _Calle Grecia_ with _Avenida Rosa_ -- we mapped the locations using decimal degrees to increase precision.
-The mapping process is straightforward.
-We used three different websites to increase the coverage of our maps and include the largest possible number of observations.
-The first website we employed is Google Maps (<http://maps.google.co.uk>), the world's most popular map software.
-We queried Google Maps' application programming interface (API) to search for street intersections and other precise addressed we retrieved from the Truth Commission Report.
-Additionally, we used the API to locate police stations and government buildings still in use.
+Our third source of geographic coordinates is Latitude
+(\url{http://latitude.to}). We used Latitude to source inexact
+coordinates, such as coordinates for cities and towns. If no precise
+information was available, our measure of latitude and longitude
+correspond to those of the control point of each city. We also used the
+Latitude.to website to convert non-decimal coordinates to decimal ones.
+For instance, some addresses in Wikimapia are in degrees, minutes, and
+seconds (DMS) format, and to keep the dataset comparable all coordinates
+follow the same standard.
 
-Next, we used Wikimapia (<https://wikimapia.org/>), a open-content collaborative platform that allows users to mark locations worldwide.
-As of March 2019, Wikimapia contains about 30 million places in its database[^wikimapia].
-We benefited from this crowdsourced information to identify shanty-towns, military bases, and small neighbourhoods that we could not find in general-purpose maps.
+\hypertarget{variables-and-patterns}{%
+\subsection{Variables and Patterns}\label{variables-and-patterns}}
 
-Our third source of geographic coordinates is Latitude (<http://latitude.to>).
-We used Latitude to source inexact coordinates, such as coordinates for cities and towns.
-If no precise information was available, our measure of latitude and longitude correspond to those of the control point of each city.
-We also used the Latitude.to website to convert non-decimal coordinates to decimal ones.
-For instance, some addresses in Wikimapia are in degrees, minutes, and seconds (DMS) format, and to keep the dataset comparable all coordinates follow the same standard.
+As mentioned previously, our dataset includes 57 variables and covers
+the seventeen years of the Pinochet Presidency (1973--1990). Figure
+\ref{trend} shows the frequency of government-perpetrated incidents per
+month from 1973 to 1990. We see that the majority of the acts of
+violence happened in the first two months after the military coup. The
+high incidence of human rights violations in the beginning of the
+Pinochet era reflects the activity of the ``Caravan of Death'', a group
+of paramilitaries established in October 1973 to torture and kill
+dissidents
+\citetext{\citealp[459]{davis1990they}; \citealp{verdugo2001chile}}.
 
-## Variables and Patterns
+In September 1973 alone, the government committed 633 acts of violence
+against civilians, of which 406 the Commission signals as ``definite and
+known deaths''. The violations continued in October, with 504 incidences
+and 317 certified deaths. The numbers decline to about 10\% of those
+levels in November and December, and then recede further in 1974.
 
-As mentioned previously, our dataset includes 57 variables and covers the seventeen years of the Pinochet Presidency (1973--1990).
-Figure \ref{trend} shows the frequency of government-perpetrated incidents per month from 1973 to 1990.
-We see that the majority of the acts of violence happened in the first two months after the military coup.
-The high incidence of human rights violations in the beginning of the Pinochet era reflects the activity of the "Caravan of Death", a group of paramilitaries established in October 1973 to torture and kill dissidents [@davis1990they, 459; @verdugo2001chile].
-
-[comment]: <> (Figure 02 - Time Trends Here)
-
-In September 1973 alone, the government committed 633 acts of violence against civilians, of which 406 the Commission signals as "definite and known deaths".
-The violations continued in October, with 504 incidences and 317 certified deaths.
-The numbers decline to about 10% of those levels in November and December, and then recede further in 1974.
-
-In absolute numbers, the Chilean dictatorship was not as violent as its Argentinian counterpart[^argentina], but the figures are high when compared to other South American countries.
-For example, the Brazilian military regime made 434 victims from 1964 to 1985; Paraguay had 424 victims of human rights abuse from 1954 to 1989; and Uruguay reported about 300 incidences from 1973 to 1985 [@elpais2014ditadura; @fsp2016argentina].
-The timing of repression also varies.
-Despite eventual periods of higher intensity, violence in authoritarian Brazil and Paraguay stretched somewhat evenly during the regime [@agenciabrasil2014ditadura; @elpais2014ditadura].
-This contrasts with the Chilean case, whose repression was concentrated in the early years of the regime.
+In absolute numbers, the Chilean dictatorship was not as violent as its
+Argentinian counterpart\footnote{The Argentinian government disputes the
+  figure of 30,000 deaths and disappearances human rights organisations
+  often suggest \citep{bbc2016argentina, elpais2016argentina}. Ceferino
+  Reato, a writer and journalist, indicates that there were 7,158
+  incidences during the regime, including 6,415 cases of disappearances
+  and 743 victims of summary executions \citep{clarin2016argentina}.
+  However, General Jorge Rafael Videla, who ruled Argentina from 1976 to
+  1981, seems to agree with those figures. He once remarked that killing
+  ``about 7 or 8 thousand people'' was ``the price of winning the war
+  against the subversives'' \citep{clarin2012videla}.}, but the figures
+are high when compared to other South American countries. For example,
+the Brazilian military regime made 434 victims from 1964 to 1985;
+Paraguay had 424 victims of human rights abuse from 1954 to 1989; and
+Uruguay reported about 300 incidences from 1973 to 1985
+\citep{elpais2014ditadura, fsp2016argentina}. The timing of repression
+also varies. Despite eventual periods of higher intensity, violence in
+authoritarian Brazil and Paraguay stretched somewhat evenly during the
+regime \citep{agenciabrasil2014ditadura, elpais2014ditadura}. This
+contrasts with the Chilean case, whose repression was concentrated in
+the early years of the regime.
 
 We can also explore how the government employed violence across space.
-We see that assassinations are concentrated around Santiago and neighbouring cities.
-The results are expected.
-Santiago has been Chile's most densely populated city since the colonial times and in 1975 the city was home to about 30% of the country's population [@un1976yearbook].
-Disappearances follow a similar pattern.
-Although there are cases in the northern parts of the country, the military government targeted more victims in Santiago and neighbouring areas.
+We see that assassinations are concentrated around Santiago and
+neighbouring cities. The results are expected. Santiago has been Chile's
+most densely populated city since the colonial times and in 1975 the
+city was home to about 30\% of the country's population
+\citep{un1976yearbook}. Disappearances follow a similar pattern.
+Although there are cases in the northern parts of the country, the
+military government targeted more victims in Santiago and neighbouring
+areas.
 
-[comment]: <> (Figure 03 - Spatial variation)
+We can further disaggregate the analysis and visualise how the regime
+persecuted particular groups. Students and members of left-wing parties
+are two categories of special interest. On 18 September 1073, only one
+week after the coup, Pinochet and his followers rounded up thousands of
+students and activists in Chile's National Stadium. Military forces
+tortured and murdered hundreds of prisoners over the course of a few
+days. An emblematic victim was Víctor Lidio Jara Martínez, a university
+teacher, theatre director and singer (id number 25). According to
+witnesses, Mr Jara had his hand crushed and his wrists broken, and he
+was later shot 44 times with a machine gun. Mr Jara's body was left in
+the front of the stadium as a warning for other political prisoners
+\citep[ 56]{mcsherry2015victor}. The government again tortured and
+killed students and activists in its campaign against the Revolutionary
+Left Movement (\emph{Movimiento de Izquierda Revolucionaria}, MIR). The
+MIR originated in student unions and in the late 1970s the movement
+promoted a series of attacks against government personnel and official
+buildings. Government forces strongly retaliated and the group resorted
+to terrorist tactics to confront the regime
+\citep{schlotterbeck2018beyond}.
 
-We can further disaggregate the analysis and visualise how the regime persecuted particular groups.
-Students and members of left-wing parties are two categories of special interest.
-On 18 September 1073, only one week after the coup, Pinochet and his followers rounded up thousands of students and activists in Chile's National Stadium.
-Military forces tortured and murdered hundreds of prisoners over the course of a few days.
-An emblematic victim was Víctor Lidio Jara Martínez, a university teacher, theatre director and singer (id number 25).
-According to witnesses, Mr Jara had his hand crushed and his wrists broken, and he was later shot 44 times with a machine gun.
-Mr Jara's body was left in the front of the stadium as a warning for other political prisoners [@mcsherry2015victor 56].
-The government again tortured and killed students and activists in its campaign against the Revolutionary Left Movement (_Movimiento de Izquierda Revolucionaria_, MIR).
-The MIR originated in student unions and in the late 1970s the movement promoted a series of attacks against government personnel and official buildings.
-Government forces strongly retaliated and the group resorted to terrorist tactics to confront the regime [@schlotterbeck2018beyond].
+\hypertarget{conclusion-new-avenues-for-research}{%
+\section{Conclusion: New Avenues for
+Research}\label{conclusion-new-avenues-for-research}}
 
-[comment]: <> (Figure 04 - Students and Communist Leaders)
+In this paper, we introduce a dataset with rich information about more
+than 2,000 victims of the Pinochet regime. Our data come from two
+sources. First, we manually coded all information available in the
+Report of the Chilean National Commission on Truth and Reconciliation
+\citeyearpar{report1991}. Second, we added the geographical locations
+and the specific dates of the human rights abuses whenever we could
+retrieve them. The graphs and maps included in this article provide some
+preliminary results about the temporal and spatial variation of state
+violence during Chile's last military government.
 
-# Conclusion: New Avenues for Research
+We believe our data open new topics of research. For instance,
+\citet{lupu2017legacy}, \citet{rozenas2017political} and
+\citet{zhukov2018stalin} highlight that state repression has enduring
+effects on political preferences and social attitudes. Researchers can
+test whether the Pinochet regime has caused similar attitudinal changes
+in direct or indirect victims. Moreover, sociologists and criminologists
+can analyse the relationship between human rights abuses and post-regime
+levels of interpersonal violence. Recent studies show that democracies
+which arise after military regimes have higher homicide rates
+\citep{frantz2018legacy, karstedt2006democracy}. Our data can show if
+areas with significant levels of military repression are more violent
+today.
 
-In this paper, we introduce a dataset with rich information about more than 2,000 victims of the Pinochet regime.
-Our data come from two sources.
-First, we manually coded all information available in the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991].
-Second, we added the geographical locations and the specific dates of the human rights abuses whenever we could retrieve them.
-The graphs and maps included in this article provide some preliminary results about the temporal and spatial variation of state violence during Chile's last military government.
+Researchers can also examine how political coalitions affect the use of
+lethal violence in authoritarian regimes. Although the topic has
+received increasing attention
+\citep[e.g.,][]{fjelde2010generals, gandhi2007authoritarian, rivera2017authoritarian},
+the internal dynamics of autocratic governments remains understudied.
+The main reason is a lack of fine-grained information
+\citep[16]{ferrara2014assessing}. By linking human right abuses to
+changes in Pinochet's coalition, scholars can explore whether civilian
+or bureaucratic support lead to higher incidence of state violence. The
+individual data presented here can be combined with government records
+at any level of aggregation.
 
-We believe our data open new topics of research.
-For instance, @lupu2017legacy, @rozenas2017political and @zhukov2018stalin highlight that state repression has enduring effects on political preferences and social attitudes.
-Researchers can test whether the Pinochet regime has caused similar attitudinal changes in direct or indirect victims.
-Moreover, sociologists and criminologists can analyse the relationship between human rights abuses and post-regime levels of interpersonal violence.
-Recent studies show that democracies which arise after military regimes have higher homicide rates [@frantz2018legacy; @karstedt2006democracy].
-Our data can show if areas with significant levels of military repression are more violent today.
+Qualitative scholars will find the personal details of the victims to be
+particularly useful. Historians willing to reconstruct the biographies
+of specific individuals are able to access pre-compiled information in a
+single digital file. Others might be interested in using our data as a
+starting point for network analysis or to collect oral testimony from
+survivors and acquaintances. In that regard, the dataset can accommodate
+future qualitative information. As we include a unique identification
+number to each victim, it is easy to update the personal record of any
+individual with new data from public archives or personal
+correspondence.
 
-Researchers can also examine how political coalitions affect the use of lethal violence in authoritarian regimes.
-Although the topic has received increasing attention [e.g., @fjelde2010generals; @gandhi2007authoritarian; @rivera2017authoritarian], the internal dynamics of autocratic governments remains understudied.
-The main reason is a lack of fine-grained information [@ferrara2014assessing, 16].
-By linking human right abuses to changes in Pinochet's coalition, scholars can explore whether civilian or bureaucratic support lead to higher incidence of state violence.
-The individual data presented here can be combined with government records at any level of aggregation.
-
-Qualitative scholars will find the personal details of the victims to be particularly useful.
-Historians willing to reconstruct the biographies of specific individuals are able to access pre-compiled information in a single digital file.
-Others might be interested in using our data as a starting point for network analysis or to collect oral testimony from survivors and acquaintances.
-In that regard, the dataset can accommodate future qualitative information.
-As we include a unique identification number to each victim, it is easy to update the personal record of any individual with new data from public archives or personal correspondence.
-
-Lastly, scholars can investigate the connections between international legitimacy and domestic politics in repressive regimes.
-This is a promising area of research as the Chilean government and American intelligence services continue to declassify documents from the Pinochet era.
-One relevant question is whether pressure from foreign governments and organisations had any influence over the levels of human rights abuses in Chile.
-We hope our dataset is useful for scholars interested in these and other questions, and that the information it contains elicits hypotheses not only about the Pinochet period, but about authoritarian governments more generally.
-
-[^df]: Postdoctoral Research Fellow, the Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), [`http://danilofreire.github.io`](http://danilofreire.github.io).
-Corresponding author.
-
-[^jm]: Senior Lecturer in Public Policy, Department of Political Economy, King's College London, [`john.meadowcroft@kcl.ac.uk`](mailto:john.meadowcroft@kcl.ac.uk), [`http://johnmeadowcroft.net`](http://johnmeadowcroft.net).
-
-[^ds]: Associate Professor, Department of Political Science and the Political Theory Project, Brown University, [`david_skarbek@brown.edu`](david_skarbek@brown.edu), [`http://davidskarbek.com`](http://davidskarbek.com).
-
-[^eg]: Software Developer, Tempo UK.
-
-[^rettig]: Former Chilean ambassador Raúl Rettig chaired the Chilean National Commission for Truth and Reconciliation.
-
-[^omissions]: The Truth and Commission Report investigated only cases of death and disappearances, leaving aside a high number of other serious human rights violations committed by the Pinochet government.
-For instance, the Report makes no mention of cases of torture or forced exile, which comprise a significant share of the reported violations [@ensalaco1994truth, 660].
-
-[^ambiguous]: Two cases are ambiguously described in the Truth and Commission Report, so we treat them as missing data.
-The victims are Ruiter Enrique Correa Arce, a news stand owner accused of facilitating message exchanges between party leaders (id number 843), and Alonso Fernando Gahona Chavez, a communist leader of municipal workers of La Cisterna (id number 847).
-
-[^wikimapia]: Wikimapia statistics for March 2019 are available at <https://wikimapia.org/stats/action_stats/?fstat=2&period=3&year=2019&month=3>.
-
-[^argentina]: The Argentinian government disputes the figure of 30,000 deaths and disappearances human rights organisations often suggest [@bbc2016argentina; @elpais2016argentina].
-Ceferino Reato, a writer and journalist, indicates that there were 7,158 incidences during the regime, including 6,415 cases of disappearances and 743 victims of summary executions [@clarin2016argentina].
-However, General Jorge Rafael Videla, who ruled Argentina from 1976 to 1981, seems to agree with those figures.
-He once remarked that killing "about 7 or 8 thousand people" was "the price of winning the war against the subversives" [@clarin2012videla].
+Lastly, scholars can investigate the connections between international
+legitimacy and domestic politics in repressive regimes. This is a
+promising area of research as the Chilean government and American
+intelligence services continue to declassify documents from the Pinochet
+era. One relevant question is whether pressure from foreign governments
+and organisations had any influence over the levels of human rights
+abuses in Chile. We hope our dataset is useful for scholars interested
+in these and other questions, and that the information it contains
+elicits hypotheses not only about the Pinochet period, but about
+authoritarian governments more generally.
 
@@
 \setlength{\parskip}{6pt}
+
+\bibliography{references.bib}
+
+\end{document}

### 2019-03-31 rap/article.Rmd
written draft v0.1

@@
 date: \today
-thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments.
-Data and replication information are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
-abstract: "This article presents a georeferenced dataset on human rights violations in General Augusto Pinochet's dictatorship in Chile.
-We manually coded the personal details of 2,398 victims mentioned in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations.
-Our dataset comprises 57 variables from 1973 to 1990 and it is available as a stand-alone spreadsheet or as the `pinochet` package for `R`.
-The data describe the initially high-level then steep decline of human rights violations, the extent to which the military government targeted the opposition, and the spatial variation of state repression.
-We also discuss our coding procedures, show descriptive statistics, and conclude with suggestions for potential uses of our dataset."
-keywords: "authoritarian regimes, Chile, georeferenced event data, Pinochet, truth commission"
+thanks: "We thank Fábio Barros, Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments.
+Replication materials and R source code are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
+abstract: "This article presents a georeferenced dataset on human rights violations in the Pinochet dictatorship in Chile.
+We coded the personal details of 2,398 victims named in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations.
+The dataset comprises 57 variables from 1973 to 1990 and is available as a stand-alone spreadsheet or as the `pinochet` package for `R`.
+As examples, we describe the major temporal and spatial patterns of the human rights abuses.
+We also discuss our coding procedures, show descriptive statistics, and conclude with suggestions for potential applications of the dataset."
+keywords: "Chile, georeferenced event data, human rights, Pinochet, truth commission"
 fontfamily: libertine
@@
 The coup marked the beginning of a seventeen-year military dictatorship which combined rapid economic liberalisation with large-scale political repression [@meadowcroft2014hayek; @valdes1995pinochet].
-In 1991, after Chile's transition to democracy, then-President Patricio Aylwin convened a commission to investigate human rights abuses in the Pinochet regime.
+In 1991, after Chile's transition to democracy, President Patricio Aylwin convened a commission to investigate human rights violations in the Pinochet regime.
 The Report of the National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], recorded over 2,000 cases of murders and disappearances from 1973 to 1990.
-The Commission confronted the abuses of the past and portrayed a careful reconstruction of state repression during the military government.
-Most important, the Report contributed to the erosion of authoritarianism in Chile, as it fostered a series of criminal investigations that culminated with the indictment of General Pinochet in 2000 [@brahm2007uncovering 26].
+The Commission confronted the abuses of the past and portrayed a careful reconstruction of state violence during the military government.
+Most importantly, the Report contributed to the erosion of authoritarianism in Chile, as it fostered a series of criminal investigations that culminated with the indictment of General Pinochet in 2000 [@brahm2007uncovering 26].
 
-Although the Report is a valuable resource to scholars, quantitative researchers cannot readily use the rich information it contains.
-In this paper, we present a manually-coded dataset with all information from the Truth Commission Report plus new variables we collected to complement the original data.
+Although the Report is a valuable source of information for scholars, quantitative researchers cannot readily use the rich data it contains.
+In this paper, we present a manually-coded dataset with all information from the Truth Commission Report plus new variables we collected to complement the original records.
 We transcribed every personal detail from the 903 pages of the English translation of the Report, assigned a unique identification number to each victim, then georeferenced the location of the human rights abuses.
-We mapped all locations that appear in the report, such as places used to torture, interrogate or murder the victims.
+We added coordinates of latitude and longitude to every location mentioned in the report, such as places used to torture, interrogate or murder the victims.
 
@@
 In some cases, autocrats declared unrestrained violence to be an explicit government policy.
-Peru's former President Alberto Fujimori openly stated that his _autogolpe_ (1992) granted him the rights to "[apply] drastic punishments" towards alleged terrorists [@dabene1997narcodemocracias, 97; @samtleben2013constitucion, 56].
+In 1992, Peru's former President Alberto Fujimori openly stated that his _autogolpe_ granted him the rights to "[apply] drastic punishments" towards any suspect of terrorism [@dabene1997narcodemocracias, 97; @samtleben2013constitucion, 56].
 
-The Pinochet regime is also notable for its brutality.
+The Pinochet regime was also notable for its brutality.
 On the day of the coup, the military opened fire against the _Palacio La Moneda_, the Presidential Office, and President Salvador Allende committed suicide after troops stormed the building [@verdugo2001chile].
@@
 
-The establishment of the Truth and Reconciliation Commission was the first attempt of
+The establishment of the Truth and Reconciliation Commission was the first attempt to make the dictatorship accountable for its crimes.
+The Commission focused on the most serious cases of human rights violations, those which resulted in deaths of disappearances[^omissions].
+One factor that limited the scope of the investigations was that the Commission did not have the legal authority to demand collaboration from the military forces [@popkin1995truth, 85].
+The Chilean Commission negotiated an amnesty to perpetrator in exchange for freedom of inquiry, and it collected testimony from victims in Chile and abroad [@quinn2001dealing, 399].
+The result was a broad yet accurate description of the events.
+Although the military contested the conclusions of the Report, they did not question its factual information [@bakiner2009denial, 54; @brahm2005beyond].
+
+While the Report did not bring legal punishment to violence perpetrators, it addressed important concerns of the victims.
+In February 1992, the Chilean Congress passed a law granting a monthly pension to individuals named in the Report, along with psychological treatment for their families and school subsidies for their children [@vasallo2002truth, 167].
+The Chilean government also created a new office to deal with returning nationals, and President Aylwin released almost all political prisoners in the country [@quinn2001dealing, 400].
+
+Data from the Truth Commission Report have also fostered academic research on the effects of authoritarian violence on social outcomes.
+For instance, Ensalaco [-@ensalaco1994truth; -@ensalaco2000chile] argues that the Report had a positive impact on human rights conditions in Chile, while @davis1990they use the Report to analyse the drivers on political violence in Chile.
+More recently, Esberg [-@esberg2018antecipating; -@esberg2018audience] employs data compiled from the Truth Commission Report to examine how the Pinochet regime applied different methods of repression to appeal to supporters and avoid political backlash.
+
+Despite these noteworthy examples, quantitative studies on the dynamics of authoritarian rule in Chile remains limited.
+We believe a major obstacle for further research in the subject is the lack of easily-available, pre-formatted data.
+Manual text coding is a time-consuming process, and data availability would allow scholars to investigate new aspects of the Pinochet regime.
 
@@
 The data also feature geographical coordinates for a number of the incidents, as well as information about methods of torture and causes of death.
-Figure 01 shows an example entry from our dataset.
+Table 1 shows an example entry from our dataset.
+
+\vspace{.5cm}
+\begin{center}
+Table 1 about here.
+
+Example entry with the victim's characteristics\\ and information about their murder or disappearance.
+\end{center}
+\vspace{.5cm}
 
@@
 
-[comment]: <> (Table 1 around here)
+\vspace{.5cm}
+\begin{center}
+Table 2 about here.
+
+Number of killings, disappearances, unresolved cases, and suicides 
+\end{center}
+\vspace{.5cm}
+
+[comment]: <> (Table 2 around here)
 
@@
 As mentioned previously, our dataset includes 57 variables and covers the seventeen years of the Pinochet Presidency (1973--1990).
-Figure \ref{trend} shows the frequency of government-perpetrated incidents per month from 1973 to 1990.
+Figure 1 shows the frequency of government-perpetrated incidents per month from 1973 to 1990.
 We see that the majority of the acts of violence happened in the first two months after the military coup.
@@
 
+\vspace{.5cm}
+\begin{center}
+Figure 1 about here.
+
+Two time trends: one with the whole period,\\ another discarding the first year to bette visualise variation in later years
+\end{center}
+\vspace{.5cm}
+
 [comment]: <> (Figure 02 - Time Trends Here)
+\newpage
 
@@
 
+\vspace{.5cm}
+\begin{center}
+Figure 2 about here.
+
+Two maps: deaths and disappearances 
+\end{center}
+\vspace{.5cm}
+
 [comment]: <> (Figure 03 - Spatial variation)
@@
 Government forces strongly retaliated and the group resorted to terrorist tactics to confront the regime [@schlotterbeck2018beyond].
+
+\vspace{.5cm}
+\begin{center}
+Figure 3 about here.
+
+Time trends: attacks against students and workers
+\end{center}
+\vspace{.5cm}
 
@@
 
-[^rettig]: Former Chilean ambassador Raúl Rettig chaired the Chilean National Commission for Truth and Reconciliation.
+[^rettig]: Former Chilean Senator and jurist Raúl Rettig chaired the Chilean National Commission for Truth and Reconciliation.
 
-[^omissions]: The Truth and Commission Report investigated only cases of death and disappearances, leaving aside a high number of other serious human rights violations committed by the Pinochet government.
-For instance, the Report makes no mention of cases of torture or forced exile, which comprise a significant share of the reported violations [@ensalaco1994truth, 660].
+[^omissions]: The Commission's mandate did not extend to cases of forced exile or torture that did not lead to death [@ensalaco1994truth, 660].
+President Alwyn and his advisors feared the inclusion of non-lethal violence episodes would make the investigation unmanageable [@vasallo2002truth, 163].
 

### 2019-04-29 rap/article.Rmd
typos
add text

@@
 date: \today
-thanks: "We thank Fábio Barros, Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments.
+thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments.
 Replication materials and R source code are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
@@
 On 11 September 1973, General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende.
-The coup marked the beginning of a seventeen-year military dictatorship which combined rapid economic liberalisation with large-scale political repression [@meadowcroft2014hayek; @valdes1995pinochet].
+The coup marked the beginning of a seventeen-year military dictatorship which combined rapid economic liberalisation with large-scale political repression [@valdes1995pinochet].
 In 1991, after Chile's transition to democracy, President Patricio Aylwin convened a commission to investigate human rights violations in the Pinochet regime.
@@
 Violence during transitions has characterised regime change in Latin America.
-Transitions to autocratic rule in Brazil (1964) and Argentina (1976) prompted frequent human rights violations that only recently have been fully documented [@agosin1987generals; @elpais2016argentina].
+Transitions to autocratic rule in Brazil (1964) and Argentina (1976) prompted frequent human rights violations that only recently have been fully documented [@agenciabrasil2014ditadura; @elpais2016argentina].
 In some cases, autocrats declared unrestrained violence to be an explicit government policy.
@@
 The establishment of the Truth and Reconciliation Commission was the first attempt to make the dictatorship accountable for its crimes.
-The Commission focused on the most serious cases of human rights violations, those which resulted in deaths of disappearances[^omissions].
+The Commission focused on the most serious cases of human rights violations, which are those that resulted in deaths of disappearances[^omissions].
 One factor that limited the scope of the investigations was that the Commission did not have the legal authority to demand collaboration from the military forces [@popkin1995truth, 85].
@@
 
-\vspace{.5cm}
-\begin{center}
-Table 1 about here.
+\newpage
 
-Example entry with the victim's characteristics\\ and information about their murder or disappearance.
-\end{center}
-\vspace{.5cm}
+| Variable                | Sample          |
+| :---------------------- | :-------------- |
+| region                  | Taparacá        |
+| id                      | 276             |
+| page                    | 265             |
+| start date              | 18/01/1974      |
+| end date                | 18/01/1974      |
+| last name               | Marques Agusto  |
+| first name              | Nelson José     |
+| minor                   | 0               |
+| age                     | 31              |
+| male                    | 1               |
+| occupation              | Office worker   |
+| victim affiliation      | Communist party |
+| violence                | 1 (killed)      |
+| method                  | 1 (gun)         |
+| interrogation           | NA (unknown)    |
+| torture                 | NA (unknown)    |
+| mistreatment            | 1 (yes)         |
+| perpetrator             | Military        |
+| place                   | 5 (in public)   |
+| start location          | Iquique         |
+| latitude 1              | -20.23070       |
+| longitude 1             | -70.13567       |
+| place 2                 | 4 (in custody)  |
+| location 2              | Pisagua prison  |
+| latitude 2              | -19.59667       |
+| longitude 2             | -70.21222       |
 
-[comment]: <> (Example Entry)
+Table: Sample event from the dataset (selected variables).
+
+\newpage
 
@@
 The dataset also shows that we obtained this information from pages 159-160 of the Truth Commission Report.
+
+| Violence         | _N_    |
+| :--------------- | -----: |
+| Killed           |   1331 |
+| Disappeared      |    853 |
+| Unresolved cases |     93 |
+| Suicides         |      7 |
+| Unknown          |        |
+
+Table: Number of killings, disappearances, unresolved cases, and suicides
 
@@
 Forty percent of the documented cases are disappearances.
-
-\vspace{.5cm}
-\begin{center}
-Table 2 about here.
-
-Number of killings, disappearances, unresolved cases, and suicides 
-\end{center}
-\vspace{.5cm}
-
-[comment]: <> (Table 2 around here)
 
@@
 
-Two time trends: one with the whole period,\\ another discarding the first year to bette visualise variation in later years
+Two time trends: one with the whole period,\\ another discarding the first year to better visualise variation in later years
 \end{center}
@@
 
-[comment]: <> (Figure 02 - Time Trends Here)
+[comment]: <> (Figure 01 - Time Trends Here)
 \newpage
@@
 
-[comment]: <> (Figure 03 - Spatial variation)
+[comment]: <> (Figure 02 - Spatial variation)
 
@@
 
-[comment]: <> (Figure 04 - Students and Communist Leaders)
+[comment]: <> (Figure 03 - Students and Communist Leaders)
 

### 2019-04-29 data-parsing.R

@@ -1,0 +1,131 @@
+########################
+### Parsing the data ###
+########################
+
+## Install required R packages
+if (!require("plyr")) {
+    install.packages("plyr")
+}
+if (!require("tidyverse")) {
+    install.packages("tidyverse")
+}
+if (!require("devtools")) {
+    install.packages("devtools")
+}
+if (!require("narnia")) {
+    install_github("njtierney/narnia")
+}
+if (!require("rnaturalearth")) {
+    install.packages("rnaturalearth")
+}
+
+## Load raw data
+df <- readxl::read_xlsx("data/raw-data.xlsx")
+
+
+## Data manipulation
+# Convert names to lowercase
+names(df) <- tolower(names(df))
+
+# Rename variables
+df <- plyr::rename(df,
+                   c("merged id" = "merged_id",
+                     "start date" = "start_date",
+                     "end date" = "end_date",
+                     "last name" = "last_name",
+                     "first name" = "first_name",
+                     "sex (c)" = "male",
+                     "occupation (c)" = "occupation",
+                     "occupation" = "occupation_2",                     
+                     "victim affiliation (c)" = "victim_affiliation",
+                     "victim affiliation" = "victim_affiliation2",
+                     "violence (c)" = "violence",
+                     "method (c)" = "method",
+                     "interrogation (c)" = "interrogation",
+                     "torture (c)" = "torture",
+                     "war tribunal" = "war_tribunal",
+                     "previous arrests (#)" = "number_previous_arrests",
+                     "perpetrator affiliation (c)" = "perpetrator_affiliation",
+                     "place-1" = "place_1",
+                     "start location-1" = "start_location_1",
+                     "latitude-1" = "latitude_1",
+                     "longitude-1" = "longitude_1",
+                     "exact coordinates-1" = "exact_coordinates_1",
+                     "place-2" = "place_2",
+                     "location-2" = "location_2",
+                     "latitude-2" = "latitude_2",
+                     "longitude-2" = "longitude_2",
+                     "exact coordinates-2" = "exact_coordinates_2",
+                     "place ©-3" = "place_3",
+                     "end location-3" = "end_location_3",
+                     "latitude-3" = "latitude_3",
+                     "longitude-3" = "longitude_3",
+                     "exact coordinates-3" = "exact_coordinates_3",
+                     "place (c)-4" = "place_4",
+                     "end location-4" = "end_location_4",
[…]

### 2019-05-02 rap/article.Rmd
update

@@
 date: \today
-thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments.
+thanks: "We thank Umberto Mignozzetti, Lucas Mingardi and Robert Myles McDonnell for their helpful comments.
 Replication materials and R source code are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
@@
 We coded the personal details of 2,398 victims named in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations.
-The dataset comprises 57 variables from 1973 to 1990 and is available as a stand-alone spreadsheet or as the `pinochet` package for `R`.
+The dataset comprises XXXXXXXXXX variables from 1973 to 1990 and is available as a stand-alone spreadsheet or as the `pinochet` package for `R`.
 As examples, we describe the major temporal and spatial patterns of the human rights abuses.
@@
 
-Our dataset comprises 2,398 observations and 57 variables.
+Our dataset comprises 2,398 observations and XXXXXXXXX variables.
 Each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number.
@@
 | :---------------------- | :-------------- |
-| region                  | Taparacá        |
 | id                      | 276             |
@@
 | first name              | Nelson José     |
-| minor                   | 0               |
+| minor                   | 0 (no)          |
 | age                     | 31              |
-| male                    | 1               |
-| occupation              | Office worker   |
-| victim affiliation      | Communist party |
-| violence                | 1 (killed)      |
-| method                  | 1 (gun)         |
+| male                    | 1 (yes)         |
+| occupation              | office worker   |
+| victim affiliation      | communist party |
+| violence                | killed          |
+| method                  | gun             |
 | interrogation           | NA (unknown)    |
@@
 | mistreatment            | 1 (yes)         |
-| perpetrator             | Military        |
-| place                   | 5 (in public)   |
+| perpetrator             | military        |
+| place                   | in public       |
 | start location          | Iquique         |
@@
 | longitude 1             | -70.13567       |
-| place 2                 | 4 (in custody)  |
+| place 2                 | in custody      |
 | location 2              | Pisagua prison  |
@@
 | Suicides         |      7 |
-| Unknown          |        |
+| Unknown          |    115 |
 
-Table: Number of killings, disappearances, unresolved cases, and suicides
+Table: Number of killings, disappearances, unresolved cases, and suicides.
 

### 2019-05-02 R/script.R

@@ -16,12 +16,12 @@ if (!require("devtools")) {
 if (!require("narnia")) {
         install_github("njtierney/narnia")
 }
+if (!require("ggplot2")) {
+        install.packages("ggplot2")
+}
 if (!require("rnaturalearth")) {
         install.packages("rnaturalearth")
 }
-
-devtools::install_github("tidyverse/ggplot2")
-library(ggplot2)
 
 ## Load data
 df <- readr::read_csv("https://goo.gl/F1tgk1")
@@ -83,12 +83,16 @@ df <- plyr::rename(df,
 
 # Transform variables
 df <- transform(df,
+                region = as.character(region),
+                id = as.integer(id),
                 merged_id = as.integer(merged_id),
+                page = as.character(page),
                 start_date = as.Date(start_date, "%d/%m/%Y"),
                 end_date = as.Date(end_date, "%d/%m/%Y"),
+                last_name = as.character(last.name),
+                first_name = as.character(first_name),
                 minor = as.factor(minor),
                 age = as.numeric(age),
-                female = female - 1,
                 violence = as.integer(violence),
                 method = as.integer(method),
                 number_previous_arrests = as.integer(number_previous_arrests),

### 2019-05-02 manuscript/online-appendix.Rmd
add

@@
 author: 
-  - "Danilo Freire^[PhD Candidate, Department of Political Economy, King's College London.
-Email:   [danilofreire@gmail.com](mailto:danilofreire@gmail.com).] &nbsp; John Meadowcroft^[Senior Lecturer in Public Policy, Department of Political Economy, King's College London.
-Email:  [john.meadowcroft@kcl.ac.uk](mailto:john.meadowcroft@kcl.ac.uk).] &nbsp; David Skarbek^[Associate Professor, Department of Political Science, Brown University.
-Email: [david_skarbek@brown.edu](david_skarbek@brown.edu).
-Corresponding author.] &nbsp; Eugenia Guerrero^[Software Developer, Tempo UK.]"
-date: "`r format(Sys.time(), '%B %d, %Y')`"
-abstract: How did the Pinochet regime use deaths and disappearances after the violent military coup in September 1973?
-The coup resulted in more than three decades of violence as General Pinochet's government fought against left-wing movements for the sake of the 'national wellbeing.' While these efforts were significant, they were not uniformly distributed spatially or over time during his regime.
-This paper provides new data on observed variation of regime-based violence.
-It codes and compiles data from two volumes of the Report of the Chilean National Commission on Truth and Reconciliation, a systematic investigation of the Pinochet regime published in 1991.
-The data documents the initially high-level and then steep decline of these atrocities, the extent to which victims were targeted, and other demographic information about the victims.
-Our data also includes geographic information system data that allows researchers to map the spatial variation of atrocities over time.
-keywords: authoritarianism, Chile, human rights, Pinochet, truth commission
+  - "Danilo Freire[^df]"
+  - "John Meadowcroft[^jm]"
+  - "David Skarbek[^ds]"
+  - "Eugenia Guerrero[^eg]"
+date: \today
+thanks: "Replication materials and R source code are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
+fontfamily: libertine
+fontawesome: yes
 fontsize: 12pt
-monospace-url: no
+monospace-url: yes
 spacing: double
@@
     fig_caption: yes
-    number_sections: no
+    number_sections: yes
     keep_tex: no
     template: template.latex
-    latex_engine: pdflatex
 ---
@@
 
-This document contains the codebook, data and code necessary to replicate the results in "Deaths and Disappearances in the Pinochet Regime" by @freire2017pinochet-regime-data.
-The `dataset.csv` file includes the complete dataset.
+This document contains the codebook, data and code necessary to replicate the results in "Deaths and Disappearances in the Pinochet Regime: A New Dataset" by @freire2019pinochet.
+The `pinochet.csv` file includes the complete dataset.
 The original Excel file, also included in the paper [GitHub repository](https://github.com/danilofreire/pinochet-victims-data), has been pre-processed to reduce compatibility problems with other statistical software such as R or Stata.
-All data pre-processing steps are described below.
-The analyses were performed in the R statistical language, version 3.4.1 [@rstats2017].
 
@@
 
-By compiling and analyzing the data from both volumes of the Report of the Chilean National Commission on Truth and Reconciliation, we seek to understand the logic and dynamics of the violence, deaths, and disappearances under Pinochet’s Chile.
+By compiling and analysing the data from both volumes of the Report of the Chilean National Commission on Truth and Reconciliation, we seek to understand the logic and dynamics of the violence, deaths, and disappearances under Pinochet’s Chile.
 

### 2019-05-29 manuscript/article/article.Rmd
update manuscript

--- 2019-05-02 rap/article.Rmd
@@
 We coded the personal details of 2,398 victims named in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations.
-The dataset comprises XXXXXXXXXX variables from 1973 to 1990 and is available as a stand-alone spreadsheet or as the `pinochet` package for `R`.
+The dataset comprises 59 variables from 1973 to 1990 and is available as a stand-alone spreadsheet or as the `pinochet` package for `R`.
 As examples, we describe the major temporal and spatial patterns of the human rights abuses.
@@
     number_sections: yes
-    keep_tex: no
+    keep_tex: yes
     template: template.latex
@@
 
-Our dataset comprises 2,398 observations and XXXXXXXXX variables.
+Our dataset comprises 2,398 observations and 59 variables.
 Each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number.
@@
 Table 1 shows an example entry from our dataset.
-
-\newpage
-
-| Variable                | Sample          |
-| :---------------------- | :-------------- |
-| id                      | 276             |
-| page                    | 265             |
-| start date              | 18/01/1974      |
-| end date                | 18/01/1974      |
-| last name               | Marques Agusto  |
-| first name              | Nelson José     |
-| minor                   | 0 (no)          |
-| age                     | 31              |
-| male                    | 1 (yes)         |
-| occupation              | office worker   |
-| victim affiliation      | communist party |
-| violence                | killed          |
-| method                  | gun             |
-| interrogation           | NA (unknown)    |
-| torture                 | NA (unknown)    |
-| mistreatment            | 1 (yes)         |
-| perpetrator             | military        |
-| place                   | in public       |
-| start location          | Iquique         |
-| latitude 1              | -20.23070       |
-| longitude 1             | -70.13567       |
-| place 2                 | in custody      |
-| location 2              | Pisagua prison  |
-| latitude 2              | -19.59667       |
-| longitude 2             | -70.21222       |
-
-Table: Sample event from the dataset (selected variables).
-
-\newpage
 
@@
 To make the data accessible, our GitHub repository has detailed installation instructions for users new to `R`.
+
+\newpage
+
+```{r sample-obs, echo=FALSE, warning=FALSE, message=FALSE,cache=TRUE}
+if (!require("kableExtra")) {
+        install.packages("kableExtra")
+}
+if (!require("tidyverse")) {
+        install.packages("tidyverse")
+}
+load("/Users/politicaltheory/Documents/github/pinochet/data/pinochet.RData")
+kable(t(pinochet[12, c(1:4,7:22,24:37)]), "latex", caption = "Sample event from the dataset (selected variables)", col.names = c("Value"), booktabs = T) %>%
+  kable_styling(position ="center", latex_options = "hold_position")
+```
+
+\vspace{.5cm}
 
@@
 
-| Violence         | _N_    |
-| :--------------- | -----: |
-| Killed           |   1331 |
-| Disappeared      |    853 |
-| Unresolved cases |     93 |
-| Suicides         |      7 |
-| Unknown          |    115 |
+\vspace{.5cm}
 
-Table: Number of killings, disappearances, unresolved cases, and suicides.
+```{r number, echo=FALSE, message=FALSE, warning=FALSE,cache=TRUE}
+kable(table(pinochet$violence), "latex", caption = "Number of killings, disappearances, unresolved cases, and suicides", booktabs = T, col.names = c("Violence", "N")) %>%
+  kable_styling(position ="center", latex_options = "hold_position")
+```
+
+\vspace{.5cm}
 
@@
 
-\vspace{.5cm}
-\begin{center}
-Figure 1 about here.
-
-Two time trends: one with the whole period,\\ another discarding the first year to better visualise variation in later years
-\end{center}
-\vspace{.5cm}
-
-[comment]: <> (Figure 01 - Time Trends Here)
-\newpage
-
 In September 1973 alone, the government committed 633 acts of violence against civilians, of which 406 the Commission signals as "definite and known deaths".
 The violations continued in October, with 504 incidences and 317 certified deaths.
+In total, the Pinochet regime made 1274 victims in its first year.
 The numbers decline to about 10% of those levels in November and December, and then recede further in 1974.
@@
 
+\vspace{.5cm}
+
+```{r time-trend, echo=FALSE, fig.align='center', fig.height=3, fig.width=5, warning=FALSE, message=FALSE, fig.cap='Human rights abuses in the Pinochet regime, 1973-1990',cache=TRUE}
+if (!require("tidyverse")) {
+        install.packages("tidyverse")
+}
+if (!require("lubridate")) {
+        install.packages("lubridate")
+}
+pinochet %>% ungroup() %>% 
+  mutate(Year = year(start_date_monthly)) %>%
+  group_by(Year) %>% tally() %>% 
+  filter(!is.na(Year)) %>% 
+  ggplot(aes(x = Year, y = n)) +
+  geom_line() +
+  theme_minimal() +
+  labs(x = NULL, y = NULL, title = "Number of Human Rights Abuses", 
+       subtitle = "Pinochet Regime, 1973-1990") +
+  scale_y_continuous(breaks = c(0, 500, 1000, 1274))
+```
+
+\vspace{.5cm}
+
 We can also explore how the government employed violence across space.
@@
 
-\vspace{.5cm}
-\begin{center}
-Figure 2 about here.
+```{r maps, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.cap="Spatial variation in human rights abuses in the Pinochet regime, 1973-1990", fig.height=6, fig.width=7,cache=TRUE}
+if (!require("sf")) {
+        install.packages("sf")
+}
+if (!require("devtools")) {
+        install.packages("devtools")
+}
+devtools::install_github("ropensci/rnaturalearthdata")
+if (!require("rnaturalearthhires")) {
+        install.packages("rnaturalearthhires", repos = "http://packages.ropensci.org", type = "source")
+}
+library(rnaturalearthdata)
+library(rnaturalearthhires)
 
-Two maps: deaths and disappearances 
-\end{center}
-\vspace{.5cm}
+chile <- rnaturalearthhires::countries10 %>%
+st_as_sf() %>%
+filter(SOVEREIGNT %in% c("Chile", "Argentina", "Peru", "Paraguay"))
 
-[comment]: <> (Figure 02 - Spatial variation)
+violent_events <- pinochet %>% 
+  select(violence, latitude_1, longitude_1) %>%
+  filter(!is.na(latitude_1), !is.na(longitude_1), !is.na(violence)) %>%
+  st_as_sf(coords = c("longitude_1", "latitude_1"), crs = 5361)
+
+coords_vio <- st_coordinates(violent_events) %>% as_tibble()
+violent_events <- bind_cols(violent_events, coords_vio)
+
+ggplot() +
+  geom_sf(data = chile) +
+  coord_sf(xlim = c(-75.6, -67), ylim = c(-55, -19)) +
+  labs(x = NULL, y = NULL) +
+  geom_point(data = violent_events, aes(x = X, y = Y, colour = violence, fill = violence), shape = 21) + 
+  scale_colour_manual(values = c("#042333FF", "#481567FF",  "#253582FF", "#2D708EFF", "#29AF7FFF")) +
+  scale_fill_manual(values = alpha(c("#042333FF", "#481567FF", "#253582FF", "#2D708EFF", "#29AF7FFF"), .6)) +
+  facet_wrap(~violence, nrow = 1) +
+  theme_minimal() +
+  theme(strip.text = element_blank()) +
+  theme(legend.position = "bottom",
+        axis.text = element_blank()) + 
+  theme(legend.title=element_blank())
+```
 
 We can further disaggregate the analysis and visualise how the regime persecuted particular groups.
-Students and members of left-wing parties are two categories of special interest.
+School and university students are two categories of special interest.
 On 18 September 1073, only one week after the coup, Pinochet and his followers rounded up thousands of students and activists in Chile's National Stadium.
-Military forces tortured and murdered hundreds of prisoners over the course of a few days.
-An emblematic victim was Víctor Lidio Jara Martínez, a university teacher, theatre director and singer (id number 25).
-According to witnesses, Mr Jara had his hand crushed and his wrists broken, and he was later shot 44 times with a machine gun.
-Mr Jara's body was left in the front of the stadium as a warning for other political prisoners [@mcsherry2015victor 56].
+Military forces tortured and murdered dozens of prisoners over the course of a few days.
 The government again tortured and killed students and activists in its campaign against the Revolutionary Left Movement (_Movimiento de Izquierda Revolucionaria_, MIR).
@@
 \vspace{.5cm}
-\begin{center}
-Figure 3 about here.
 
-Time trends: attacks against students and workers
-\end{center}
+```{r time-trend-students, echo=FALSE, fig.align='center', fig.height=3, fig.width=5, warning=FALSE, message=FALSE, fig.cap='Human rights abuses against students, 1973-1990',cache=TRUE}
+
+pinochet %>% ungroup() %>% 
+  mutate(Year = year(start_date_monthly)) %>%
+  group_by(Year) %>% tally() %>% 
+  filter(!is.na(Year)) %>% 
+  ggplot() +
+  geom_line(aes(x = Year, y = n)) +
+  geom_line(aes(x = Year, y = n)) +
+  theme_minimal() +
+  labs(x = NULL, y = NULL, title = "Number of Human Rights Abuses Against Students", 
+       subtitle = "Pinochet Regime, 1973-1990") +
+  scale_y_continuous(breaks = c(0, 500, 1000))
+```
+
 \vspace{.5cm}

### 2019-05-29 manuscript/article/article.tex
update manuscript

--- 2019-03-29 rap/article.tex
@@
       pdfauthor={Danilo Freire; John Meadowcroft; David Skarbek; Eugenia Guerrero},
-      pdfkeywords={authoritarian regimes, Chile, georeferenced event data, Pinochet, truth
+      pdfkeywords={Chile, georeferenced event data, human rights, Pinochet, truth
 commission},
@@
 
-\title{\textbf{Deaths and Disappearances in the Pinochet Regime: A New Dataset}\thanks{We thank Umberto Mignozzetti and Robert Myles McDonnell for their
-helpful comments. Data and replication information are available at
+\title{\textbf{Deaths and Disappearances in the Pinochet Regime: A New Dataset}\thanks{We thank Umberto Mignozzetti, Lucas Mingardi and Robert Myles McDonnell
+for their helpful comments. Replication materials and R source code are
+available at
 \href{http://github.com/danilofreire/pinochet}{\texttt{http://github.com/danilofreire/pinochet}}.}}
@@
 \noindent This article presents a georeferenced dataset on human rights violations
-in General Augusto Pinochet's dictatorship in Chile. We manually coded
-the personal details of 2,398 victims mentioned in the Chilean Truth
-Commission Report and added geographical coordinates for all
-identifiable atrocity locations. Our dataset comprises 57 variables from
-1973 to 1990 and it is available as a stand-alone spreadsheet or as the
-\texttt{pinochet} package for \texttt{R}. The data describe the
-initially high-level then steep decline of human rights violations, the
-extent to which the military government targeted the opposition, and the
-spatial variation of state repression. We also discuss our coding
+in the Pinochet dictatorship in Chile. We coded the personal details of
+2,398 victims named in the Chilean Truth Commission Report and added
+geographical coordinates for all identifiable atrocity locations. The
+dataset comprises 59 variables from 1973 to 1990 and is available as a
+stand-alone spreadsheet or as the \texttt{pinochet} package for
+\texttt{R}. As examples, we describe the major temporal and spatial
+patterns of the human rights abuses. We also discuss our coding
 procedures, show descriptive statistics, and conclude with suggestions
-for potential uses of our dataset.
+for potential applications of the dataset.
 \vspace{.5cm}
 
-\noindent \textbf{Keywords}: authoritarian regimes, Chile, georeferenced event data, Pinochet, truth
+\noindent \textbf{Keywords}: Chile, georeferenced event data, human rights, Pinochet, truth
 commission
@@
 economic liberalisation with large-scale political repression
-\citep{meadowcroft2014hayek, valdes1995pinochet}. In 1991, after Chile's
-transition to democracy, then-President Patricio Aylwin convened a
-commission to investigate human rights abuses in the Pinochet regime.
-The Report of the National Commission on Truth and Reconciliation
+\citep{valdes1995pinochet}. In 1991, after Chile's transition to
+democracy, President Patricio Aylwin convened a commission to
+investigate human rights violations in the Pinochet regime. The Report
+of the National Commission on Truth and Reconciliation
 \citeyearpar{report1991}, also known as the Rettig Report\footnote{Former
-  Chilean ambassador Raúl Rettig chaired the Chilean National Commission
-  for Truth and Reconciliation.}, recorded over 2,000 cases of murders
-and disappearances from 1973 to 1990. The Commission confronted the
-abuses of the past and portrayed a careful reconstruction of state
-repression during the military government. Most important, the Report
-contributed to the erosion of authoritarianism in Chile, as it fostered
-a series of criminal investigations that culminated with the indictment
-of General Pinochet in 2000 \citep[ 26]{brahm2007uncovering}.
+  Chilean Senator and jurist Raúl Rettig chaired the Chilean National
+  Commission for Truth and Reconciliation.}, recorded over 2,000 cases
+of murders and disappearances from 1973 to 1990. The Commission
+confronted the abuses of the past and portrayed a careful reconstruction
+of state violence during the military government. Most importantly, the
+Report contributed to the erosion of authoritarianism in Chile, as it
+fostered a series of criminal investigations that culminated with the
+indictment of General Pinochet in 2000 \citep[ 26]{brahm2007uncovering}.
 
-Although the Report is a valuable resource to scholars, quantitative
-researchers cannot readily use the rich information it contains. In this
-paper, we present a manually-coded dataset with all information from the
-Truth Commission Report plus new variables we collected to complement
-the original data. We transcribed every personal detail from the 903
-pages of the English translation of the Report, assigned a unique
-identification number to each victim, then georeferenced the location of
-the human rights abuses. We mapped all locations that appear in the
-report, such as places used to torture, interrogate or murder the
-victims.
+Although the Report is a valuable source of information for scholars,
+quantitative researchers cannot readily use the rich data it contains.
+In this paper, we present a manually-coded dataset with all information
+from the Truth Commission Report plus new variables we collected to
+complement the original records. We transcribed every personal detail
+from the 903 pages of the English translation of the Report, assigned a
+unique identification number to each victim, then georeferenced the
+location of the human rights abuses. We added coordinates of latitude
+and longitude to every location mentioned in the report, such as places
+used to torture, interrogate or murder the victims.
 
@@
 (1976) prompted frequent human rights violations that only recently have
-been fully documented \citep{agosin1987generals, elpais2016argentina}.
-In some cases, autocrats declared unrestrained violence to be an
-explicit government policy. Peru's former President Alberto Fujimori
-openly stated that his \emph{autogolpe} (1992) granted him the rights to
-``{[}apply{]} drastic punishments'' towards alleged terrorists
+been fully documented
+\citep{agenciabrasil2014ditadura, elpais2016argentina}. In some cases,
+autocrats declared unrestrained violence to be an explicit government
+policy. In 1992, Peru's former President Alberto Fujimori openly stated
+that his \emph{autogolpe} granted him the rights to ``{[}apply{]}
+drastic punishments'' towards any suspect of terrorism
 \citetext{\citealp[97]{dabene1997narcodemocracias}; \citealp[56]{samtleben2013constitucion}}.
 
-The Pinochet regime is also notable for its brutality. On the day of the
-coup, the military opened fire against the \emph{Palacio La Moneda}, the
-Presidential Office, and President Salvador Allende committed suicide
-after troops stormed the building \citep{verdugo2001chile}. The violence
-continued during the military government. The \emph{junta} killed
-thousands of students and members of labour organisations in 1973, with
-up to 200,000 Chileans -- about 2\% of the country's population -- led
-into exile \citep[31]{wright2007chilean}.
+The Pinochet regime was also notable for its brutality. On the day of
+the coup, the military opened fire against the \emph{Palacio La Moneda},
+the Presidential Office, and President Salvador Allende committed
+suicide after troops stormed the building \citep{verdugo2001chile}. The
+violence continued during the military government. The \emph{junta}
+killed thousands of students and members of labour organisations in
+1973, with up to 200,000 Chileans -- about 2\% of the country's
+population -- led into exile \citep[31]{wright2007chilean}.
 
 The establishment of the Truth and Reconciliation Commission was the
-first attempt of
+first attempt to make the dictatorship accountable for its crimes. The
+Commission focused on the most serious cases of human rights violations,
+which are those that resulted in deaths of disappearances\footnote{The
+  Commission's mandate did not extend to cases of forced exile or
+  torture that did not lead to death \citep[660]{ensalaco1994truth}.
+  President Alwyn and his advisors feared the inclusion of non-lethal
+  violence episodes would make the investigation unmanageable
+  \citep[163]{vasallo2002truth}.}. One factor that limited the scope of
+the investigations was that the Commission did not have the legal
+authority to demand collaboration from the military forces
+\citep[85]{popkin1995truth}. The Chilean Commission negotiated an
+amnesty to perpetrator in exchange for freedom of inquiry, and it
+collected testimony from victims in Chile and abroad
+\citep[399]{quinn2001dealing}. The result was a broad yet accurate
+description of the events. Although the military contested the
+conclusions of the Report, they did not question its factual information
+\citetext{\citealp[54]{bakiner2009denial}; \citealp{brahm2005beyond}}.
+
+While the Report did not bring legal punishment to violence
+perpetrators, it addressed important concerns of the victims. In
+February 1992, the Chilean Congress passed a law granting a monthly
+pension to individuals named in the Report, along with psychological
+treatment for their families and school subsidies for their children
+\citep[167]{vasallo2002truth}. The Chilean government also created a new
+office to deal with returning nationals, and President Aylwin released
+almost all political prisoners in the country
+\citep[400]{quinn2001dealing}.
+
+Data from the Truth Commission Report have also fostered academic
+research on the effects of authoritarian violence on social outcomes.
+For instance, Ensalaco
+\citetext{\citeyear{ensalaco1994truth}; \citeyear{ensalaco2000chile}}
+argues that the Report had a positive impact on human rights conditions
+in Chile, while \citet{davis1990they} use the Report to analyse the
+drivers on political violence in Chile. More recently, Esberg
+\citetext{\citeyear{esberg2018antecipating}; \citeyear{esberg2018audience}}
+employs data compiled from the Truth Commission Report to examine how
+the Pinochet regime applied different methods of repression to appeal to
+supporters and avoid political backlash.
+
+Despite these noteworthy examples, quantitative studies on the dynamics
+of authoritarian rule in Chile remains limited. We believe a major
+obstacle for further research in the subject is the lack of
+easily-available, pre-formatted data. Manual text coding is a
+time-consuming process, and data availability would allow scholars to
+investigate new aspects of the Pinochet regime.
 
@@
 
-Our dataset comprises 2,398 observations and 57 variables. Each
+Our dataset comprises 2,398 observations and 59 variables. Each
 observation corresponds to a victim of the Pinochet regime and every
@@
 incidents, as well as information about methods of torture and causes of
-death. Figure 01 shows an example entry from our dataset.
+death. Table 1 shows an example entry from our dataset.
 
@@
 instructions for users new to \texttt{R}.
+
+\newpage
+
+\begin{table}[!h]
+
+\caption{\label{tab:sample-obs}Sample event from the dataset (selected variables)}
+\centering
+\begin{tabular}{ll}
+\toprule
+  & Value\\
+\midrule
+individual\_id & 12\\
+group\_id & 12\\
+start\_date\_daily & 1973-09-12\\
+end\_date\_daily & 1973-09-12\\
+last\_name & Gojanovic Arias\\
+\addlinespace
+first\_name & Drago Vinko\\
+minor & 0\\
+age & 23\\
+male & 1\\
+occupation & Blue Collar\\
+\addlinespace
+occupation\_detail & driver at German Democratic Republic embassy\\
+victim\_affiliation & Opposition\\
+victim\_affiliation\_detail & Communist party\\
+violence & Killed\\
+method & Gun\\
+\addlinespace
+interrogation & NA\\
+torture & NA\\
+mistreatment & NA\\
+targeted & NA\\
+press & 0\\
+\addlinespace
+number\_previous\_arrests & NA\\
+perpetrator\_affiliation & Regime\\
+perpetrator\_affiliation\_detail & Military patrol\\
+nationality & Chilean\\
+place\_1 & Home\\
+\addlinespace
+location\_1 & Las Condes\\
+latitude\_1 & -33.39901\\
+longitude\_1 & -70.55731\\
+exact\_coordinates\_1 & 0\\
+place\_2 & In Public\\
+\addlinespace
+location\_2 & Intersection of Calle Tabancura and Avenida Kennedy\\
+latitude\_2 & -33.38366\\
+longitude\_2 & -70.53468\\
+exact\_coordinates\_2 & 1\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\vspace{.5cm}
 
@@
 information from pages 159-160 of the Truth Commission Report.
+
+\vspace{.5cm}
+
+\begin{table}[!h]
+
+\caption{\label{tab:number}Number of killings, disappearances, unresolved cases, and suicides}
+\centering
+\begin{tabular}{lr}
+\toprule
+Violence & N\\
+\midrule
+Disappearance & 853\\
+Disappearance, Information of Death & 108\\
+Killed & 1331\\
+Suicide & 8\\
+Unresolved & 93\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\vspace{.5cm}
 
@@
 As mentioned previously, our dataset includes 57 variables and covers
-the seventeen years of the Pinochet Presidency (1973--1990). Figure
-\ref{trend} shows the frequency of government-perpetrated incidents per
-month from 1973 to 1990. We see that the majority of the acts of
-violence happened in the first two months after the military coup. The
-high incidence of human rights violations in the beginning of the
-Pinochet era reflects the activity of the ``Caravan of Death'', a group
-of paramilitaries established in October 1973 to torture and kill
-dissidents
+the seventeen years of the Pinochet Presidency (1973--1990). Figure 1
+shows the frequency of government-perpetrated incidents per month from
+1973 to 1990. We see that the majority of the acts of violence happened
+in the first two months after the military coup. The high incidence of
+human rights violations in the beginning of the Pinochet era reflects
+the activity of the ``Caravan of Death'', a group of paramilitaries
+established in October 1973 to torture and kill dissidents
 \citetext{\citealp[459]{davis1990they}; \citealp{verdugo2001chile}}.
@@
 known deaths''. The violations continued in October, with 504 incidences
-and 317 certified deaths. The numbers decline to about 10\% of those
+and 317 certified deaths. In total, the Pinochet regime made 1274
+victims in its first year. The numbers decline to about 10\% of those
 levels in November and December, and then recede further in 1974.
@@
 
+\vspace{.5cm}
+
+\begin{figure}
+
+{\centering \includegraphics{article_files/figure-latex/time-trend-1}
+
+}
+
+\caption{Human rights abuses in the Pinochet regime, 1973-1990}\label{fig:time-trend}
+\end{figure}
+
+\vspace{.5cm}
+
 We can also explore how the government employed violence across space.
@@
 
+\begin{figure}
+
+{\centering \includegraphics{article_files/figure-latex/maps-1}
+
+}
+
+\caption{Spatial variation in human rights abuses in the Pinochet regime, 1973-1990}\label{fig:maps}
+\end{figure}
+
 We can further disaggregate the analysis and visualise how the regime
-persecuted particular groups. Students and members of left-wing parties
-are two categories of special interest. On 18 September 1073, only one
-week after the coup, Pinochet and his followers rounded up thousands of
+persecuted particular groups. School and university students are two
+categories of special interest. On 18 September 1073, only one week
+after the coup, Pinochet and his followers rounded up thousands of
 students and activists in Chile's National Stadium. Military forces
-tortured and murdered hundreds of prisoners over the course of a few
-days. An emblematic victim was Víctor Lidio Jara Martínez, a university
-teacher, theatre director and singer (id number 25). According to
-witnesses, Mr Jara had his hand crushed and his wrists broken, and he
-was later shot 44 times with a machine gun. Mr Jara's body was left in
-the front of the stadium as a warning for other political prisoners
-\citep[ 56]{mcsherry2015victor}. The government again tortured and
-killed students and activists in its campaign against the Revolutionary
-Left Movement (\emph{Movimiento de Izquierda Revolucionaria}, MIR). The
-MIR originated in student unions and in the late 1970s the movement
-promoted a series of attacks against government personnel and official
-buildings. Government forces strongly retaliated and the group resorted
-to terrorist tactics to confront the regime
-\citep{schlotterbeck2018beyond}.
+tortured and murdered dozens of prisoners over the course of a few days.
+The government again tortured and killed students and activists in its
+campaign against the Revolutionary Left Movement (\emph{Movimiento de
+Izquierda Revolucionaria}, MIR). The MIR originated in student unions
+and in the late 1970s the movement promoted a series of attacks against
+government personnel and official buildings. Government forces strongly
+retaliated and the group resorted to terrorist tactics to confront the
+regime \citep{schlotterbeck2018beyond}.
+
+\vspace{.5cm}
+
+\begin{figure}
+
+{\centering \includegraphics{article_files/figure-latex/time-trend-students-1}
+
+}
+
+\caption{Human rights abuses against students, 1973-1990}\label{fig:time-trend-students}
+\end{figure}
+
+\vspace{.5cm}
 

### 2019-05-29 R/script.R

@@ -1,369 +1,0 @@
-###########################################################################
-############ Deaths and Disappearances in the Pinochet Regime #############
-### Danilo Freire, John Meadowcroft, David Skarbek and Eugenia Guerrero ###
-###########################################################################
-
-## Load required packages
-if (!require("plyr")) {
-        install.packages("plyr")
-}
-if (!require("tidyverse")) {
-        install.packages("tidyverse")
-}
-if (!require("devtools")) {
-        install.packages("devtools")
-}
-if (!require("narnia")) {
-        install_github("njtierney/narnia")
-}
-if (!require("ggplot2")) {
-        install.packages("ggplot2")
-}
-if (!require("rnaturalearth")) {
-        install.packages("rnaturalearth")
-}
-
-## Load data
-df <- readr::read_csv("https://goo.gl/F1tgk1")
-
-## Data manipulation
-# Convert names to lowercase
-names(df) <- tolower(names(df))
-
-# Rename variables
-df <- plyr::rename(df,
-                   c("merged id" = "merged_id",
-                     "start date" = "start_date",
-                     "end date" = "end_date",
-                     "last name" = "last_name",
-                     "first name" = "first_name",
-                     "sex (c)" = "female",
-                     "occupation (c)" = "occupation",
-                     "occupation" = "occupation_2",                     
-                     "victim affiliation (c)" = "victim_affiliation",
-                     "victim affiliation" = "victim_affiliation2",
-                     "violence (c)" = "violence",
-                     "method (c)" = "method",
-                     "interrogation (c)" = "interrogation",
-                     "torture (c)" = "torture",
-                     "war tribunal" = "war_tribunal",
-                     "previous arrests (#)" = "number_previous_arrests",
-                     "perpetrator affiliation (c)" = "perpetrator_affiliation",
-                     "place-1" = "place_1",
-                     "start location-1" = "start_location_1",
-                     "latitude-1" = "latitude_1",
-                     "longitude-1" = "longitude_1",
-                     "exact coordinates-1" = "exact_coordinates_1",
-                     "place-2" = "place_2",
-                     "location-2" = "location_2",
-                     "latitude-2" = "latitude_2",
-                     "longitude-2" = "longitude_2",
[…]

### 2019-05-29 data-parsing.R

@@ -1,131 +1,0 @@
-########################
-### Parsing the data ###
-########################
-
-## Install required R packages
-if (!require("plyr")) {
-    install.packages("plyr")
-}
-if (!require("tidyverse")) {
-    install.packages("tidyverse")
-}
-if (!require("devtools")) {
-    install.packages("devtools")
-}
-if (!require("narnia")) {
-    install_github("njtierney/narnia")
-}
-if (!require("rnaturalearth")) {
-    install.packages("rnaturalearth")
-}
-
-## Load raw data
-df <- readxl::read_xlsx("data/raw-data.xlsx")
-
-
-## Data manipulation
-# Convert names to lowercase
-names(df) <- tolower(names(df))
-
-# Rename variables
-df <- plyr::rename(df,
-                   c("merged id" = "merged_id",
-                     "start date" = "start_date",
-                     "end date" = "end_date",
-                     "last name" = "last_name",
-                     "first name" = "first_name",
-                     "sex (c)" = "male",
-                     "occupation (c)" = "occupation",
-                     "occupation" = "occupation_2",                     
-                     "victim affiliation (c)" = "victim_affiliation",
-                     "victim affiliation" = "victim_affiliation2",
-                     "violence (c)" = "violence",
-                     "method (c)" = "method",
-                     "interrogation (c)" = "interrogation",
-                     "torture (c)" = "torture",
-                     "war tribunal" = "war_tribunal",
-                     "previous arrests (#)" = "number_previous_arrests",
-                     "perpetrator affiliation (c)" = "perpetrator_affiliation",
-                     "place-1" = "place_1",
-                     "start location-1" = "start_location_1",
-                     "latitude-1" = "latitude_1",
-                     "longitude-1" = "longitude_1",
-                     "exact coordinates-1" = "exact_coordinates_1",
-                     "place-2" = "place_2",
-                     "location-2" = "location_2",
-                     "latitude-2" = "latitude_2",
-                     "longitude-2" = "longitude_2",
-                     "exact coordinates-2" = "exact_coordinates_2",
-                     "place ©-3" = "place_3",
-                     "end location-3" = "end_location_3",
-                     "latitude-3" = "latitude_3",
-                     "longitude-3" = "longitude_3",
-                     "exact coordinates-3" = "exact_coordinates_3",
-                     "place (c)-4" = "place_4",
-                     "end location-4" = "end_location_4",
[…]

### 2019-05-30 manuscript/article/article.Rmd
v 0.1.0

@@
   select(violence, latitude_1, longitude_1) %>%
-  filter(!is.na(latitude_1), !is.na(longitude_1), !is.na(violence)) %>%
+  filter(!is.na(violence), !is.na(latitude_1), !is.na(longitude_1)) %>%
   st_as_sf(coords = c("longitude_1", "latitude_1"), crs = 5361)
@@
 
-We can further disaggregate the analysis and visualise how the regime persecuted particular groups.
-School and university students are two categories of special interest.
-On 18 September 1073, only one week after the coup, Pinochet and his followers rounded up thousands of students and activists in Chile's National Stadium.
-Military forces tortured and murdered dozens of prisoners over the course of a few days.
-The government again tortured and killed students and activists in its campaign against the Revolutionary Left Movement (_Movimiento de Izquierda Revolucionaria_, MIR).
-The MIR originated in student unions and in the late 1970s the movement promoted a series of attacks against government personnel and official buildings.
-Government forces strongly retaliated and the group resorted to terrorist tactics to confront the regime [@schlotterbeck2018beyond].
+We can further disaggregate the analysis and visualise how particular groups perpetrated violence during the regime.
+Figure~\ref{fig:time-trend-perpetrators} shows how the regime forces and the opposition attacked each other after the consolidation of the military rule.
+The mid-1970s is a period of severe state repression, but the government violence lessens around the end of the decade.
+In 1977, the National Intelligence Directorate (_Dirección de Inteligencia Nacional_), the Chilean secret police, was dissolved after killing the main leaders of the Revolutionary Left Movement (_Movimiento de Izquierda Revolucionaria_, MIR), a left-wing guerrilla organisation [@loveman1986military, 12].
+In contrast, the number of atrocities by the opposition increases in the late 1970s and early 1980s, when the MIR regrouped and promoted a series of attacks against government personnel and official buildings.
+In the mid-1980s, we see a retaliation from the Chilean government against a wave of protests in the wake of an economic crisis.
+State violence would only recede in the years preceding the democratisation [@valdivia2003terrorism, 188].
 
@@
 
-```{r time-trend-students, echo=FALSE, fig.align='center', fig.height=3, fig.width=5, warning=FALSE, message=FALSE, fig.cap='Human rights abuses against students, 1973-1990',cache=TRUE}
-
-pinochet %>% ungroup() %>% 
+```{r time-trend-perpetrators, echo=FALSE, fig.align='center', fig.height=4, fig.width=5, warning=FALSE, message=FALSE, fig.cap='Violence perpetrators in the Pinochet regime, 1973-1990'}
+# Load requied packages
+if (!require("tidyverse")) {
+        install.packages("tidyverse")
+}
+if (!require("lubridate")) {
+        install.packages("lubridate")
+}
+# Plot
+pinochet %>% 
+  select(c(perpetrator_affiliation, start_date_monthly)) %>%
   mutate(Year = year(start_date_monthly)) %>%
-  group_by(Year) %>% tally() %>% 
+  filter(!is.na(perpetrator_affiliation)) %>% 
   filter(!is.na(Year)) %>%
-  ggplot() +
-  geom_line(aes(x = Year, y = n)) +
-  geom_line(aes(x = Year, y = n)) +
+  filter(Year >= 1975) %>%
+  group_by(perpetrator_affiliation, Year) %>%
+  summarise(no_rows = length(perpetrator_affiliation)) %>%
+  ggplot(aes(x = Year, y = no_rows, colour = perpetrator_affiliation)) +
+  viridis::scale_color_viridis(discrete = TRUE,
+                               end = 0.5, direction = -1) + 
+  geom_line() +
   theme_minimal() +
-  labs(x = NULL, y = NULL, title = "Number of Human Rights Abuses Against Students", 
-       subtitle = "Pinochet Regime, 1973-1990") +
-  scale_y_continuous(breaks = c(0, 500, 1000))
+  labs(x = NULL, y = NULL, title = "Violence Perpetrators", 
+       subtitle = "Pinochet Regime, 1975-1990") +
+  scale_x_continuous(breaks = c(1975, 1980, 1985, 1990)) +
+  facet_grid(perpetrator_affiliation ~ ., scales = "fixed") + 
+  theme(legend.position = "bottom") + 
+  theme(legend.title=element_blank())
 ```
-
-\vspace{.5cm}
-
-[comment]: <> (Figure 03 - Students and Communist Leaders)
 
@@
 
-[^eg]: Software Developer, Tempo UK.
+[^eg]: Software Developer, Attest, UK.
 

### 2019-05-30 manuscript/online-appendix/online-appendix.Rmd
v 0.1.0

---
title:  |
        | **Data Codebook and Appendix for**
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author: 
  - "Danilo Freire[^df]"
  - "John Meadowcroft[^jm]"
  - "David Skarbek[^ds]"
  - "Eugenia Guerrero[^eg]"
date: \today
thanks: "Replication materials and R source code are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    toc: true
    number_sections: yes
    keep_tex: no
    template: template.latex
---

\newpage

# Introduction

\setlength{\parindent}{0pt}

This document contains the code necessary to replicate the results in "Deaths and Disappearances in the Pinochet Regime: A New Dataset" by @freire2019dataset. The data are available as the `pinochet` package for the `R` statistical language [@freire2019pinochet] and as an Excel spreadsheet at <http://github.com/danilofreire/pinochet>.

# Background

The violent military coup in Chile in September 1973 resulted in over three decades of sustained violence, deaths, and forced disappearances as General Pinochet’s government fought against left-wing movements for the sake of the 'national wellbeing'. The fall of Pinochet’s regime and the traumatic consequences of his authoritarian government left Chile deeply divided. In order to move forward and avoid a repetition of the past, the Chilean government highlighted the need of an impartial and comprehensive report detailing the extent of the deaths and disappearances.

By compiling and analysing the data from both volumes of the Report of the Chilean National Commission on Truth and Reconciliation, we seek to understand the logic and dynamics of the violence, deaths, and disappearances under Pinochet’s Chile.

# Sources

Report of the Chilean National Commission on Truth and Reconciliation, Vol. 1.

Report of the Chilean National Commission on Truth and Reconciliation, Vol. 2.

# Variables

The dataset has 59 variables with information about the victims, the perpetrators, and geographical coordinates of each incident. We provide a brief description of each variable below. Here, we sorted them alphabetically.

* `additional_comments`: This variable gives additional information about cases that are too ambiguous to fit precisely into a single category. For instance, when the Truth and Commission Report is unsure whether the victim was killed or committed suicide, or if a given human rights abuse may have occurred in Santiago or elsewhere. Only 10 cases have additional information.

* `age`: Victim's age, in years. The variable ranges from 0.22 (about 4 months) to 85. The mean is 29.7 years.

* `end_date_daily`: Day when the human rights abuse ended.

* `end_date_monthly`: Not all instances of human rights abuses have exact dates. This variable indicates the month where the atrocity occurred, when this information is available. We used the first day of the month as a reference.

* `exact_coordinates_n`: We matched the event sites with coordinates of latitude and longitude. As the report does not have the precise location of all events, we used the closest reference available. This is a dummy variable stating whether coordinates are precise (street level) or not. 1 = yes. There are six variables in the dataset, each pertaining to one location where the individual was found or taken to.

* `first_name`: Victim's first name.

* `group_id`: Unique identification number for groups of individuals attacked together.

* `individual_id`: Unique individual identification number.

* `interrogation`: Individual was interrogated, 1 = yes.

* `last_name`: Victim's last name.

* `latitude_n`: Latitudinal coordinates pertaining to the string `place_n` variable.

* `location_n`: Where the individual was seen or found. There are up to 6 locations, so we coded them as `location_1` to `location_6`. The same pattern repeats in the variables below. The compilation of the `location_n` variable was based completely on information given in the Truth Report. However, since this information was in a string format (e.g. intersection of Calle Grecia and Avenida Rosa), creating a new variable incorporating each location’s latitude and longitude was necessary to pursue further analysis of the trends in deaths and disappearances. The format chosen was decimal coordinates. Three main websites were used to find the decimal coordinates:
  - Latitude: http://latitude.to/. We used Latitude for sourcing more inexact coordinates, such as the coordinates for cities or towns. We also used it to convert non-decimal coordinates into the decimal format.
   - Wikimapia: http://wikimapia.org/. Wikimapia was mostly used for locating shantytowns, neighborhoods, military bases, or police stations that were not found on Google maps. Its strength is that it is an editable map, meaning that individual users can collaborate so as to produce a more comprehensive map (filled with local knowledge) than other standard maps such as Google maps.
  - Google maps: http://maps.google.co.uk/. Google Maps was useful for finding intersections of roads or precise addresses, as well as for searching for police stations, some military bases, and other governmental buildings.

* `longitude_n`: Longitudinal coordinates pertaining to the string `place_n` variable.

* `male`: Victim is male, 1 = yes.

* `method`: Method of violence used by the perpetrator. There are 25 categories.

* `minor`: Victim is a minor, 1 = yes.

* `mistreatment`: Individual was mistreated by the government or the opposition, 1 = yes.

* `nationality`: Individual's nationality. They are:
  - Argentinean, Austrian, Bolivian, Brazilian, Chilean, Dual British and Chilean, Dual French and Chilean, Dual Swiss and Chilean, Ecuadorian, French, Mexican, Spanish, Uruguayan, Vietnamese

* `number_previous_arrests`: Number of individual's previous arrests. 0 indicates that the individual was reported to the authorities and was allowed to go free (no arrest).

* `occupation`: Victim occupation. We grouped all occupations into 10 categories:
  - White collar, Blue collar, School student, University student, Military, Non-military government, Unemployed, Unknown, Ex-military, Housewife

* `occupation_detail`: Details about victim's occupation, when available. The details usually describe the company or public sector office where the individual worked.

* `page`: Denotes which page of the English version of the Truth and Commission Report describes the victim.

* `perpetrator_affiliation`: Perpetrator's political affiliation. It is a binary variable: either individuals are in the government or in the opposition. We follow the Report's broad definition the opposition: “private citizens who committed terrorist acts or other kinds of politically motivated attacks” (p.43). Thus, the opposition includes supporters of Presidente Allende, terrorist groups, etc.

* `perpetrator_affiliation_detail`: Details about perpetrator's political affiliation. As with `occupation_detail`, this variable contains extra information about the perpetrator.

* `place_n`: Place where the individual was spotted/reported to be seen. (in chronological order, from 1 to 6 places). Categories:
  - Home; Work; University; In custody; In public; In hospital; Unknown

* `press`: Individual disappearance or death appeared in the press.

* `start_date_daily`: Date when the victim suffered abuse.

* `start_date_monthly`: Calendar month when victim suffered abuse. Similar to what we did in `end_date_monthly`, we used the first day of the month as a reference.

* `victim_affiliation`: individual's political affiliation. Categories:
  - Opposition (Allende/Left): this category encompasses individuals loyal to the ex-president Allende, as well as individuals who were affiliated or leaders of left-wing groups
  - Regime (Pinochet): this category encompasses individuals who belonged to the Pinochet government.
  - Unknown: this category refers to individuals whose political affiliation was unknown or unspecified.
  - None: this category pertains to individuals whom the report states had no known political affiliation.

* `victim_affiliation_detail`: details about the individual's political affiliation, when available. This complements the previous variable by giving as much information as possible on the victims’ political affiliations.

* `violence`: Type of violence. Categories: 
  - Killed: cases where the Commission signals a definite and known death.
  - Suicide
  - Disappearance: cases where the victims are presumed to be dead at the hands of government agents. Government agents are assumed to have disposed of their bodies in a covert fashion, making their retrieval impossible at the time of the publication of the Report.
  - Disappearance, information of death: cases which the Commission has information that signals that “the victims are dead; that they died at the hands of the government agents, or persons in their service; and that these or other agents disposed of the victims’ mortal remains by throwing them into a river or a sea, by covertly burying them, or by disposing of them in some other secret fashion” (p.44).
  - Unresolved: insufficient information or evidence

* `targeted`: Individual was targeted. Categories:
  - Denounced; Summoned; Listed; Voluntarily reported; At home

* `torture`: individual was tortured, 1 = yes. The definition of torture the Report uses is the following: “[...] torture means any act by which severe pain or suffering, whether physical or mental, is intentionally inflicted on a person for such purposes as obtaining from him or a third person information or confession, punishing him for an act he or a third person has committed or is suspected of having committed, or intimidating or coercing him or a third person, or for any reason based on discrimination of any kind, when such pain or suffering is inflicted by or at the instigation of, or with the consent or acquiescence of a public official or other person acting in an official capacity. It does not include pain or suffering arising only from, inherent in or incidental to, lawful sanctions” (pp. 38-39).

* `war_tribunal`: Individual was subject to a war tribunal, 1 = yes.

\newpage

# Replication Code

## Table 01 - Sample Event

The code below downloads the `pinochet` package, loads the dataset into `R` and reproduces table 01 in the article.

```{r sample-obs-no-eval, eval=FALSE, warning=FALSE, message=FALSE, tidy=FALSE}
## Download the pinochet package
if (!require("pinochet")) {
        install.packages("pinochet")
}

# Load the package and the dataset
library(pinochet)
data(pinochet)

# Load required packages
if (!require("kableExtra")) {
        install.packages("kableExtra")
}
if (!require("tidyverse")) {
        install.packages("tidyverse")
}

# Table 01
kable(t(pinochet[12, c(1:4,7:22,24:37)]), "latex",
      caption = "Sample event from the dataset (selected variables)",
      col.names = c("Value"), booktabs = T) %>%
  kable_styling(position ="center", latex_options = "hold_position")
```

```{r sample-obs01, echo=FALSE, warning=FALSE, message=FALSE}
if (!require("kableExtra")) {
        install.packages("kableExtra")
}
if (!require("tidyverse")) {
        install.packages("tidyverse")
}
load("/Users/politicaltheory/Documents/github/pinochet/data/pinochet.RData")

kable(t(pinochet[12, c(1:4,7:22,24:37)]), "latex", 
      caption = "Sample event from the dataset (selected variables)",
      col.names = c("Value"), booktabs = T) %>%
  kable_styling(position ="center", latex_options = "hold_position")
```

\newpage

## Table 02 - Number of Events

```{r number,  message=FALSE, warning=FALSE,cache=TRUE}
kable(table(pinochet$violence), "latex",
      caption = "Number of killings, disappearances, unresolved cases, and suicides",
      booktabs = T, col.names = c("Violence", "N")) %>%
  kable_styling(position ="center", latex_options = "hold_position")
```

## Figure 01 - Number of Human Rights Abuses

The code below reproduces figure 01. It shows the aggregate number of victims in the regime (1973--1990).

```{r time-trend, fig.align='center', fig.height=3, fig.width=5, warning=FALSE, message=FALSE, fig.cap='Human rights abuses in the Pinochet regime, 1973-1990'}
# Install necessary packages
if (!require("tidyverse")) {
        install.packages("tidyverse")
}
if (!require("lubridate")) {
        install.packages("lubridate")
}

# Plot
pinochet %>% 
  ungroup() %>% 
  mutate(Year = year(start_date_monthly)) %>%
  group_by(Year) %>% tally() %>% 
  filter(!is.na(Year)) %>% 
  ggplot(aes(x = Year, y = n)) +
  geom_line() +
  theme_minimal() +
  labs(x = NULL, y = NULL, title = "Number of Human Rights Abuses", 
       subtitle = "Pinochet Regime, 1973-1990") +
  scale_y_continuous(breaks = c(0, 500, 1000, 1274))
```

## Figure 02 - Spatial Variation

```{r maps, warning=FALSE, message=FALSE, fig.align='center', fig.cap="Spatial variation in human rights abuses in the Pinochet regime, 1973-1990", fig.height=6, fig.width=7, tidy=FALSE}
# Load necessary packages
if (!require("sf")) {
        install.packages("sf")
}
if (!require("devtools")) {
        install.packages("devtools")
}
devtools::install_github("ropensci/rnaturalearthdata")
if (!require("rnaturalearthhires")) {
        install.packages("rnaturalearthhires",
                         repos = "http://packages.ropensci.org",
                         type = "source")
}
library(rnaturalearthdata)
library(rnaturalearthhires)

# Save events
chile <- rnaturalearthhires::countries10 %>%
st_as_sf() %>%
filter(SOVEREIGNT %in% c("Chile", "Argentina", "Peru", "Paraguay"))

violent_events <- pinochet %>% 
  select(violence, latitude_1, longitude_1) %>%
  filter(!is.na(violence), !is.na(latitude_1), !is.na(longitude_1)) %>%
  st_as_sf(coords = c("longitude_1", "latitude_1"), crs = 5361)

coords_vio <- st_coordinates(violent_events) %>% as_tibble()
violent_events <- bind_cols(violent_events, coords_vio)

# Create map
ggplot() +
  geom_sf(data = chile) +
  coord_sf(xlim = c(-75.6, -67), ylim = c(-55, -19)) +
  labs(x = NULL, y = NULL) +
  geom_point(data = violent_events, aes(x = X, y = Y,
                                        colour = violence,
                                        fill = violence),
             shape = 21) + 
  scale_colour_manual(values = c("#042333FF", "#481567FF",
                                 "#253582FF", "#2D708EFF",
                                 "#29AF7FFF")) +
  scale_fill_manual(values = alpha(c("#042333FF", "#481567FF",
                                     "#253582FF", "#2D708EFF",
                                     "#29AF7FFF"), .6)) +
  facet_wrap(~violence, nrow = 1) +
  theme_minimal() +
  theme(strip.text = element_blank()) +
  theme(legend.position = "bottom",
        axis.text = element_blank()) + 
  theme(legend.title=element_blank())
```

## Figure 03 - Violence Perpetrators

```{r time-trend-perpetrators, fig.align='center', fig.height=4, fig.width=5, warning=FALSE, message=FALSE, fig.cap='Violence perpetrators in the Pinochet regime, 1973-1990'}
# Load requied packages
if (!require("tidyverse")) {
        install.packages("tidyverse")
}

# Plot
pinochet %>% 
  select(c(perpetrator_affiliation, start_date_monthly)) %>%
  mutate(Year = year(start_date_monthly)) %>%
  filter(!is.na(perpetrator_affiliation)) %>% 
  filter(!is.na(Year)) %>%
  filter(Year >= 1975) %>%
  group_by(perpetrator_affiliation, Year) %>%
  summarise(no_rows = length(perpetrator_affiliation)) %>%
  ggplot(aes(x = Year, y = no_rows, colour = perpetrator_affiliation)) +
  viridis::scale_color_viridis(discrete = TRUE, end = 0.5, direction = -1) + 
  geom_line() +
  theme_minimal() +
  labs(x = NULL, y = NULL, title = "Violence Perpetrators", 
       subtitle = "Pinochet Regime, 1975-1990") +
  scale_x_continuous(breaks = c(1975, 1980, 1985, 1990)) +
  facet_grid(perpetrator_affiliation ~ ., scales = "fixed") + 
  theme(legend.position = "bottom") + 
  theme(legend.title=element_blank())

```

\newpage

# Session Information

```{r session}
sessionInfo()
```

[^df]: Postdoctoral Research Fellow, the Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), [`http://danilofreire.github.io`](http://danilofreire.github.io). Corresponding author.

[^jm]: Senior Lecturer in Public Policy, Department of Political Economy, King's College London, [`john.meadowcroft@kcl.ac.uk`](mailto:john.meadowcroft@kcl.ac.uk), [`http://johnmeadowcroft.net`](http://johnmeadowcroft.net).

[^ds]: Associate Professor, Department of Political Science and the Political Theory Project, Brown University, [`david_skarbek@brown.edu`](david_skarbek@brown.edu), [`http://davidskarbek.com`](http://davidskarbek.com).

[^eg]: Software Developer, Attest, UK.

\setlength{\parindent}{0cm}
\setlength{\parskip}{6pt}

### 2019-05-31 R/pinochet.R
fix issues cran

@@ -2,7 +2,7 @@
 #'
 #' A dataset containing information about victims of the Pinochet regime
 #' in Chile (1973-1990). The data comes from the National Truth and Reconciliation
-#' Commission Report (1991). We included georeferenced information about all incidents
+#' Commission Report (1991, ISBN:9780268016463). We included georeferenced information about all incidents
 #' where the information could be retrieved. More information about the coding process
 #' is available at http://github.com/danilofreire/pinochet.
 #'
@@ -14,8 +14,8 @@
 #' \describe{
 #'   \item{additional_comments}{any further information on the victim}
 #'   \item{age}{victim's age}
-#'   \item{end_date_daily}{date when the human rights abuse ended}
-#'   \item{end_date_monthly}{calendar month when human rights abuse end}
+#'   \item{end_date_daily}{calendar day when the human rights abuse ended}
+#'   \item{end_date_monthly}{calendar month when human rights abuse ended}
 #'   \item{exact_coordinates_1}{exactitude of coordinates for place_1, 1 = exact}
 #'   \item{exact_coordinates_2}{exactitude of coordinates for place_2, 1 = exact}
 #'   \item{exact_coordinates_3}{exactitude of coordinates for place_3, 1 = exact}
@@ -51,9 +51,9 @@
 #'   \item{mistreatment}{individual was mistreated, 1 = yes}
 #'   \item{nationality}{individual's nationality}
 #'   \item{number_previous_arrests}{number of individual's previous arrests}
-#'   \item{occupation}{victim occupation}
+#'   \item{occupation}{victim's occupation}
 #'   \item{occupation_detail}{details about victim's occupation, when available}
-#'   \item{page}{report page which describes the victim}
+#'   \item{page}{report page where we retrieved information about the victim}
 #'   \item{perpetrator_affiliation}{perpetrator's political affiliation}
 #'   \item{perpetrator_affiliation_detail}{details about perpetrator's political affiliation}
 #'   \item{place_1}{place where the individual was spotted/reported to be seen (in chronological order, from 1 to 6 places)}

### 2019-06-03 manuscript/article/article.Rmd
update R code manuscript

@@
 The files follow the principles of "tidy data", where each column represents one variable and each row is one case [@wickham2014tidy].
-The `pinochet` package is available for download on the Comprehensive R Archive Network (CRAN) and at <http://github.com/danilofreire/pinochet>.
+The `pinochet` package is available for download on the [Comprehensive R Archive Network (CRAN)](https://cran.r-project.org/package=pinochet) and at <http://github.com/danilofreire/pinochet>.
 To make the data accessible, our GitHub repository has detailed installation instructions for users new to `R`.
@@
 
-```{r sample-obs, echo=FALSE, warning=FALSE, message=FALSE,cache=TRUE}
+```{r sample-obs, echo=FALSE, warning=FALSE, message=FALSE}
 if (!require("kableExtra")) {
@@
 }
-load("/Users/politicaltheory/Documents/github/pinochet/data/pinochet.RData")
+if (!require("pinochet")) {
+        install.packages("pinochet")
+}
 kable(t(pinochet[12, c(1:4,7:22,24:37)]), "latex", caption = "Sample event from the dataset (selected variables)", col.names = c("Value"), booktabs = T) %>%
@@
   scale_x_continuous(breaks = c(1975, 1980, 1985, 1990)) +
+  scale_y_continuous(breaks = c(0, 50, 100, 135)) + 
   facet_grid(perpetrator_affiliation ~ ., scales = "fixed") + 

### 2019-06-03 manuscript/online-appendix/online-appendix.Rmd
update R code manuscript

@@
 }
-load("/Users/politicaltheory/Documents/github/pinochet/data/pinochet.RData")
+if (!require("pinochet")) {
+        install.packages("pinochet")
+}
 
@@
 }
+if (!require("lubridate")) {
+        install.packages("lubridate")
+}
 

### 2021-03-01 manuscript/article.Rmd
add intro

--- 2019-03-08 rap/article.Rmd
@@
 title:  |
-        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
+        | Human Rights Violations in the Pinochet Regime:
+        | A Quantitative Presentation of the Rettig and Valech Reports
 author: 
-  - "Danilo Freire[^df]"
-  - "John Meadowcroft[^jm]"
-  - "David Skarbek[^ds]"
-  - "Eugenia Guerrero[^eg]"
+  - Mauricio Carrasco[^auth01]
+  - Danilo Freire[^auth02]
+  - Pablo Seguel Gutierrez[^auth03]
 date: \today
-thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments.
-Data and replication information are available at [`http://github.com/danilofreire/pinochet-dataset`](http://github.com/danilofreire/pinochet-dataset)."
-abstract: "... However, . This paper presents a georeferenced event dataset on "
-keywords: "authoritarian regimes, Chile, geocoded event data, Pinochet, truth commission"
-fontfamily: libertine
-fontawesome: yes
+thanks: "The authors thank Eugenia Guerrero, John Meadowcroft, and David Skarbek for their work on the Rettig Report dataset.
+All replication materials and the source code for the `pinochet R` package are available at <https://github.com/danilofreire/pinochet>."
+abstract: "This article introduces a new dataset on human rights violations during the Pinochet regime in Chile (1973-1990).
+Based on previous work by @freire2019dataset, we code all cases of deaths and disappearances recorded in the Rettig (1991), Valech I (2004), and Valech II (2005) reports.
+Our dataset includes XX variables and presents information on XXXX individuals, of which XXX appear in at least two reports.
+We also provide geographical coordinates for all locations of human rights abused mentioned in the three reports.
+The data are available in several formats, including an `R` package built especifically for our purposes.
+The paper shows the temporal and spatial variation of human rights violations during the regime, and proposes further avenues of research for those interested in Chilean history and cross-national studies of state-sponsored violence."
+abstractspacing: double
+keywords: Chile; human rights violations; Pinochet; state-sponsored violence; truth commission
+jelcodes: 
 fontsize: 12pt
-monospace-url: yes
+margin: 2cm
+urlcolor: darkblue
+linkcolor: Mahogany
+citecolor: Mahogany
 spacing: double
@@
     keep_tex: no
-    template: template.latex
+    toc: no
+    toc_depth: 3
+    template: article-template.latex
 ---
 
-# Introduction\label{sec:intro}
+```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
+# If you need to install any package while knitting the document
+r <- getOption("repos")
+r["CRAN"] <- "https://cran.rstudio.com/"
+options(repos = r)
 
-\setlength{\parindent}{1cm}
-\setlength{\parskip}{0pt}
+# Chunk definitions
+def_chunk_hook  <- knitr::knit_hooks$get("chunk")
+knitr::knit_hooks$set(chunk = function(x, options) {
+  x <- def_chunk_hook(x, options)
+  ifelse(options$size != "normalsize",
+         paste0("\n \\", options$size, "\n\n", x,
+                "\n\n \\normalsize"), x)
+})
 
-# Conclusion: New Avenues for Research
+# Install packages
+packages <- c("kableExtra", "scales", "tidyverse")
+installed_packages <- packages %in% rownames(installed.packages())
+if (any(installed_packages == FALSE)) {
+  install.packages(packages[!installed_packages])
+}
+invisible(lapply(packages, library, character.only = TRUE))
 
-In this paper, we discuss a quantitative presentation of the Report of the Chilean National Commission on Truth and Reconciliation [-@report1991], which described more than two thousand human rights abuses perpetrated by the Pinochet regime from 1973 to 1990.
+# Read dataset
+library(readr)
+df <- read_csv("../data/victims.csv")
+```
 
-The dataset includes rich information about more than 2,0000 victims of the Pinochet regime, the geographical location of the incidents, and the specific date when the violations occurred.
-The graphs and maps shown in this article provide some preliminary results about the temporal and spatial variation of state-sponsored violence during the recent military period in Chile.
+[^auth01]: Author information. 
+[^auth02]: Independent Researcher,
+[`danilofreire@gmail.com`](mailto:danilofreire@gmail.com),
+<https://danilofreire.github.io>.
+[^auth03]: Author information.
 
-We believe our data opens a number of potential research avenues.
-First, scholars can use the dataset to evaluate the long-term impacts of human rights violations on several outcomes of interest.
-For example, XXX and XXX highlight that state-led violence has enduring effects on social polarisation, political outcomes, and economic development.
-Researchers can test whether the Pinochet regime caused a similar impact on Chile's economy and society.
-Moreover, sociologists and criminologists may be interested in the relationship between the intensity of the human rights abuse and post-regime levels of social unrest or interpersonal violence.
+\newpage
 
-Second, the dataset enables researchers to analyse the inner dynamics of authoritarian governments.
-This topic remains understudied in political science due to a lack of disaggregated information.
-By linking human right abuses to changes in Pinochet's support coalition, scholars can explore whether the violations affected levels of civilian or bureaucratic collaboration to the regime.
-XXXXXXXXXXXXX.
+# Introduction
+\label{sec:intro}
 
-Finally, scholars can investigate the connections between international legitimacy and domestic politics in repressive regimes.
-This is a promising area of research as the Chilean government and American intelligence services continue to declassify documents from the Pinochet era.
-One relevant question is whether pressure from foreign governments and organisations had any influence over the levels of human rights abuses in Chile.
-We hope our dataset is useful for scholars interested in these and other questions, and that the information it contains elicits hypotheses not only about the Pinochet period, but about authoritarian governments more generally.
+General Augusto Pinochet ruled Chile with an iron fist from 1973 to 1990. Later
+findings show that the Chilean Armed Forces and the government's secret police
+carried out gross human rights violations against political dissidents,
+including murders, beatings, waterboarding, and kidnappings [@ensalaco2000chile;
+@huneeus2007pinochet; @verdugo2001chile]. Most of the repression occurred in the
+first months of the military regime, but state terrorism continued until the
+last years of the dictatorship, often in collaboration with other Latin American
+countries [@dinges2005condor; @mcsherry2012predatory]. While the true extension
+of the brutality of the Pinochet regime remains unknown, estimates suggests that
+the number of victims may be close to 30,000 [@weissbrodt1992report].
 
-[^df]: Postdoctoral Research Fellow, The Political Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912, [`danilofreire@gmail.com`](mailto:danilofreire@gmail.com), [`http://danilofreire.github.io`](http://danilofreire.github.io).
-Corresponding author.
+In this paper, we provide a quantitative presentation of all cases of deaths and
+disappearances recorded in the Rettig and Valech reports. Both reports were
+elaborated at the request of the Chilean government and are the official sources
+of information on serious human rights abuses during the Pinochet years. The
+National Commission for Truth and Reconciliation Report, also known as the
+Rettig Report after its main editor Raúl Rettig, was compiled in 1991 and
+recorded more than 2,000 cases of murder and kidnappings from 1973 to 1990
+[@rettig1991report]. In 2004, a commission chaired by Bishop Sergio Valech
+produced another compilation of victims of the military dictatorship, which
+complemented and expanded the original findings [@valech2004report]. The Valech
+report was updated in 2010 to include more cases.
 
-[^jm]: Senior Lecturer in Public Policy, Department of Political Economy, King's College London, [`john.meadowcroft@kcl.ac.uk`](mailto:john.meadowcroft@kcl.ac.uk), [`http://johnmeadowcroft.net`](http://johnmeadowcroft.net).
-
-[^ds]: Associate Professor, Department of Political Science and the Political Theory Project, Brown University, [`david_skarbek@brown.edu`](david_skarbek@brown.edu), [`http://davidskarbek.com`](http://davidskarbek.com).
-
-[^eg]: Software Developer, Tempo UK.
+Our dataset contains `r ncol(df)` variables and `r comma(nrow(df))`
+observations. The data describes personal characteristics of the victims -- such
+as unique id, name, age, gender, and occupation -- as well as information about
+the crimes, including the type, perpetrator, and geolocation whenever such
+information was available in the reports. We formatted the dataset following the
+principles of "tidy data", in which each row correspondents to one case, each
+variable forms a column, and the data are represented as a table
+[@wickham2014tidy]. Moreover, we provide three different presentations of the
+data. The first table includes only the cases mentioned in Rettig Report, the
+second is the data from the Valech Report, and in the final table we merged all
+cases together. To make the data easily available to users, we provide a
+companion `R` package, named `pinochet` [@freire2019pinochet], alongside `csv`
+and `xlsx` files that scholars can download at
+<http://github.com/danilofreire/pinochet>.
 
@@
 \setlength{\parindent}{0cm}
-\setlength{\parskip}{6pt}
+\setlength{\parskip}{5pt}

### 2021-03-01 R/pinochet.R

@@ -1,80 +1,0 @@
-#' Human rights abuses in the Pinochet regime, 1973-1990
-#'
-#' A dataset containing information about victims of the Pinochet regime
-#' in Chile (1973-1990). The data comes from the National Truth and Reconciliation
-#' Commission Report (1991, ISBN:9780268016463). We included georeferenced information about all incidents
-#' where the information could be retrieved. More information about the coding process
-#' is available at http://github.com/danilofreire/pinochet.
-#'
-#' @docType data
-#'
-#' @usage data(pinochet)
-#'
-#' @format A data frame with 2398 rows and 59 variables:
-#' \describe{
-#'   \item{additional_comments}{any further information on the victim}
-#'   \item{age}{victim's age}
-#'   \item{end_date_daily}{calendar day when the human rights abuse ended}
-#'   \item{end_date_monthly}{calendar month when human rights abuse ended}
-#'   \item{exact_coordinates_1}{exactitude of coordinates for place_1, 1 = exact}
-#'   \item{exact_coordinates_2}{exactitude of coordinates for place_2, 1 = exact}
-#'   \item{exact_coordinates_3}{exactitude of coordinates for place_3, 1 = exact}
-#'   \item{exact_coordinates_4}{exactitude of coordinates for place_4, 1 = exact}
-#'   \item{exact_coordinates_5}{exactitude of coordinates for place_5, 1 = exact}
-#'   \item{exact_coordinates_6}{exactitude of coordinates for place_6, 1 = exact}
-#'   \item{first_name}{victim's first name}
-#'   \item{group_id}{unique id for groups of individuals attacked together}
-#'   \item{individual_id}{unique individual id}
-#'   \item{interrogation}{individual was interrogated, 1 = yes}
-#'   \item{last_name}{victim's last name}
-#'   \item{latitude_1}{latitudinal coordinates pertaining to the string place_1 variable}
-#'   \item{latitude_2}{latitudinal coordinates pertaining to the string place_2 variable}
-#'   \item{latitude_3}{latitudinal coordinates pertaining to the string place_3 variable}
-#'   \item{latitude_4}{latitudinal coordinates pertaining to the string place_4 variable}
-#'   \item{latitude_5}{latitudinal coordinates pertaining to the string place_5 variable}
-#'   \item{latitude_6}{latitudinal coordinates pertaining to the string place_6 variable}
-#'   \item{location_1}{where the individual was seen or found}
-#'   \item{location_2}{where the individual was seen or found}
[…]
